### 2022-02-03 ms_body/_ms_template/svm-rmarkdown-article-example.tex
First big commit

\documentclass[12pt,]{article}
\usepackage[left=1in,top=1in,right=1in,bottom=1in]{geometry}
\newcommand*{\authorfont}{\fontfamily{phv}\selectfont}
\usepackage[]{mathpazo}

  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}

\usepackage{abstract}
\renewcommand{\abstractname}{}    % clear the title
\renewcommand{\absnamepos}{empty} % originally center

\renewenvironment{abstract}
 {{%
    \setlength{\leftmargin}{0mm}
    \setlength{\rightmargin}{\leftmargin}%
  }%
  \relax}
 {\endlist}

\makeatletter
\def\@maketitle{%
  \newpage
%  \null
%  \vskip 2em%
%  \begin{center}%
  \let \footnote \thanks
    {\fontsize{18}{20}\selectfont\raggedright  \setlength{\parindent}{0pt} \@title \par}%
}
%\fi
\makeatother

\setcounter{secnumdepth}{0}

\author{}

\date{}

\usepackage{titlesec}

\titleformat*{\section}{\normalsize\bfseries}
\titleformat*{\subsection}{\normalsize\itshape}
\titleformat*{\subsubsection}{\normalsize\itshape}
\titleformat*{\paragraph}{\normalsize\itshape}
\titleformat*{\subparagraph}{\normalsize\itshape}

\newtheorem{hypothesis}{Hypothesis}
\usepackage{setspace}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{hyperref}
\setlength\parindent{24pt}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

% move the hyperref stuff down here, after header-includes, to allow for - \usepackage{hyperref}

\makeatletter
\@ifpackageloaded{hyperref}{}{%
\ifxetex
  \PassOptionsToPackage{hyphens}{url}\usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \PassOptionsToPackage{hyphens}{url}\usepackage[draft,unicode=true]{hyperref}
\fi
}

\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
             pdfkeywords = {},  
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

% Add an option for endnotes. -----

% add tightlist ----------
\providecommand{\tightlist}{%
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% add some other packages ----------

% \usepackage{multicol}
% This should regulate where figures float
% See: https://tex.stackexchange.com/questions/2275/keeping-tables-figures-close-to-where-they-are-mentioned
\usepackage[section]{placeins}

\begin{document}

% \pagenumbering{arabic}% resets `page` counter to 1 
%

\vskip -8.5pt

 % removetitleabstract

\noindent \doublespacing

\pagebreak

\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}
\setlength{\parskip}{8pt}
\vspace*{-0.2in}

\noindent

\textbf{Cryptic diversity and host-specificity of \emph{Tetramesa}
(Hymenoptera: Eurytomidae) associated with two African grasses:
implications for biological control}

Sutton, G. F.\textsuperscript{1}\(\dag\), van Steenderen,
C.\textsuperscript{1}, Canavan, K\textsuperscript{1}, Day, M.
D.\textsuperscript{2}, Paterson, I. D.\textsuperscript{1}

~

\begingroup
\fontsize{10}{12}\selectfont

\textsuperscript{1} Center for Biological Control, Department of Zoology
and Entomology, Rhodes University, Makhanda, 6140, South Africa

\textsuperscript{2} Department of Agriculture and Fisheries, GPO Box
267, Brisbane, Australia \endgroup

~

\textbf{Corresponding author} \(\dag\)
\href{mailto:g.sutton@ru.ac.za}{\nolinkurl{g.sutton@ru.ac.za}}

\pagebreak

\setlength\parindent{24pt}

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

\emph{Sporobolus pyramidalis} Beauv. and \emph{Sporobolus natalensis}
(Steud.) Th. Dur. \& Schinz. (Poaceae) (giant rat tail grass), are two
perennial African grasses that have become invasive in Australia. Field
host-range surveys on 47 non-target grasses in the weeds native
distribution (South Africa), between 2017 and 2019, found an apparently
specialised and highly damaging stem-boring wasp, \emph{Tetramesa} sp.
1. Subsequent field surveys in late 2019 found a morphologically
indistinguishable \emph{Tetramesa} sp. on a closely-related non-target
grass, \emph{Eragrostis curvula} (Schrad.) Nees (African lovegrass). If
the wasps reared from the different host-plants (\emph{S. pyramidalis}
or \emph{E. curvula}) represent the same genetic entity, this would
preclude its use as a biological control agent. In this study, we showed
that two morphologically cryptic \emph{Tetramesa} sp. populations
collected from \emph{E. curvula} and \emph{S. pyramidalis},
respectively, represent two distinct genetic clades, likely representing
distinct species. Cross-inoculation trials indicated that each wasp
could survive on the host plant off which it was collected. Our findings
indicate that practitioners need to be aware of cryptic species when
conducting field host-range surveys. Morphologically indistinguishable
candidate agents might in fact be crytic species with their own
restricted host ranges, and if we do not take this into account, then we
may exclude potentially effective and host-specific agents.

~

\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}
\setlength{\parskip}{8pt}
\vspace*{-0.2in}

\noindent

\textbf{Keywords}: cryptic species; DNA barcoding; \emph{Eragrostis
curvula}; giant rat's tail grass'; invasive grass; \emph{Sporobolus}

\pagebreak
\setlength\parindent{24pt}

\hypertarget{introduction}{%
\section{1. Introduction}\label{introduction}}

\emph{Sporobolus pyramidalis} Beauv.

Recent estimates indicate that there are approximately 7 million insect
species on earth, albeit only about 20\% have been described to date
(Stork, 2018). There are strong geographic and taxonomic biases in
sampling coverage and intensity, with Afrotropical Hymenoptera being
particularly under-studied and poorly known (van Noort et al., 2015).
There are approximately 205 described \emph{Tetramesa} species known,
worldwide (Al-Barrak et al., 2004), although only four species have been
described in the Afroptropical region, none of which are known from
South Africa (van Noort, 2019). Our study reports on two potentially
novel \emph{Tetramesa} species from South Africa. While these wasps have
not been formally described yet, these specimens have been sent to an
expert taxonomist for an official scientific description, but our
\emph{COI} barcoding and rearing trials provide strong support for the
species status of both \emph{Tetramesa} identified during the current
study. An additional five putative \emph{Tetramesa} species have been
identified during our field surveys in South Africa (G.F. Sutton,
unpublished data), and are currently undergoing further study. Surveys
for potential biological control agents of weeds provide an exciting
opportunity for documenting and discovering previously unknown
biodiversity, which is particularly relevant for surveys performed in
Africa and other relatively understudied geographic regions.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\vspace*{-0.2in}

\noindent

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{CSLReferences}%
{\setlength{\parindent}{0pt}%
\everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
{\par}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Al-Barrak2004}{}%
Al-Barrak, M., Loxdale, H.D., Brookes, C.P., Dawah, H.A., Biron, D.G.,
Alsagair, O., 2004. Molecular evidence using enzyme and {RAPD} markers
for sympatric evolution in {British} species of {\emph{Tetramesa}}
({Hymenoptera}: {Eurytomidae}). Biological Journal of the Linnean
Society 83, 509--525.
\url{https://doi.org/10.1111/j.1095-8312.2004.00408.x}

\leavevmode\hypertarget{ref-Stork2018}{}%
Stork, N.E., 2018. How many species of insects and other terrestrial
arthropods are there on earth? Annual Review of Entomology 63, 31--45.

\leavevmode\hypertarget{ref-VanNoort2015}{}%
van Noort, S., Buffington, M.L., Forshage, M., 2015. Afrotropical
{Cynipoidea} ({Hymenoptera}). ZooKeys 1--176.
\url{https://doi.org/10.3897/zookeys.493.6353}

\end{CSLReferences}

\newpage

\hypertarget{figure-and-table-legend}{%
\section{Figure and table legend}\label{figure-and-table-legend}}

\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\vspace*{-0.2in}

\noindent

\textbf{Table 1} Results from no-choice cross-inoculation trials.

\textbf{Table 2} Results from no-choice greenhouse-based host-range
trials for \emph{Tetramesa} sp. reared from \emph{Sporobolus
pyramidalis} and \emph{S. natalensis}.

\textbf{Figure 1} Genetics of \emph{Tetramesa} spp.

\newpage

\newpage
\singlespacing 
\end{document}

### 2022-02-03 ms_body/manuscript.Rmd
First big commit

---
output:
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    latex_engine: pdflatex
    template: ../svm-latex-ms.tex
  html_document:
    df_print: paged
  word_document: default
header-includes:
- \usepackage{hyperref}
- \setlength\parindent{24pt}
- \usepackage{multirow}
- \usepackage{helvet}
- \usepackage{setspace}
- \renewcommand\familydefault{\sfdefault}
biblio-style: apsr
geometry: margin=1in
fontfamily: mathpazo
bibliography: refs.bib
csl: biological-control.csl
fontsize: 12pt
spacing: double
indent: yes
endnote: no
include-before: '`\newpage{}`{=latex}'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE,
                      message=FALSE, 
                      warning=FALSE,
                      fig.path='figs/',
                      cache.path = '_cache/',
                      fig.process = function(x) {
                      x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
                      if (file.rename(x, x2)) x2 else x
                      })

# Load required packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, 
               tidyr,
               readr,
               here,
               gt,
               ggtext,
               gtsummary,
               kableExtra)
```

\pagebreak

\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}
\setlength{\parskip}{8pt}
\vspace*{-0.2in}
\noindent

**Testing MaxEnt model performance in a novel geographic region using an intentionally introduced insect**

Sutton, G. F.^1^$\dag$; Martin, G.D. ^1,^^2^

\

\begingroup
\fontsize{10}{12}\selectfont
^1^ Center for Biological Control, Department of Zoology and Entomology, Rhodes University, Makhanda, 6140, South Africa

^2^ Afromontane Research Unit and Department Zoology and Entomology, University of the Free State, Phuthaditjhaba, 9866, South Africa
\endgroup

\

**Corresponding author** $\dag$ g.sutton@ru.ac.za

\pagebreak

\setlength\parindent{24pt}

# Abstract

Ecological models are important tools to support and guide the development and implementation of environmental policies and management programs. Species distributions models (SDM's), such as the popular MaxEnt software, are frequently used to guide conservation programmes, predict the potential distribution of invasive species, forecast the impacts of climate change and develop applied ecological research (e.g. biological control). Many applications of these models require that SDM's are transferred in either space and/or time. However, few studies to date have tested the transferability, and thus usefulness, of MaxEnt models using fully independent data. Moreover, numerous authors have raised concerns over how model complexity (controlled primarily by feature class and regularisation multiplier settings) may affect MaxEnt model transferability. In this paper, we evaluated the usefulness of MaxEnt models transferred in space using a native Australian insect, *Dasineura rubiformis* Kolesik (Diptera: Cecidomyiidae) that has been intentionally introduced as a biological control agent of the invasive plant, *Acacia mearnsii* De Wild (Fabaceae) in South Africa. MaxEnt models were developed using native-range records only (Australia) and projected over South Africa to identify the potential climatic suitability for the insect, using a range of model settings configurations. Model transferability was assessed using independent post-release data from South Africa. Our results demonstrated that MaxEnt scores were positively correlated with increased establishment rates for *D. rubiformis* in South Africa. However, we found that this pattern held only when MaxEnt models were specified with user-defined settings configurations. A MaxEnt model calibrated with default settings was overfit to the training data, and performed poorly when transfered in space. Our study demonstrates that MaxEnt models can prove useful when transferred in space, however, user-optimised model settings likely need to be used to control model complexity, and reduce the likelihood of overfitting.

 \

\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}
\setlength{\parskip}{8pt}
\vspace*{-0.2in}
\noindent

**Keywords**: Species distribution model, biological control, *Acacia*, *Dasineura rubiformis*, model complexity, model transferability

\setlength\parindent{24pt}

# 1. Introduction

Ecological models are important tools to support and guide the development and implementation of environmental policies and management programs [@Addison2013; @Schuwirth2019]. For example, ecological models can be used for planning conservation programs [@Guisan2013], predicting the establishment of invasive species [@Martin2020], implementing biological control programmes [@Mukherjee2021], or forecasting species responses to environmental change [@Bocedi2014]. Species distribution models (SDM’s) are an example of ecological models that have become increasingly popular in recent years [@Elith2009]. SDM’s typically take the form of correlative or mechanistic models that relate species presence/absences to environmental covariates to map habitats that may be suitable for a taxon [@Elith2011]. The Maximum Entropy Species Distribution Model (hereafter ‘MaxEnt’) [@Phillips2017] is among the most popular methods for species distribution modeling and has been shown to perform well compared to alternative modeling algorithms [@Elith2006]. It uses maximum entropy to distinguish between environmental conditions where the focal taxon is present from environmental conditions at background locations where the taxons' presence is absent (or assumed absent) [@Elith2011].

Despite its popularity, the usefulness of MaxEnt has been questioned when used to generate predictions for novel geographic regions (transferred in space) or time periods (transferred in time) from those for which the data used to build the model were collected [@Yates2018; @Blasi2021]. To date, studies assessing transferability have produced contrasting results [@Randin2006; @Bahn2013; @Capinha2018]. Model transferability has typically been assessed by cross validation, whereby a fraction of the data is withheld to calibrate the model (i.e. training data) and the remaining data is retained to validate the model (i.e. the testing data) [@Wenger2012]. However, cross validation may not provide an estimate of model transferability as a selected test sample may not provide an unbiased estimate of the full data set, and therefore may not reflect the heterogeneity that may exist in the underlying data in space or time [see @Wenger2012]. Ideally, a direct test of model transferability should use an independent dataset of species presence/absences and environmental covariates [@Fielding1997; @West2016]. Unfortunately, very few studies, to date, have tested model transferability using independent datasets [e.g. @Costa2010; @Rebelo2010], which may have limited our understanding of how useful SDM’s can be when transferred in space or time. Issues with model transferability may lead to misleading conclusions, which may hinder scientists abilities to address important ecological questions (e.g. where could an invasive species establish?, where should a conservation priority be translocated?) [@Merow2013].

Model complexity can affect uncertainty of model transferability in space and time [@Warren2014]. Model complexity may be affected by numerous parameters, notably the feature classes and regularisation (beta) multiplier settings used to specify the MaxEnt model [@Elith2011; @Merow2013]. MaxEnt allows users to specify combinations of different feature classes that modify the shape and complexity of species-environment functions. Moreover, MaxEnt uses a regularisation multiplier (a form of L~1~-regularisation) to balance model complexity with the model fit, penalising unnecessarily complex models [@Merow2013]. The default MaxEnt settings use all feature classes in model fitting except 'threshold' features (depending on sample size of training data) and a regularisation multiplier of 1 [@Phillips2017]. Several recent studies have highlighted the importance of species-specific tuning of feature classes and regularisation multipliers for optimising model complexity [@Shcheglovitova2013; @Radosavljevic2014], with default settings often producing models that are overly complex, and thus, may perform poorly when transferred [@Elith2011; @Merow2013]. Despite the potential implications, surprisingly few studies use user-defined feature class and regularisation settings [e.g. @Sutton2019b], and as such, little is known about how model settings configurations may influence model transferability [@Moreno-Amat2015; @Low2021].

In this study, we performed a retrospective assessment of the transferability of SDM’s built using MaxEnt [@Phillips2017]. Here, we modelled the potential climatic suitability of an insect that is native to Australia, and that has been introduced into South Africa as a biological control agent. The insect, *Dasineura rubiformis* Kolesik (Diptera: Cecidomyiidae) was introduced into South Africa from Australia in 2013, to curb the spread of the invasive Australian *Acacia mearnsii* De Wild [@Impson2013]. Extensive field surveys have been performed in Australia (characterizing the native distribution of *D. rubiformis*) [@Adair2004] and long-term post-release monitoring has been performed in South Africa to measure establishment rates of the insect, following experimental releases of the insect (Impson et al., 2021). As such, the availability of independent testing data from a separate continent to the data used to generate the models makes this an ideal study system for assessing the transferability of SDM’s.

We built MaxEnt models for *D. rubiformis* using distribution records from its native distribution (Australia), and projected the resulting MaxEnt suitability rasters over South Africa. MaxEnt scores range from 0 to 100, whereby 0 represents regions predicted to be highly unsuitable and 100 indicating highly suitable regions. To test model transferability, we fit the MaxEnt scores as a covariate in a logistic regression model to determine whether MaxEnt scores were statistically relevant predictors of insect establishment rates in South Africa (with these data being entirely independent of the data used to build the MaxEnt models). A strong, positive correlation between MaxEnt scores and insect establishment rates would provide support that the SDM’s were useful when transferred in space (from Australia to South Africa). Moreover, we tested whether model complexity and parameter tuning could improve the accuracy and transferability of MaxEnt by comparing the predictive power of MaxEnt models calibrated with default settings and an array of optimised settings configurations.

# 2. Methods and Materials

## 2.1. Species occurences records

A total of 103 native-range records for *D. rubiformis* were obtained from [@Adair2004]. A total of 452 invaded-range records to evaluate model transferability were taken from Impson, Kleinjan and Hoffmann (unpublished). Distribution maps for *D. rubiformis* for both native and invaded range are provided in Fig. 1.

Spatial autocorrelation is an important factor that may affect model outputs. Filtering of species occurrence data may limit the inherent biases in the data and improve model quality [@Veloz2009]. To avoid pseudo-replication, only one occurrence record per 2.5 minute grid cell was used for model calibration. Species occurrence datasets were thinned using the ‘spThin’ package [@Aiello-Lammens2015], and spatial autocorrelation analyses were performed using the ‘ecospat’ package [@DiCola2017].

## 2.2. Environmental predictors

Climate data were obtained by downloading the standard set of 19 bioclimatic variables from the WorldClim ver. 1.4 database [@Hijmans2005] (data available at: www.worldclim.org/download.html). This dataset is representative of annual and seasonal means and
variation of temperature and precipitation metrics averaged over the 1950–2000 time period (current climate) at a 5 arc minute resolution.

To reduce multicollinearity between environmental predictors and thereby decrease the risk of calibrating overfit models, Pearson’s correlation coefficients were computed for all pairs of predictors, whereby predictors which were highly correlated (|*r*| > 0.85) were excluded from the final predictor set [@Capinha2011]. The reduced set of environmental predictors consisted of seven climatic variables, including: bio1 – annual mean temperature, bio2 – mean
diurnal temperature range (mean of monthly [max temp – min temp], bio3 – isothermality, bio9 – mean temperature of the driest quarter, bio12 – mean annual precipitation, bio14 – precipitation of the driest month and bio16 – precipitation of the wettest quarter (see @Hijmans2005 for further details).

## 2.3. Model calibration

Maxent was implemented in the `dismo` package in `R` [@hijmans2017package]. MaxEnt was selected as it consistently outperforms other modelling algorithms [@Wisz2008]. Given that MaxEnt is a presence-only modelling algorithm, model calibration requires a user-defined geographic background to sample the climate of representative grid cells where the focal species is absent (i.e. background points). Background definition can have a significant effect on model output [@VanDerWal2009]. The background should ideally represent the geographic areas available to the focal species, omitting areas where species absence is due to historical factors, dispersal constraints and/or biotic interactions [@Sanin2018]. Following @Webber2011, we defined the model background using the Koppen-Geiger climate classification (Available at: http://koeppen-geiger.vu-wien.ac.at). Only Koppen-Geiger climate zones that contained at least one native-range occurrence record for *D. rubiformis* were used as the background area from which pseudo-absences were drawn for model calibration. Koppen-Geiger climate zones were intersected using the `extract` function from the `R` package `raster` [@hijmans2017package].

## 2.4. Optimal model settings

Model performance and optimal settings configurations were assessed using multiple metrics that reflected aspects of model (1) discriminatory ability (AUC~test~), (2) overfitting (AUC~diff~), (3) omission rates (OR~10~), and (4) overall parsimony (AICc) [@Low2021].

(1) AUC~test~ assesses the models ability to discriminate between predicted presence/absence at withheld portions of the data used to test the model versus background points. The area under the receiver operating characteristic curve (AUC) is one of the most popular metrics used to evaluate MaxEnt models. An AUC of less than 0.8 is considered a poor model, between 0.8 and 0.9 is a fair model, between 0.9 and 0.995 a good model, and > 0.995 an excellent model [@Fielding1997]. Thus, higher AUC~test~ values indicate increased ability to discriminate between training and background points.

(2) AUC~diff~ calculates the difference between AUC values calculated on training points only (AUC~train~) and AUC~test~ using cross-validation [@Warren2011]. Thus, higher AUC~diff~ values indicate whether the MaxEnt model is overfit on the training data, and thus, performs poorly when evaluated against testing points.

(3) OR~10~ is the proportion of testing points that are not predicted to fall within the projected model surface once the model is converted into a binary prediction output [@Boria2014]. Overfit models have omission rates higher than the theoretical expectation for the threshold applied [@Shcheglovitova2013]. OR10 sets the binary prediction threshold at a value that excludes the 10% of the calibration localities from the model with the lowest prediction values, and therefore has an expected omission rate of 0.10 [@Boria2014]. As such, the OR~10~ criterion selected models calibrated with MaxEnt settings which best approximated the expected 0.10 omission rate. Models with omission rates increasingly higher than the expected value were considered as more overfit [@Boria2017].

(4) Lastly, optimal model settings were determined by selecting model configurations which produced the lowest value for the Akaike Information Criterion corrected for small sample sizes (AICc) (i.e. AICc=0; following @Muscarella2014). The AICc criterion simultaneously scores models according to their complexity and goodness-of-fit, whereby models with the lowest AICc are selected as the best models. AICc was used as the primary evaluation metric as it is calculated using MaxEnt models built using the entire species occurrence dataset, unlike AUC and OR~10~ (and numerous other metrics frequently used for model evaluation) which may be spatially biased due to the partitioning of the species occurrence dataset into training and evaluation sets [@Sanin2018].

## 2.5. Assessing model extrapolation

Several authors have raised concerns over the use of correlative SDM's, such as MaxEnt, when projected into new geographic regions or time periods due to issues with extrapolation into novel/non-analogous climate [@Elith2009; @Elith2011; @Yates2018]. To address these concerns, Multivariate environmental similarity surfaces (hereafter ‘MESS’) [@Elith2010] were computed to assess whether MaxEnt models were extrapolating or interpolating. MESS analyses measure the similarity of any given point to a set of reference points [reference points were defined as only the species presence points used to calibrate models following @Kriticos2014]. Negative MESS values indicate geographic regions outside the range of climate variables used to calibrate the model (i.e. extrapolation space or MESS-), while MESS values between 0 and 100 indicate geographic regions inside the range of climatic variables used to calibrate the model (i.e. interpolation space or MESS+). MESS maps are used as a measure of prediction uncertainty or caution against inferences in extrapolation space [@Elith2010].

## 2.6. Statistical analyses

Five candidate MaxEnt models were specified using native-range (Australia) training data. The five models differed in the feature classes and regularisation multiplier settings used by MaxEnt to build the models. The five models were specified with (1) default MaxEnt settings, and four optimised settings configurations that maximised performance based on (2) AUC~test~, (3) AUC~diff~, (4) OR~10~ and (5) AICc.

To evaluate the predictive performance and transferability of these models, we evaluated whether MaxEnt climatic suitability scores were correlated with increased probabilities of *D. rubiformis* establishment in South Africa. To do so, MaxEnt scores were specified as a continuous fixed effect and whether *D. rubiformis* was recorded or not (present/absent) as a boolean response variable in a logistic general linear model (GLM), using a logit link function. MaxEnt scores, bounded between 0 (not climatically suitable) and 1 (highly climatically suitable) indicating climatic similarity between the native range occupied by *D. rubiformis* in Australia and sites where it was released as a biological control agent in South Africa, were derived by projecting MaxEnt suitability rasters from models calibrated using native-range GPS records only (Australian native-range records) over South Africa, and extracting the MaxEnt suitability scores for each site where *D. rubiformis* was surveyed in South Africa using the `extract` function from the `R` package `raster` [@hijmans2017package]. GLM's were specified using the `rms` package in `R` [@Harrell2017package]. Likelihood Ratio Tests (LRT) were performed to determine whether there was a statistical correlation between MaxEnt scores and the probability of *D. rubiformis* establishment. A statistically significant, positive correlation between MaxEnt scores and insect establishment rates would provide support that the SDM’s were useful and predictive of insect establishment, even when the insect was transferred in space. Out-of-sample AUC was calculated using the `rms` package, using 500 bootstrap replicates [@Harrell2017package] to quantify the predictive performance of MaxEnt models, whereby AUC values less than 0.80 are typically considered poor fitting models (lacks predictive power) [@Fielding1997].

A similar approach was adopted to test if each occurrence record in South Africa was classified as being in interpolation (within the climate space occupied by the insect in its native range) versus extrapolation (outside the climate space occupied by the insect in its native range) was predictive of *D. rubiformis* establishment. Interpolation/extrapolation were quantified using MESS maps using the `mess` function from the `dismo` package in `R` [@hijmans2017package]. Interpolation versus extrapolation was fit as a categorical predictor and whether *D. rubiformis* was recorded or not (present/absent) as a boolean response variable in a logistic GLM, using a logit link function.

All modelling and statistical analyses were conducted in `R` ver. 4.0.3 [@RCoreTeam2020].

# 3. Results

## 3.1. Native-range models (model calibration)

Five candidate MaxEnt models were developed using native-range (Australia) training data for *D. rubiformis*. The five models differed in the feature classes and regularisation multiplier settings used by MaxEnt to build the models. The five models were specified with (1) default MaxEnt settings, and four optimised settings configurations that (2) maximised performance based on AUC~test~, (3) AUC~diff~, (4) OR~10~ and (5) AICc (Table S1). The AUC values for these five models ranged were ranged between 0.82 and 0.97, depending on the MaxEnt settings configuration used. As such, all five candidate models  were considered useful and had high predictive accuracy in distinguishing between *D. rubiformis* recorded presences and background points in its native range. While default MaxEnt settings produced the highest discriminatory power when tested in the training area of Australia (AUC = 0.97), models calibrated with tuned settings configurations led to less overfitting (lower AUC~diff~), better omission rates (best approximated the expectation for OR~10~), and more parsimonious models (AICc) (Fig. 2).

## 3.2. Model transferability

When evaluated on independent testing data in South Africa, the MaxEnt models showed relatively high, but variable, predictive accuracy, depending on the MaxEnt settings used to calibrate the model. For all models, the probability that *D. rubiformis* was recorded at a field site in South Africa increased with higher MaxEnt climatic suitability scores (Fig. 3). The default MaxEnt model found no statistically significant relationship between MaxEnt climatic suitability scores and the probability of *D. rubiformis* establishment in South Africa (Table. 1), and displayed low power to discriminate between sites where *D. rubiformis* had established or not (AUC: 0.64 $\pm$ 0.01) (Fig. 4). In contrast, the four models calibrated using optimised model settings configurations all demonstrated a statistically significant positive relationship between MaxEnt suitability scores and the probability of *D. rubiformis* establishment in South Africa (Table. 1). All four tuned models displayed high power to discriminate between sites where *D. rubiformis* had established or not (AUC: 0.83 - 0.84) than the MaxEnt model calibrated with default settings (AUC: 0.64 $\pm$ 0.01) (Fig. 4). The predicted climatic suitability for *D. rubiformis* across South Africa varied depending on the MaxEnt settings used to calibrate the model (Fig. 5). Default MaxEnt settings yielded higher climatic suitability estimates for *D. rubiformis* than the other MaxEnt models with tuned model settings, notably in regions where *D. rubiformis* is absent (e.g. the north-west coastal region bordering Namibia).

Despite the high performance of the MaxEnt models when transferred in space, MESS maps indicated that the MaxEnt models were extrapolating climatic suitability scores for *D. rubiformis* over the vast majority of South Africa (Fig. S1). Indeed, 54 of the 78 field sites surveyed for *D. rubiformis* in South Africa occur in geographic regions where the MaxEnt models were extrapolating. *Dasineura rubiformis* established at 23 of 24 sites in interpolation (96%), while it established at 44 of 54 sites in extrapolation (82%) (Fig. 6). We found no evidence that a site being in extrapolation/interpolation was a statistically significant predictor of *D. rubiformis* establishment in South Africa (X~2~ = 3.40, d.f = 1, *P* = 0.07). Moreover, classifying sites as being in extrapolation/interpolation yielded lower power to discriminate to discriminate between sites where *D. rubiformis* had established or not (AUC: 0.67 $\pm$ 0.02) than all of the MaxEnt models, barring the model with default settings.

# 4. Discussion

Species distribution models (SDM's) are becoming increasingly important tools to support and guide the development and implementation of environmental policies and management programs [@Addison2013; @Schuwirth2019]. The usefulness of SDM's is dependent on model performance, which is typically measured by its ability to correctly predict where a species is established and its capacity to distinguish between sites where the species is established versus a site where the species is absent (or assumed to be absent) [@Smith2021]. However, a number of authors have called in question the usefulness of SDM's, notably when models are transferred in space or time [@Yates2018; @Blasi2021]. Poor transferability may lead to erroneous model predictions that may have serious negative consequences for the application of the model, and resulting policies and management programmes (Soultan and Safi, 2017). Here, we report on the validation of a SDM for an insect, (*D. rubiformis*), native to Australia, that has been intentionally introduced as a biological control agent into South Africa to reduce the invasiveness of its host-plant, *A. mearnsii* (Impson et al., 2013). The the availability of independent testing data from South Africa, a separate continent to the data used to generate the models (Australia), makes this an ideal study system for assessing the transferability of SDM’s. This study represents one of only a handful of studies to validate SDM's using independent testing data [@Costa2010; @Rebelo2010; @West2016; @Smith2021].

Our results indicated that MaxEnt SDM's were able to provide robust estimates of climatic suitability for *D. rubiformis* in South Africa. AUC values greater than 0.80 are typically considered good fitting models (having high predictive power) [@Fielding1997]. When MaxEnt models, calibrated using native-range records for *D. rubiformis* from Australia, were projected over South Africa, they had AUC scores of 0.83 - 0.84, indicative of high power to discriminate between sites where *D. rubiformis* had established in South Africa or not. Additionally, we showed that the probability that *D. rubiformis* was recorded at a field site in South Africa increased with higher MaxEnt climatic suitability scores. This result demonstrates the usefulness of the MaxEnt SDM and that, at least for our case-study, SDM's can be transferred into novel geographic regions, and provide valuable information to guide policy making and the development of management programmes. For example, biological control programmes such as the programme in which *D. rubiformis* is used to curtail the spread of the highly invasive weed *A. mearnsii* expend many resources (time and money) to mass-rear insect populations that are then released into the field. The positive correlation between MaxEnt suitability scores and *D. rubiformis* establishment means that researchers can develop and implement mass-rearing and experimental release programmes for the insect that prioritise releasing the insect at sites that are highly climatically suitable, and thus, more likely the insect will establish viable populations. This may assist in improving the level of control over the target weed and increasing the efficiency of resources expended by increasing the chances of establishing viable insect populations.

Our study adds to a growing body of literature indicating that SDM model complexity affects model performance and transferability when being projected into novel geographic regions. While the default MaxEnt model had the highest predictive power in the training area (Australia), this model performed extremely poorly when transferred over South Africa. In contrast, the four MaxEnt models calibrated with tuned settings showed high predictive power when transferred. The poor transferability of the Default model is likely explained by default settings overfitting the model on the training points [@Warren2014]. This finding corroborates other studies showing that high predictive power in the model training area is not a guarantee of transferability [@Warren2014; @Duque-Lazo2016]. We echo that recommendations of @Shcheglovitova2013 and @Muscarella2014 that model tuning should be performed to estimate optimal model complexity when transferring MaxEnt models in space and/or time in order to maximise their usefulness.

An additional concern with transferring MaxEnt models in space or time is the potential for models to significantly extrapolate predictions when the model is projected into non-analogous conditions to those under which the model was calibrated (Elith et al., 2010; Mesgaran et al., 2014). However, the applied use of SDM's in developing environmental policy, developing management programmes for invasive species and implementing conservation programmes often necessitate that models are transferred and predictions made, even when models are extrapolating (Elith et al., 2010). Here, we showed that *D. rubiformis* was able to establish at many sites (82%) in South Africa that were outside the range of climate space it occupies in its native range (i.e. extrapolation), which was not substantially different from establishment rates at sites in interpolation. While our MaxEnt models were able to make accurate predictions of *D. rubiformis* establishment in South Africa, even when models were extrapolating, this finding in no way discounts the importance of assessing model extrapolation when transferring models in space or time. A number of tools have been developed in recent years to assist SDM users with identifying model extrapolation, e.g. MESS [@Elith2010] and ExDet (Mesgaran et al., 2014), and which should be an integral component of any SDM study.

Given that the *D. rubiformis* MaxEnt models showed high predictive accuracy when transferred in space over South Africa and the statistically significant positive correlation between MaxEnt scores and *D. rubiformis* establishment, we wish to highlight how these models can be used in an applied setting. To visualise the potential geographic regions where *D. rubiformis* may establish in South Africa, we thresholded the MaxEnt map that optimised AICc where the probability of establishment was > 50% (i.e. MaxEnt score = 0.30), given in Fig. 9. The map demonstrates that the entire south and east coast regions of South Africa are climatically suitable for *D. rubiformis*. Mass-releasing and releasing weed biological control agents is a time-consuming and costly endeavor (Coetzee et al., 2021). As such, our models provide a statistically validated approach to optimising the redistribution of *D. rubiformis* in South Africa by prioritising releases at field sites where MaxEnt scores are > 0.30, and therefore where the probability of establishment was > 50%.

# 5. Conclusions

Species distribution models are becoming increasingly popular with researchers, stakeholders, policy makers and land managers alike, particularly for making predictions in novel geographic geographic regions or forecasting species responses to climate change. Transfering SDM's in space and/or time, however, has been met with justifiable resilience and concerns over how well these models perform upon transfer. In this study, we presented a case-study demonstrating that MaxEnt SDM's were able to accurately predict the distribution of an Australian insect when introduced into South Africa. Our study is one of only a few studies to date validating MaxEnt models using spatially independent data. However, the accuracy of the MaxEnt models was dependent on the settings used to calibrate the models, with default MaxEnt programme settings having low predictive power. Our results demonstrate that MaxEnt can be useful in predicting where a species may establish, even when projected into a novel geographic area, but that users need to be aware of how model settings can influence model outputs and usefulness. Demonstrating how SDM's can be useful, while communicating uncertainty in their predictions and outputs, is essential for building confidence in the use of SDM's (Blasi et al. 2020). We showed how validating predictive models can be used to guide the implementation of a biological control programme by optimising the redistribution and release of *D. rubiformis* in geographic areas that are climatically suitable and where the probability of the insect establishing is greater than 50%.

# Acknowledgements

GFS and GDM acknowledge funding from the South African Working for Water (WfW) programme of the Department of Forestry, Fisheries and the Environment: Natural Resource Management Programmes (DFFE: NRMP). Funding was also provided by the South African Research Chairs Initiative of the Department of Science and Technology and the National Research Foundation (NRF) of South Africa. Any opinions, finding, conclusions or recommendations expressed in this material are those of the authors and the NRF does not accept any liability in this regard.

\newpage

# References

\setlength{\parindent}{-0.0in}
\setlength{\leftskip}{0.0in}
\setlength{\parskip}{8pt}
\vspace*{-0.0in}
\noindent

\newlength{\cslhangindent}
\newenvironment{CSLReferences}%
{\setlength{\parindent}{0pt}%
\everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
{\par}

<div id="refs" ></div>

\newpage

# Figure and table legend

\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\vspace*{-0.2in}
\noindent

\newpage

## *Figure 1*

```{r import_fig_1, echo = FALSE}
knitr::include_graphics(here::here("./ms_body/ms_figs/fig_1.png"), dpi = NA)
```

\newpage

## *Figure 2*

```{r import_fig_2, echo = FALSE}
knitr::include_graphics(here::here("./ms_body/ms_figs/fig_2.png"), dpi = NA)
```

\newpage

## *Figure 3*

```{r import_fig_3, echo = FALSE}
knitr::include_graphics(here::here("./ms_body/ms_figs/fig_3.png"), dpi = NA)
```

\newpage

## *Figure 4*

```{r import_fig_4, echo = FALSE}
knitr::include_graphics(here::here("./ms_body/ms_figs/fig_4.png"), dpi = NA)
```

\newpage

## *Figure 5*

```{r import_fig_5, echo = FALSE}
knitr::include_graphics(here::here("./ms_body/ms_figs/fig_5.png"), dpi = NA)
```

\newpage

## *Figure 6*

```{r import_fig_6, echo = FALSE}
knitr::include_graphics(here::here("./ms_body/ms_figs/fig_6.png"), dpi = NA)
```

\newpage

**Table 1**

 \

```{r table_1, echo = FALSE, results = "asis"}
# Import data for Table 1
table1_data <- readr::read_csv(here::here("./data/data_clean/table1_data.csv"))

# Add footnote
table1_footnote <- table1_data
names(table1_footnote)[2] <- paste0(names(table1_footnote)[2], 
                                    footnote_marker_alphabet(1))

# Make Table 1
table1 <- table1_footnote %>% 
  knitr::kable(
    format = "latex",
    align = "l",
    booktabs = TRUE,
    longtable = TRUE,
    linesep = "",
    escape = FALSE,
    col.names = c(
      "Model",
      "$\\beta$\\textsuperscript{a}", 
      "$\\chi$\\textsubscript{2}",
      "d.f.",
      "\\textit{P}"
    )
    ) %>%
  kableExtra::kable_styling(
      position = "left",
      latex_options = c("striped", "repeat_header")
      #,       stripe_color = "gray!15"
    ) 
#table1

# Add footnote
table1 <- table1 %>%
    footnote(
    alphabet = c("Adjusted to indicate change in Y for each 1% unit increase in covariate"),
    escape = TRUE,
    footnote_as_chunk = FALSE,
    threeparttable = TRUE) 
table1
```

\newpage

**Table S1**

 \

```{r table_S1, echo = FALSE}
# Import data to make Table S1
tableS1_data <- readr::read_csv(here::here("./data/data_clean/table2_data.csv")) %>%
  janitor::clean_names()

# Create Table S1
tableS1 <- tableS1_data %>%
  knitr::kable(
    format = "latex",
    align = "l",
    booktabs = TRUE,
    longtable = TRUE,
    linesep = "",
    escape = FALSE,
    col.names = c(
      "Model",
      "RM",
      "FC",
      "AUC\\textsubscript{test}",
      "AUC\\textsubscript{diff}",
      "OR\\textsubscript{10}"
    )
  ) %>%
  kableExtra::kable_styling(
    position = "left",
    latex_options = c("striped", "repeat_header")
    #,    stripe_color = "gray!15"
  )

# Display Table S1
tableS1
```

### 2022-02-03 R/functions/function_convert_GPS.R

@@ -1,0 +1,45 @@
+# Custom function to convert DMS to decimal degrees
+# I can't remember where I found this function online.
+# - Please contact me if you are the author or know who the author is 
+#   so that I can give you the appropriate credit. 
+
+# Converting coordinates from DMS or DdM formats to decimal
+# DMS = "degrees minutes seconds"; DdM = "degrees decimal minutes" 
+# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+dg2dec <- function(varb, Dg=NA, Min=NA, Sec=NA, SW.Hemisphere="S|W") {
+  # Dg=decimal, Min=minutes and Sec=seconds; 
+  # NOTE 1 - if the format is "degrees decimal minutes - DdM" (e.g. 40?? 26.767?? N) and not 
+  # "degrees minutes seconds - DMS" (e.g. 40?? 26?? 46?? N), then call the function only with 
+  # Dg and Min arguments, like dg2dec(varb, Dg="??", Min="??N"). 
+  # Same applies when there is no seconds symbol (e.g. 45??12'7.38).
+  # Note that one should not use blank spaces in Dg, Min or Sec arguments (will return NA).
+  # For more details on formats see: 
+  # https://en.wikipedia.org/wiki/Geographic_coordinate_conversion#Coordinate_format_conversion
+  
+  # Use paste0("[", Dg, Min, Sec, "]") to build regex [] pattern
+  # therefore, strsplit() will split string "varb" by what symbols you give to Dg, Min, Sec
+  DMS <- sapply(strsplit(varb, paste0('[', Dg, Min, Sec, ']')), as.numeric)
+  
+  # DMS is a matrix; first row contains degrees; second - minutes; third - seconds.
+  # If the format is "degrees decimal minutes" (e.g. 40?? 26.767?? N) and not 
+  # "degrees minutes seconds" (e.g. 40?? 26?? 46?? N), then the matrix has only two valid rows: 
+  # first row contains degrees; the second - minutes;
+  # therefore, compute conversion for seconds only if there are more than 2 rows in DMS 
+  # and Sec is different from NA (if there are seconds in the DMS format)
+  decdg <- abs(DMS[1, ]) + DMS[2, ]/60 + ifelse(dim(DMS)[1] > 2  & !is.na(Sec), DMS[3, ]/3600, 0)
+  
[…]

### 2022-02-03 R/scripts/01-clean-invaded-range-gps.R

@@ -1,0 +1,95 @@
+###########################################################################
+# Script: 01 - Clean invaded range records  -------------------------------
+#
+# AUTHOR: Guy F. Sutton
+# AFFILIATION: Centre for Biological Control, Rhodes University, South Africa
+# DATE MODIFIED: 03/02/2021
+# CONTACT: g.sutton@ru.ac.za
+###########################################################################
+
+###########################################################################
+# Setup -------------------------------------------------------------------
+###########################################################################
+
+# Load required packages
+if (!require("pacman"))
+  install.packages("pacman")
+pacman::p_load(tidyverse,
+               tidyr,
+               DHARMa,
+               here,
+               readxl,
+               glmmTMB)
+
+# Change ggplot theme
+theme_set(
+  theme_classic() +
+    theme(
+      panel.border = element_rect(colour = "black",
+                                  fill = NA),
+      axis.text = element_text(colour = "black"),
+      axis.title.x = element_text(margin = unit(c(2, 0, 0, 0),
+                                                "mm")),
+      axis.title.y = element_text(margin = unit(c(0, 4, 0, 0),
+                                                "mm")),
+      legend.position = "none"
+    )
+)
+
+# Load source functions
+# - Load in function to process GPS records
+source(here::here("./R/functions/function_convert_GPS.R"))
+
+###########################################################################
+# Load data ---------------------------------------------------------------
+###########################################################################
+
+# Load invaded-range GPS points
+raw_data <-
+  readxl::read_xlsx(
+    here::here(
+      "./data/data_raw/dasineura_rubiformis_invaded_range_database_cleaned.xlsx"
+    )
+  )
+head(raw_data)
+
+###########################################################################
+# Process data ------------------------------------------------------------
+###########################################################################
+
+# Remove NA values in latitude column
+raw_data <- raw_data %>%
+  tidyr::drop_na(lat2)
+raw_data
+
+# Need to convert the decimal minutes into decimal degrees
+data_proc <- raw_data %>%
+  dplyr::mutate(
+    lat_converted = dg2dec(
+      varb = raw_data$lat2,
+      Dg = "°",
+      Min = "'"
+    ),
+    lon_converted = dg2dec(
+      varb = raw_data$lon2,
+      Dg = "°",
+      Min = "'"
+    )
+  )
+head(data_proc)
+
[…]

### 2022-02-03 R/scripts/02-maxent-native-range.R

@@ -1,0 +1,430 @@
+# -----------------------------------------------------------------------------
+# Script: 02 - MaxEnt models - native-range (Australia)  
+#
+# AUTHOR: Guy F. Sutton
+# AFFILIATION: Centre for Biological Control, Rhodes University, South Africa
+# DATE MODIFIED: 03/02/2021
+# CONTACT: g.sutton@ru.ac.za
+# -----------------------------------------------------------------------------
+
+# -----------------------------------------------------------------------------
+# Session setup
+# -----------------------------------------------------------------------------
+
+# Load required packages
+if (!require("pacman"))
+  install.packages("pacman")
+pacman::p_load(
+  tidyverse,
+  dismo,
+  raster,
+  here,
+  corrplot,
+  Hmisc,
+  patchwork,
+  ecospat,
+  gridSVG,
+  gridExtra,
+  grid,
+  ENMeval,
+  spThin,
+  viridis,
+  viridisLite,
+  mapdata,
+  maptools,
+  scales,
+  geosphere,
+  rgdal,
+  ggtext,
+  rJava,
+  rgeos,
+  sp,
+  sf,
+  ggspatial,
+  ecospat,
+  rnaturalearth,
+  rnaturalearthdata
+)
+
+# Change ggplot theme
+theme_set(
+  theme_classic() +
+    theme(
+      panel.border = element_rect(colour = "black",
+                                  fill = NA),
+      axis.text = element_text(colour = "black"),
+      axis.title.x = element_text(margin = unit(c(2, 0, 0, 0),
+                                                "mm")),
+      axis.title.y = element_text(margin = unit(c(0, 4, 0, 0),
+                                                "mm")),
+      legend.position = "none"
+    )
+)
+
+# Set the theme for the maps
+theme_opts <- list(
+  theme(
+    panel.grid.minor = element_blank(),
+    panel.grid.major = element_blank(),
+    panel.background = element_rect(fill = 'white', colour = NA),
+    plot.background = element_rect(),
+    axis.line = element_blank(),
+    axis.text.x = element_text(colour = "black"),
+    axis.text.y = element_text(colour = "black"),
+    axis.ticks = element_line(colour = "black"),
+    axis.title.x = element_text(colour = "black"),
+    axis.title.y = element_text(colour = "black"),
+    plot.title = element_text(colour = "black"),
+    panel.border = element_rect(fill = NA),
+    legend.key = element_blank()
+  )
+)
+
+# Load source functions
[…]

### 2022-02-03 R/scripts/02-maxent-native-range.R
Add model tuning results .csv file

@@ -234,7 +234,6 @@ plot(bgPoints,
 bgPoints <- raster::as.data.frame(bgPoints)
 head(bgPoints)
 
-
 # --------------------------------------------------------------------------
 # Reduce multicollinearity between environmental predictors:
 #
@@ -328,7 +327,8 @@ speciesThinned <- spThin::thin(
 )
 
 # Import the thinned GPS records
-speciesThinned <- readr::read_csv(here::here("./data/data_clean/dasi_rubi_native_thinned.csv"))
+speciesThinned <-
+  readr::read_csv(here::here("./data/data_clean/dasi_rubi_native_thinned.csv"))
 head(speciesThinned)
 
 # How many records were removed?
@@ -337,7 +337,6 @@ nrow(speciesThinned)
 
 # Just for ease, make species_thinned = species
 species <- speciesThinned
-
 
 # --------------------------------------------------------------------
 # Set up data to run MaxEnt models:
@@ -422,7 +421,45 @@ modelTuning <- ENMevaluate(
   algorithm = "maxent.jar"
 )
 
+# Extract the model tuning results 
+# - Pretty messy code but meh, it works. 
 
+# Extract all the variables and statistics to identify optimal RM/FC 
+settings <- modelTuning@results$settings
+feat.class <- modelTuning@results$features
+features <- modelTuning@results$features
+rm <- modelTuning@results$rm
+AUC.calibrate <- modelTuning@results$train.AUC
+AUC.test <- modelTuning@results$avg.test.AUC
+AUC.diff <- modelTuning@results$avg.diff.AUC
+OR.10 <- modelTuning@results$avg.test.or10pct
+OR.min <- modelTuning@results$avg.test.orMTP
+AICc <- modelTuning@results$AICc
+delta.AIC <- modelTuning@results$delta.AICc
+parameters <-  modelTuning@results$parameters
+modelEvaluation <- as.data.frame(cbind(settings, 
+                                        rm, 
+                                        AUC.calibrate, 
+                                        AUC.test, 
+                                        AUC.diff, 
+                                        OR.10, 
+                                        OR.min, 
+                                        AICc, 
+                                        delta.AIC,
+                                        parameters))
+modelEvaluation$feat.class <- as.factor(feat.class)
+modelEvaluation$rm <- as.numeric(rm)
[…]

### 2022-02-03 R/scripts/02-maxent-native-range.R
Add model tuning experiment plots

@@ -423,8 +423,6 @@ modelTuning <- ENMevaluate(
 
 # Extract the model tuning results 
 # - Pretty messy code but meh, it works. 
-
-# Extract all the variables and statistics to identify optimal RM/FC 
 settings <- modelTuning@results$settings
 feat.class <- modelTuning@results$features
 features <- modelTuning@results$features
@@ -462,6 +460,71 @@ readr::write_csv(
 )
 
 
+# Plot AUCtest
+plot_aucTest <- ggplot(data = modelEvaluation, aes(x = rm, 
+                                                y = AUC.test,
+                                                group = feat.class)) +
+  geom_point(aes(colour = feat.class)) + 
+  geom_line(aes(colour = feat.class)) +
+  scale_colour_grey(start = 0.4, end = 0.8) +
+  #scale_y_continuous(breaks = seq(0.80, 1.00, 0.04),
+  #                   limits = c(0.80, 1.00)) +
+  labs(x = "Regularisation multiplier",
+       y = expression(paste(AUC[test])),
+       subtitle = "(a)") 
+plot_aucTest
 
+# Plot AUCdiff
+plot_aucDiff <- ggplot(data = modelEvaluation, aes(x = rm, 
+                                                y = AUC.diff,
+                                                group = feat.class)) +
+  geom_point(aes(colour = feat.class)) + 
+  geom_line(aes(colour = feat.class)) +
+  scale_colour_grey(start = 0.4, end = 0.8) +
+  #scale_y_continuous(breaks = seq(0.80, 1.00, 0.04),
+  #                   limits = c(0.80, 1.00)) +
+  labs(x = "Regularisation multiplier",
+       y = expression(paste(AUC[diff])),
+       subtitle = "(b)",
+       colour = "Feature class") +
+  theme(legend.position = c(0.8, 0.7))
+plot_aucDiff
+
+# Plot 10th percentile omission rates (OR10)
+plot_or10 <- ggplot(data = modelEvaluation, aes(x = rm, 
+                                            y = OR.10,
+                                            group = feat.class)) +
+  geom_point(aes(colour = feat.class)) + 
+  geom_line(aes(colour = feat.class)) +
+  scale_colour_grey(start = 0.4, end = 0.8) +
+  geom_hline(yintercept = 0.10, linetype = "dashed") + 
+  labs(x = "Regularisation multiplier",
[…]

### 2022-02-03 R/scripts/02-maxent-native-range.R
Add default settings MaxEnt model

@@ -527,4 +527,59 @@ plots_modelTuning <- cowplot::plot_grid(
 )
 plots_modelTuning
 
+# Extract settings that maximised AUCtest
+eval_tab <- modelEvaluation %>%
+  dplyr::arrange(desc(AUC.test)) %>%
+  dplyr::slice(1)
+eval_tab
 
+# Extract settings that minimised AUCdiff
+eval_tab <- modelEvaluation %>%
+  dplyr::arrange(AUC.diff) %>%
+  dplyr::slice(1)
+eval_tab
+
+# Extract settings that best approximated OR10
+eval_tab <- modelEvaluation %>%
+  dplyr::arrange(OR.10) %>%
+  dplyr::slice(1)
+eval_tab
+
+# Extract settings that minimised AICc
+eval_tab <- modelEvaluation %>%
+  dplyr::arrange(delta.AIC) %>%
+  dplyr::slice(1)
+eval_tab
+
+
+# -----------------------------------------------------------------------------
+# Run full MaxEnt models:
+# -----------------------------------------------------------------------------
+
+# Model 1: Default MaxEnt settings
+# - The default model for n = 77 has rm = 1 and HLQ features.
+model_Default <- dismo::maxent(
+  x = trainData,
+  p = presentBg,
+  path = paste(getwd(),
+               './models/models/optimal_settings_default',
+               sep = ''),
+  args = c(
+    'betamultiplier=1.0',
+    'linear=true',
+    'quadratic=true',
+    'product=false',
+    'threshold=false',
+    'hinge=true',
+    'threads=2',
+    'doclamp=true',
+    #'fadebyclamp=true',
+    'responsecurves=true',
+    'jackknife=true',
+    'askoverwrite=false',
+    'responsecurves=true',
+    'writemess=true',
+    'writeplotdata=true',
+    'writebackgroundpredictions=true'
+  )
+)

### 2022-02-03 R/scripts/02-maxent-native-range.R
Update Default model code flow

@@ -551,7 +551,6 @@ eval_tab <- modelEvaluation %>%
   dplyr::slice(1)
 eval_tab
 
-
 # -----------------------------------------------------------------------------
 # Run full MaxEnt models:
 # -----------------------------------------------------------------------------

### 2022-02-04 R/scripts/02-maxent-native-range.R
Set seed for model settings evaluation

@@ -291,7 +291,7 @@ corData <- speciesEnv[ , redPred] %>%
 
 # Combine background points with species GPS data 
 spatialData <- dplyr::bind_cols(corData,
-                                 species) 
+                                species) 
 head(spatialData)
 
 # At what distance does spatial autocorrelation occur?
@@ -391,7 +391,6 @@ presentBg <- ifelse(dummyTraindata$presence=='present', 1, 0)
 trainData <- rbind(speciesSwd[ , climPred], 
                    bgPoints[ , climPred])
 
-
 # -----------------------------------------------------------------------------
 # Tune models using the 'ENMeval' package
 # -----------------------------------------------------------------------------
@@ -405,7 +404,8 @@ trainData <- rbind(speciesSwd[ , climPred],
 reducedPred <- raster::subset(predictors, climPred)
 
 # Run evaluation
-# - This can take > 1 hour on my terrible PC
+# - This can take 30 minutes on my terrible PC
+set.seed(2012)
 modelTuning <- ENMevaluate(
   occ = species[, 2:3],
   env = reducedPred,
@@ -459,104 +459,130 @@ readr::write_csv(
   file = here::here("./models/model_tuning/model_tuning_results.csv")
 )
 
-
 # Plot AUCtest
-plot_aucTest <- ggplot(data = modelEvaluation, aes(x = rm, 
-                                                y = AUC.test,
-                                                group = feat.class)) +
-  geom_point(aes(colour = feat.class)) + 
+plot_aucTest <- ggplot(data = modelEvaluation, aes(x = rm,
+                                                   y = AUC.test,
+                                                   group = feat.class)) +
+  geom_point(aes(colour = feat.class)) +
   geom_line(aes(colour = feat.class)) +
   scale_colour_grey(start = 0.4, end = 0.8) +
   #scale_y_continuous(breaks = seq(0.80, 1.00, 0.04),
   #                   limits = c(0.80, 1.00)) +
   labs(x = "Regularisation multiplier",
        y = expression(paste(AUC[test])),
-       subtitle = "(a)") 
+       subtitle = "(a)")
 plot_aucTest
 
 # Plot AUCdiff
-plot_aucDiff <- ggplot(data = modelEvaluation, aes(x = rm, 
-                                                y = AUC.diff,
[…]

### 2022-02-04 R/scripts/03-logistic-regression.R

@@ -1,0 +1,824 @@
+# -----------------------------------------------------------------------------
+# Script: 03 - Logistic regression
+#            - Do MaxEnt scores correlate with higher insect establishment?
+#
+# AUTHOR: Guy F. Sutton
+# AFFILIATION: Centre for Biological Control, Rhodes University, South Africa
+# DATE MODIFIED: 03/02/2021
+# CONTACT: g.sutton@ru.ac.za
+# -----------------------------------------------------------------------------
+
+# -----------------------------------------------------------------------------
+# Session setup
+# -----------------------------------------------------------------------------
+
+# Load required packages
+if (!require("pacman"))
+  install.packages("pacman")
+pacman::p_load(
+  tidyverse,
+  spocc,
+  ggmap,
+  maptools,
+  maps,
+  ggplot2,
+  scrubr,
+  mapr,
+  tidyr,
+  stringr,
+  rnaturalearth,
+  rnaturalearthdata,
+  rlang,
+  sf,
+  ggspatial,
+  raster,
+  rms,
+  here,
+  tidytext,
+  ggtext,
+  tidymodels,
+  themis,
+  ranger,
+  yardstick,
+  vip,
+  DHARMa,
+  rms
+)
+
+# Define function to calculate uncertainty in out-of-sample metrics
+boot_val <- function(data, formula, ...) {
+  out <- list()
+  for (i in 1:500) {
+    df <- sample_n(data, nrow(data), replace = TRUE)
+    md <- glm(formula, data = df, family = binomial)
+    out[[i]] <- val.prob(predict(md, type = "response"),
+                         as.numeric(df$established) - 1,
+                         pl = FALSE) %>% round(3)
+  }
+  return(out)
+}
+
+# ----------------------------------------------------------------------------
+# Evaluate performance of default settings model: 
+# ----------------------------------------------------------------------------
+
+# Import Maxent prediction raster 
+clim_map <- raster::raster("./models/rasters/raster_default.asc")
+
+# Import invaded-range GPS points 
+inv_data <- readr::read_csv("./data/data_clean/dasi_rubi_invaded_range_gps_clean.csv")
+head(inv_data)
+
+# Keep just the columns of interest
+inv_clean <- inv_data %>%
+  dplyr::select(lon = lon_converted,
+                lat = lat_converted,
+                established) %>%
+  mutate(lat = lat * -1)
+inv_clean
+
+# Keep only the sites with establishment data
[…]

### 2022-02-04 R/scripts/02-maxent-native-range.R
Update MaxEnt models with reproducible settings

@@ -610,7 +610,7 @@ model_Default <- dismo::maxent(
 )
 
 # Model 2: Optimise AUC.test
-# - The top performing model by AUC.test had rm = 2 and all features (HLPQT).
+# - The top performing model by AUC.test had rm = 6 and hinge features only.
 # Fit model with optimally tuned settings based on AICc
 model_AUCtest <- dismo::maxent(
   x = trainData,
@@ -619,7 +619,35 @@ model_AUCtest <- dismo::maxent(
                './models/models/optimal_settings_AUCtest',
                sep = ''),
   args = c(
-    'betamultiplier=2.0',
+    'betamultiplier=6.0',
+    'linear=false',
+    'quadratic=false',
+    'product=false',
+    'threshold=false',
+    'hinge=true',
+    'threads=2',
+    'doclamp=true',
+    #'fadebyclamp=true',
+    'responsecurves=true',
+    'jackknife=true',
+    'askoverwrite=false',
+    'responsecurves=true',
+    'writemess=true',
+    'writeplotdata=true',
+    'writebackgroundpredictions=true'
+  )
+)
+
+# Model 3: Optimise AUC.diff
+# - The top performing model by AUC.diff had rm = 4 and HLQPT features.
+model_AUCdiff <- dismo::maxent(
+  x = trainData,
+  p = presentBg,
+  path = paste(getwd(),
+               './models/models/optimal_settings_AUCdiff',
+               sep = ''),
+  args = c(
+    'betamultiplier=4',
     'linear=true',
     'quadratic=true',
     'product=true',
@@ -638,37 +666,8 @@ model_AUCtest <- dismo::maxent(
   )
 )
 
-# Model 3: Optimise AUC.diff
-# - The top performing model by AUC.diff had rm = 1.5 and only H features.
-model_AUCdiff <- dismo::maxent(
-  x = trainData,
-  p = presentBg,
-  path = paste(getwd(),
-               './models/models/optimal_settings_AUCdiff',
-               sep = ''),
-  args = c(
-    'betamultiplier=1.5',
-    'linear=false',
-    'quadratic=false',
-    'product=false',
-    'threshold=false',
-    'hinge=true',
-    'threads=2',
-    'doclamp=true',
-    #'fadebyclamp=true',
-    'responsecurves=true',
-    'jackknife=true',
-    'askoverwrite=false',
-    'responsecurves=true',
-    'writemess=true',
-    'writeplotdata=true',
-    'writebackgroundpredictions=true'
-  )
[…]

### 2022-02-04 R/scripts/03-logistic-regression.R
Make Figure 6

@@ -32,7 +32,6 @@ pacman::p_load(
   sf,
   ggspatial,
   raster,
-  rms,
   here,
   tidytext,
   ggtext,
@@ -110,19 +109,8 @@ print(mod_lrm_default)
 summary(mod_lrm_default)
 
 # Calculate beta for 0.01 unit increase in clim_match
-beta = 7.53 # Input the effect from clim_match row here 
+beta = 4.65 # Input the effect from clim_match row here 
 exp(beta * 0.01)
-
-# Validate model on out-of-sample AUC
-d <- clim_data %>%
-  dplyr::select(clim_match, established)
-y <- clim_data %>%
-  dplyr::select(established)
-pred.logit <- predict(mod_lrm_default, d)
-phat <- 1/(1+exp(-pred.logit))
-v <- val.prob(phat, 
-              y[1:77, ])
-print(v)
 
 # Calculate uncertainty
 set.seed(48572)
@@ -168,7 +156,6 @@ ndata <- bind_cols(ndata, setNames(as_tibble(predict(mod,
                                    c("fit_link", "se_link")))
 head(ndata)
 
-
 # Create the confidence interval and back-transform
 ndata <- ndata %>%
   dplyr::mutate(
@@ -187,12 +174,12 @@ logit_default <- ggplot(data = ndata) +
               alpha = 0.2) +
   geom_line(aes(x = clim_match, 
                 y = fit_resp)) +
-  geom_rug(data = subset(clim_data, established == 1),
-           aes(x = clim_match),
-           sides = "t") +
-  geom_rug(data = subset(clim_data, established == 0),
-           aes(x = clim_match),
-           sides = "b") +
+  #geom_rug(data = subset(clim_data, established == 1),
+  #         aes(x = clim_match),
+  #         sides = "t") +
+  #geom_rug(data = subset(clim_data, established == 0),
+  #         aes(x = clim_match),
+  #         sides = "b") +
   coord_cartesian(clip = "off") + 
   labs(x = "Maxent suitability score \n(climatic similarity to native range)",
        y = expression(paste("Probability of ",italic("D. rubiformis "), establishment)),
@@ -257,19 +244,8 @@ print(mod_lrm_AUCtest)
 summary(mod_lrm_AUCtest)
 
 # Calculate beta for 0.01 unit increase in clim_match
-beta = 9.02 # Input the effect from clim_match row here 
+beta = 2.76 # Input the effect from clim_match row here 
 exp(beta * 0.01)
-
[…]

### 2022-02-04 R/scripts/figure-6-make.R

@@ -1,0 +1,398 @@
+# -----------------------------------------------------------------------------
+# Script: - Reproduce Figure 6
+#
+# AUTHOR: Guy F. Sutton
+# AFFILIATION: Centre for Biological Control, Rhodes University, South Africa
+# DATE MODIFIED: 03/02/2021
+# CONTACT: g.sutton@ru.ac.za
+# -----------------------------------------------------------------------------
+
+# -----------------------------------------------------------------------------
+# Session setup
+# -----------------------------------------------------------------------------
+
+# Load packages
+if (!require("pacman"))
+  install.packages("pacman")
+pacman::p_load(
+  tidyverse,
+  spocc,
+  ggmap,
+  maptools,
+  maps,
+  ggplot2,
+  scrubr,
+  mapr,
+  tidyr,
+  stringr,
+  patchwork,
+  rnaturalearth,
+  rnaturalearthdata,
+  rlang,
+  sf,
+  ggspatial,
+  raster,
+  here,
+  tidytext,
+  ggtext,
+  tidymodels,
+  themis,
+  ranger,
+  viridis,
+  yardstick,
+  vip,
+  DHARMa,
+  rms
+)
+
+# Set the theme for the figures
+theme_set(
+  theme_classic() +
+    theme(
+      panel.border = element_rect(colour = "black", fill = NA),
+      axis.text = element_text(colour = "black"),
+      axis.title.x = element_text(margin = unit(c(2, 0, 0, 0), "mm")),
+      axis.title.y = element_text(margin = unit(c(0, 4, 0, 0), "mm")),
+      legend.position = "none"
+    )
+)
+
+# Set the theme for the maps
+theme_opts <- list(
+  theme(
+    panel.grid.minor = element_blank(),
+    panel.grid.major = element_blank(),
+    panel.background = element_rect(fill = 'white', colour = NA),
+    plot.background = element_rect(),
+    axis.line = element_blank(),
+    axis.text.x = element_text(colour = "black"),
+    axis.text.y = element_text(colour = "black"),
+    axis.ticks = element_line(colour = "black"),
+    axis.title.x = element_text(colour = "black"),
+    axis.title.y = element_text(colour = "black"),
+    plot.title = element_text(colour = "black"),
+    panel.border = element_rect(fill = NA),
+    legend.key = element_blank()
+  )
+)
+
+# ----------------------------------------------------------------------------
+# Plot default map: 
[…]

### 2022-02-05 R/scripts/figure-1-make.R
Make Figure 1

@@ -1,0 +1,219 @@
+# -----------------------------------------------------------------------------
+# Script: - Reproduce Figure 1
+#
+# AUTHOR: Guy F. Sutton
+# AFFILIATION: Centre for Biological Control, Rhodes University, South Africa
+# DATE MODIFIED: 03/02/2021
+# CONTACT: g.sutton@ru.ac.za
+# -----------------------------------------------------------------------------
+
+# -----------------------------------------------------------------------------
+# Session setup
+# -----------------------------------------------------------------------------
+
+# Load required packages
+if (!require("pacman"))
+  install.packages("pacman")
+pacman::p_load(
+  tidyverse,
+  spocc,
+  ggmap,
+  maptools,
+  maps,
+  ggplot2,
+  scrubr,
+  mapr,
+  tidyr,
+  stringr,
+  rnaturalearth,
+  rnaturalearthdata,
+  rlang,
+  sf,
+  ggspatial,
+  raster,
+  here,
+  tidytext,
+  cowplot,
+  tidymodels,
+  themis,
+  ranger,
+  yardstick,
+  vip,
+  DHARMa
+)
+
+# Set theme for ggplot
+theme_set(
+  theme_classic() +
+    theme(
+      panel.border = element_rect(colour = "black", fill = NA),
+      axis.text = element_text(colour = "black"),
+      axis.title.x = element_text(margin = unit(c(2, 0, 0, 0), "mm")),
+      axis.title.y = element_text(margin = unit(c(0, 4, 0, 0), "mm")),
+      legend.position = "none"
+    )
+)
+
+# ----------------------------------------------------------------------------
+# Make invaded-range map:
+# ----------------------------------------------------------------------------
+
+# Load invaded-range GPS points
+data <-
+  readr::read_csv("./data/data_clean/dasi_rubi_invaded_range_gps_clean.csv")
+
+# Clean records
+data <- data %>%
+  janitor::clean_names() %>%
+  dplyr::select(lat_converted, lon_converted, established) %>%
+  # Make lat negative to be in southern hemisphere
+  dplyr::mutate(lat_converted = lat_converted * -1) %>%
+  dplyr::filter(lat_converted < -23)
+head(data)
+
+
+# Get world map
+world <- rnaturalearth::ne_countries(scale = "medium",
+                                     returnclass = "sf")
+
+# Keep only South Africa
+world <- world %>%
+  dplyr::filter(name %in% c("South Africa"))
+
+# Classify sites as established (1) or absent (0)
[…]

### 2022-02-05 R/scripts/figure-6-make.R

@@ -392,7 +392,7 @@ all_maps
 # Save the plot to your PC
 ggsave(
   "./ms_body/ms_figs/fig_6_maxent_maps.png",
-  width = 14,
-  height = 10,
+  width = 18,
+  height = 12,
   dpi = 600
 )

### 2022-02-06 R/scripts/02-maxent-native-range.R
Make Figure 2

@@ -295,36 +295,36 @@ spatialData <- dplyr::bind_cols(corData,
 head(spatialData)
 
 # At what distance does spatial autocorrelation occur?
-spatialCorAnalysis <- ecospat::ecospat.mantel.correlogram(
-  # Data frame with environmental variables
-  dfvar = spatialData,
-  # Columns containing lat/long
-  colxy = 9:10,
-  # Number of random occurrences
-  n = 100,
-  # Climate variables
-  colvar = 1:8,
-  max = 1000,
-  nclass = 100,
-  nperm = 100
-)
+#spatialCorAnalysis <- ecospat::ecospat.mantel.correlogram(
+#  # Data frame with environmental variables
+#  dfvar = spatialData,
+#  # Columns containing lat/long
+#  colxy = 9:10,
+#  # Number of random occurrences
+#  n = 100,
+#  # Climate variables
+#  colvar = 1:8,
+#  max = 1000,
+#  nclass = 100,
+#  nperm = 100
+#)
 
 # Add a species column to the GPS data
-species <- species %>%
-  dplyr::mutate(species = "D. rubiformis")
+#species <- species %>%
+#  dplyr::mutate(species = "D. rubiformis")
 
 # Thin by spatial autocorrelation value
-speciesThinned <- spThin::thin(
-  loc.data = species,
-  lat.col = "lat",
-  long.col = "lon",
-  spec.col = "species",
-  # Km unit of correlogram
-  thin.par = 0.2,
-  reps = 100,
-  max.files = 1,
-  out.dir = here::here("./data/data_clean/")
-)
+#speciesThinned <- spThin::thin(
+#  loc.data = species,
+#  lat.col = "lat",
+#  long.col = "lon",
+#  spec.col = "species",
+#  # Km unit of correlogram
+#  thin.par = 0.2,
+#  reps = 100,
+#  max.files = 1,
+#  out.dir = here::here("./data/data_clean/")
+#)
 
 # Import the thinned GPS records
 speciesThinned <-
@@ -667,7 +667,7 @@ model_AUCdiff <- dismo::maxent(
 )
 
 # Model 4: Optimise OR10
-# - The top performing model by OR10 had rm = 1 and only HLQPT features.
+# - The top performing model by OR10 had rm = 6 and only HLQPT features.
 model_OR10 <- dismo::maxent(
   x = trainData,
   p = presentBg,
@@ -675,7 +675,7 @@ model_OR10 <- dismo::maxent(
[…]

### 2022-02-06 R/scripts/figure-2-make.R

@@ -1,0 +1,148 @@
+# -----------------------------------------------------------------------------
+# Script: - Reproduce Figure 2
+#
+# AUTHOR: Guy F. Sutton
+# AFFILIATION: Centre for Biological Control, Rhodes University, South Africa
+# DATE MODIFIED: 03/02/2021
+# CONTACT: g.sutton@ru.ac.za
+# -----------------------------------------------------------------------------
+
+# -----------------------------------------------------------------------------
+# Session setup
+# -----------------------------------------------------------------------------
+
+# Load required packages
+if (!require("pacman"))
+  install.packages("pacman")
+pacman::p_load(
+  tidyverse,
+  spocc,
+  ggmap,
+  maptools,
+  maps,
+  ggplot2,
+  scrubr,
+  mapr,
+  tidyr,
+  stringr,
+  rnaturalearth,
+  rnaturalearthdata,
+  rlang,
+  sf,
+  ggspatial,
+  raster,
+  here,
+  tidytext,
+  cowplot,
+  tidymodels,
+  themis,
+  ranger,
+  yardstick,
+  vip,
+  DHARMa
+)
+
+# Set theme for ggplot
+theme_set(
+  theme_classic() +
+    theme(
+      panel.border = element_rect(colour = "black", fill = NA),
+      axis.text = element_text(colour = "black"),
+      axis.title.x = element_text(margin = unit(c(2, 0, 0, 0), "mm")),
+      axis.title.y = element_text(margin = unit(c(0, 4, 0, 0), "mm")),
+      legend.position = "none"
+    )
+)
+
+# ----------------------------------------------------------------------------
+# Make Figure 2:
+
+# The raw data required to reproduce Figure 2 is produced in script 
+# '02-maxent-native-range.R'. Below, we import this raw data and then make 
+# the final figure.
+# ----------------------------------------------------------------------------
+
+# Load raw data 
+modelEvaluation <- readr::read_csv("./models/model_tuning/model_tuning_results.csv")
+head(modelEvaluation)
+
+# Plot AUCtest
+plot_aucTest <- ggplot(data = modelEvaluation, aes(x = rm,
+                                                   y = AUC.test,
+                                                   group = feat.class)) +
+  geom_point(aes(colour = feat.class)) +
+  geom_line(aes(colour = feat.class)) +
+  scale_colour_grey(start = 0.4, end = 0.8) +
[…]

### 2022-02-06 R/scripts/figure-3-make.R
Make Figure 3

@@ -1,0 +1,593 @@
+# -----------------------------------------------------------------------------
+# Script: - Reproduce Figure 3
+#
+# AUTHOR: Guy F. Sutton
+# AFFILIATION: Centre for Biological Control, Rhodes University, South Africa
+# DATE MODIFIED: 03/02/2021
+# CONTACT: g.sutton@ru.ac.za
+# -----------------------------------------------------------------------------
+
+# -----------------------------------------------------------------------------
+# Session setup
+# -----------------------------------------------------------------------------
+
+# Load required packages
+if (!require("pacman"))
+  install.packages("pacman")
+pacman::p_load(
+  tidyverse,
+  spocc,
+  ggmap,
+  maptools,
+  maps,
+  ggplot2,
+  scrubr,
+  mapr,
+  tidyr,
+  stringr,
+  rnaturalearth,
+  rnaturalearthdata,
+  rlang,
+  sf,
+  ggspatial,
+  raster,
+  here,
+  tidytext,
+  ggtext,
+  tidymodels,
+  themis,
+  ranger,
+  yardstick,
+  vip,
+  DHARMa,
+  rms
+)
+
+# Define function to calculate uncertainty in out-of-sample metrics
+boot_val <- function(data, formula, ...) {
+  out <- list()
+  for (i in 1:500) {
+    df <- sample_n(data, nrow(data), replace = TRUE)
+    md <- glm(formula, data = df, family = binomial)
+    out[[i]] <- val.prob(predict(md, type = "response"),
+                         as.numeric(df$established) - 1,
+                         pl = FALSE) %>% round(3)
+  }
+  return(out)
+}
+
+# ----------------------------------------------------------------------------
+# Plot for default settings model: 
+# ----------------------------------------------------------------------------
+
+# Import Maxent prediction raster 
+clim_map <- raster::raster("./models/rasters/raster_default.asc")
+
+# Import invaded-range GPS points 
+inv_data <- readr::read_csv("./data/data_clean/dasi_rubi_invaded_range_gps_clean.csv")
+head(inv_data)
+
+# Keep just the columns of interest
+inv_clean <- inv_data %>%
+  dplyr::select(lon = lon_converted,
+                lat = lat_converted,
+                established) %>%
+  mutate(lat = lat * -1)
+inv_clean
+
+# Keep only the sites with establishment data
+inv_clean <- inv_clean %>%
+  tidyr::drop_na(established)
+
+# Extract the maxent scores for each site
+clim_data <- raster::extract(clim_map,
[…]

### 2022-02-06 ms_body/manuscript.Rmd
Add LaTex manuscript document and template

@@
     latex_engine: pdflatex
-    template: ../svm-latex-ms.tex
+    template: "./_ms_template/svm-latex-ms.tex"
   html_document:
@@
 fontfamily: mathpazo
-bibliography: refs.bib
-csl: biological-control.csl
+bibliography: "./ref_list/refs.bib"
+csl: "./ref_list/biological-control.csl"
 fontsize: 12pt
@@
 ```{r import_fig_1, echo = FALSE}
-knitr::include_graphics(here::here("./ms_body/ms_figs/fig_1.png"), dpi = NA)
+# knitr::include_graphics(here::here("./ms_body/ms_figs/fig_1.png"), dpi = NA)
 ```
@@
 ```{r import_fig_2, echo = FALSE}
-knitr::include_graphics(here::here("./ms_body/ms_figs/fig_2.png"), dpi = NA)
+# knitr::include_graphics(here::here("./ms_body/ms_figs/fig_2.png"), dpi = NA)
 ```
@@
 ```{r import_fig_3, echo = FALSE}
-knitr::include_graphics(here::here("./ms_body/ms_figs/fig_3.png"), dpi = NA)
+# knitr::include_graphics(here::here("./ms_body/ms_figs/fig_3.png"), dpi = NA)
 ```
@@
 ```{r import_fig_4, echo = FALSE}
-knitr::include_graphics(here::here("./ms_body/ms_figs/fig_4.png"), dpi = NA)
+# knitr::include_graphics(here::here("./ms_body/ms_figs/fig_4.png"), dpi = NA)
 ```
@@
 ```{r import_fig_5, echo = FALSE}
-knitr::include_graphics(here::here("./ms_body/ms_figs/fig_5.png"), dpi = NA)
+# knitr::include_graphics(here::here("./ms_body/ms_figs/fig_5.png"), dpi = NA)
 ```
@@
 ```{r import_fig_6, echo = FALSE}
-knitr::include_graphics(here::here("./ms_body/ms_figs/fig_6.png"), dpi = NA)
+# knitr::include_graphics(here::here("./ms_body/ms_figs/fig_6.png"), dpi = NA)
 ```

### 2022-02-06 ms_body/manuscript.tex
Add LaTex manuscript document and template

\documentclass[12pt,]{article}
\usepackage[left=1in,top=1in,right=1in,bottom=1in]{geometry}
\newcommand*{\authorfont}{\fontfamily{phv}\selectfont}
\usepackage[]{mathpazo}

  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}

\usepackage{abstract}
\renewcommand{\abstractname}{}    % clear the title
\renewcommand{\absnamepos}{empty} % originally center

\renewenvironment{abstract}
 {{%
    \setlength{\leftmargin}{0mm}
    \setlength{\rightmargin}{\leftmargin}%
  }%
  \relax}
 {\endlist}

\makeatletter
\def\@maketitle{%
  \newpage
%  \null
%  \vskip 2em%
%  \begin{center}%
  \let \footnote \thanks
    {\fontsize{18}{20}\selectfont\raggedright  \setlength{\parindent}{0pt} \@title \par}%
}
%\fi
\makeatother

\setcounter{secnumdepth}{0}

\author{}

\date{}

\usepackage{titlesec}

\titleformat*{\section}{\normalsize\bfseries}
\titleformat*{\subsection}{\normalsize\itshape}
\titleformat*{\subsubsection}{\normalsize\itshape}
\titleformat*{\paragraph}{\normalsize\itshape}
\titleformat*{\subparagraph}{\normalsize\itshape}

\newtheorem{hypothesis}{Hypothesis}
\usepackage{setspace}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{hyperref}
\setlength\parindent{24pt}
\usepackage{multirow}
\usepackage{helvet}
\usepackage{setspace}
\renewcommand\familydefault{\sfdefault}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

% move the hyperref stuff down here, after header-includes, to allow for - \usepackage{hyperref}

\makeatletter
\@ifpackageloaded{hyperref}{}{%
\ifxetex
  \PassOptionsToPackage{hyphens}{url}\usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \PassOptionsToPackage{hyphens}{url}\usepackage[draft,unicode=true]{hyperref}
\fi
}

\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
             pdfkeywords = {},  
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

% Add an option for endnotes. -----

% add tightlist ----------
\providecommand{\tightlist}{%
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% add some other packages ----------

% \usepackage{multicol}
% This should regulate where figures float
% See: https://tex.stackexchange.com/questions/2275/keeping-tables-figures-close-to-where-they-are-mentioned
\usepackage[section]{placeins}

\begin{document}

% \pagenumbering{arabic}% resets `page` counter to 1 
%

\vskip -8.5pt

 % removetitleabstract

\noindent \doublespacing

\pagebreak

\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}
\setlength{\parskip}{8pt}
\vspace*{-0.2in}

\noindent

\textbf{Testing MaxEnt model performance in a novel geographic region
using an intentionally introduced insect}

Sutton, G. F.\textsuperscript{1}\(\dag\); Martin, G.D.
\textsuperscript{1,2}

\hfill\break

\begingroup
\fontsize{10}{12}\selectfont

\textsuperscript{1} Center for Biological Control, Department of Zoology
and Entomology, Rhodes University, Makhanda, 6140, South Africa

\textsuperscript{2} Afromontane Research Unit and Department Zoology and
Entomology, University of the Free State, Phuthaditjhaba, 9866, South
Africa \endgroup

\hfill\break

\textbf{Corresponding author} \(\dag\)
\href{mailto:g.sutton@ru.ac.za}{\nolinkurl{g.sutton@ru.ac.za}}

\pagebreak

\setlength\parindent{24pt}

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

Ecological models are important tools to support and guide the
development and implementation of environmental policies and management
programs. Species distributions models (SDM's), such as the popular
MaxEnt software, are frequently used to guide conservation programmes,
predict the potential distribution of invasive species, forecast the
impacts of climate change and develop applied ecological research
(e.g.~biological control). Many applications of these models require
that SDM's are transferred in either space and/or time. However, few
studies to date have tested the transferability, and thus usefulness, of
MaxEnt models using fully independent data. Moreover, numerous authors
have raised concerns over how model complexity (controlled primarily by
feature class and regularisation multiplier settings) may affect MaxEnt
model transferability. In this paper, we evaluated the usefulness of
MaxEnt models transferred in space using a native Australian insect,
\emph{Dasineura rubiformis} Kolesik (Diptera: Cecidomyiidae) that has
been intentionally introduced as a biological control agent of the
invasive plant, \emph{Acacia mearnsii} De Wild (Fabaceae) in South
Africa. MaxEnt models were developed using native-range records only
(Australia) and projected over South Africa to identify the potential
climatic suitability for the insect, using a range of model settings
configurations. Model transferability was assessed using independent
post-release data from South Africa. Our results demonstrated that
MaxEnt scores were positively correlated with increased establishment
rates for \emph{D. rubiformis} in South Africa. However, we found that
this pattern held only when MaxEnt models were specified with
user-defined settings configurations. A MaxEnt model calibrated with
default settings was overfit to the training data, and performed poorly
when transfered in space. Our study demonstrates that MaxEnt models can
prove useful when transferred in space, however, user-optimised model
settings likely need to be used to control model complexity, and reduce
the likelihood of overfitting.

~

\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}
\setlength{\parskip}{8pt}
\vspace*{-0.2in}

\noindent

\textbf{Keywords}: Species distribution model, biological control,
\emph{Acacia}, \emph{Dasineura rubiformis}, model complexity, model
transferability

\setlength\parindent{24pt}

\hypertarget{introduction}{%
\section{1. Introduction}\label{introduction}}

Ecological models are important tools to support and guide the
development and implementation of environmental policies and management
programs (\textbf{Addison2013?}; \textbf{Schuwirth2019?}). For example,
ecological models can be used for planning conservation programs
(\textbf{Guisan2013?}), predicting the establishment of invasive species
(\textbf{Martin2020?}), implementing biological control programmes
(\textbf{Mukherjee2021?}), or forecasting species responses to
environmental change (\textbf{Bocedi2014?}). Species distribution models
(SDM's) are an example of ecological models that have become
increasingly popular in recent years (\textbf{Elith2009?}). SDM's
typically take the form of correlative or mechanistic models that relate
species presence/absences to environmental covariates to map habitats
that may be suitable for a taxon (\textbf{Elith2011?}). The Maximum
Entropy Species Distribution Model (hereafter `MaxEnt')
(\textbf{Phillips2017?}) is among the most popular methods for species
distribution modeling and has been shown to perform well compared to
alternative modeling algorithms (\textbf{Elith2006?}). It uses maximum
entropy to distinguish between environmental conditions where the focal
taxon is present from environmental conditions at background locations
where the taxons' presence is absent (or assumed absent)
(\textbf{Elith2011?}).

Despite its popularity, the usefulness of MaxEnt has been questioned
when used to generate predictions for novel geographic regions
(transferred in space) or time periods (transferred in time) from those
for which the data used to build the model were collected
(\textbf{Yates2018?}; \textbf{Blasi2021?}). To date, studies assessing
transferability have produced contrasting results (\textbf{Randin2006?};
\textbf{Bahn2013?}; \textbf{Capinha2018?}). Model transferability has
typically been assessed by cross validation, whereby a fraction of the
data is withheld to calibrate the model (i.e.~training data) and the
remaining data is retained to validate the model (i.e.~the testing data)
(\textbf{Wenger2012?}). However, cross validation may not provide an
estimate of model transferability as a selected test sample may not
provide an unbiased estimate of the full data set, and therefore may not
reflect the heterogeneity that may exist in the underlying data in space
or time (see \textbf{Wenger2012?}). Ideally, a direct test of model
transferability should use an independent dataset of species
presence/absences and environmental covariates (\textbf{Fielding1997?};
\textbf{West2016?}). Unfortunately, very few studies, to date, have
tested model transferability using independent datasets (e.g.
\textbf{Costa2010?}; \textbf{Rebelo2010?}), which may have limited our
understanding of how useful SDM's can be when transferred in space or
time. Issues with model transferability may lead to misleading
conclusions, which may hinder scientists abilities to address important
ecological questions (e.g.~where could an invasive species establish?,
where should a conservation priority be translocated?)
(\textbf{Merow2013?}).

Model complexity can affect uncertainty of model transferability in
space and time (\textbf{Warren2014?}). Model complexity may be affected
by numerous parameters, notably the feature classes and regularisation
(beta) multiplier settings used to specify the MaxEnt model
(\textbf{Elith2011?}; \textbf{Merow2013?}). MaxEnt allows users to
specify combinations of different feature classes that modify the shape
and complexity of species-environment functions. Moreover, MaxEnt uses a
regularisation multiplier (a form of L\textsubscript{1}-regularisation)
to balance model complexity with the model fit, penalising unnecessarily
complex models (\textbf{Merow2013?}). The default MaxEnt settings use
all feature classes in model fitting except `threshold' features
(depending on sample size of training data) and a regularisation
multiplier of 1 (\textbf{Phillips2017?}). Several recent studies have
highlighted the importance of species-specific tuning of feature classes
and regularisation multipliers for optimising model complexity
(\textbf{Shcheglovitova2013?}; \textbf{Radosavljevic2014?}), with
default settings often producing models that are overly complex, and
thus, may perform poorly when transferred (\textbf{Elith2011?};
\textbf{Merow2013?}). Despite the potential implications, surprisingly
few studies use user-defined feature class and regularisation settings
(e.g. \textbf{Sutton2019b?}), and as such, little is known about how
model settings configurations may influence model transferability
(\textbf{Moreno-Amat2015?}; \textbf{Low2021?}).

In this study, we performed a retrospective assessment of the
transferability of SDM's built using MaxEnt (\textbf{Phillips2017?}).
Here, we modelled the potential climatic suitability of an insect that
is native to Australia, and that has been introduced into South Africa
as a biological control agent. The insect, \emph{Dasineura rubiformis}
Kolesik (Diptera: Cecidomyiidae) was introduced into South Africa from
Australia in 2013, to curb the spread of the invasive Australian
\emph{Acacia mearnsii} De Wild (\textbf{Impson2013?}). Extensive field
surveys have been performed in Australia (characterizing the native
distribution of \emph{D. rubiformis}) (\textbf{Adair2004?}) and
long-term post-release monitoring has been performed in South Africa to
measure establishment rates of the insect, following experimental
releases of the insect (Impson et al., 2021). As such, the availability
of independent testing data from a separate continent to the data used
to generate the models makes this an ideal study system for assessing
the transferability of SDM's.

We built MaxEnt models for \emph{D. rubiformis} using distribution
records from its native distribution (Australia), and projected the
resulting MaxEnt suitability rasters over South Africa. MaxEnt scores
range from 0 to 100, whereby 0 represents regions predicted to be highly
unsuitable and 100 indicating highly suitable regions. To test model
transferability, we fit the MaxEnt scores as a covariate in a logistic
regression model to determine whether MaxEnt scores were statistically
relevant predictors of insect establishment rates in South Africa (with
these data being entirely independent of the data used to build the
MaxEnt models). A strong, positive correlation between MaxEnt scores and
insect establishment rates would provide support that the SDM's were
useful when transferred in space (from Australia to South Africa).
Moreover, we tested whether model complexity and parameter tuning could
improve the accuracy and transferability of MaxEnt by comparing the
predictive power of MaxEnt models calibrated with default settings and
an array of optimised settings configurations.

\hypertarget{methods-and-materials}{%
\section{2. Methods and Materials}\label{methods-and-materials}}

\hypertarget{species-occurences-records}{%
\subsection{2.1. Species occurences
records}\label{species-occurences-records}}

A total of 103 native-range records for \emph{D. rubiformis} were
obtained from (\textbf{Adair2004?}). A total of 452 invaded-range
records to evaluate model transferability were taken from Impson,
Kleinjan and Hoffmann (unpublished). Distribution maps for \emph{D.
rubiformis} for both native and invaded range are provided in Fig. 1.

Spatial autocorrelation is an important factor that may affect model
outputs. Filtering of species occurrence data may limit the inherent
biases in the data and improve model quality (\textbf{Veloz2009?}). To
avoid pseudo-replication, only one occurrence record per 2.5 minute grid
cell was used for model calibration. Species occurrence datasets were
thinned using the `spThin' package (\textbf{Aiello-Lammens2015?}), and
spatial autocorrelation analyses were performed using the `ecospat'
package (\textbf{DiCola2017?}).

\hypertarget{environmental-predictors}{%
\subsection{2.2. Environmental
predictors}\label{environmental-predictors}}

Climate data were obtained by downloading the standard set of 19
bioclimatic variables from the WorldClim ver. 1.4 database
(\textbf{Hijmans2005?}) (data available at:
www.worldclim.org/download.html). This dataset is representative of
annual and seasonal means and variation of temperature and precipitation
metrics averaged over the 1950--2000 time period (current climate) at a
5 arc minute resolution.

To reduce multicollinearity between environmental predictors and thereby
decrease the risk of calibrating overfit models, Pearson's correlation
coefficients were computed for all pairs of predictors, whereby
predictors which were highly correlated (\textbar{}\emph{r}\textbar{}
\textgreater{} 0.85) were excluded from the final predictor set
(\textbf{Capinha2011?}). The reduced set of environmental predictors
consisted of seven climatic variables, including: bio1 -- annual mean
temperature, bio2 -- mean diurnal temperature range (mean of monthly
{[}max temp -- min temp{]}, bio3 -- isothermality, bio9 -- mean
temperature of the driest quarter, bio12 -- mean annual precipitation,
bio14 -- precipitation of the driest month and bio16 -- precipitation of
the wettest quarter (see (\textbf{Hijmans2005?}) for further details).

\hypertarget{model-calibration}{%
\subsection{2.3. Model calibration}\label{model-calibration}}

Maxent was implemented in the \texttt{dismo} package in \texttt{R}
(\textbf{hijmans2017package?}). MaxEnt was selected as it consistently
outperforms other modelling algorithms (\textbf{Wisz2008?}). Given that
MaxEnt is a presence-only modelling algorithm, model calibration
requires a user-defined geographic background to sample the climate of
representative grid cells where the focal species is absent
(i.e.~background points). Background definition can have a significant
effect on model output (\textbf{VanDerWal2009?}). The background should
ideally represent the geographic areas available to the focal species,
omitting areas where species absence is due to historical factors,
dispersal constraints and/or biotic interactions (\textbf{Sanin2018?}).
Following (\textbf{Webber2011?}), we defined the model background using
the Koppen-Geiger climate classification (Available at:
\url{http://koeppen-geiger.vu-wien.ac.at}). Only Koppen-Geiger climate
zones that contained at least one native-range occurrence record for
\emph{D. rubiformis} were used as the background area from which
pseudo-absences were drawn for model calibration. Koppen-Geiger climate
zones were intersected using the \texttt{extract} function from the
\texttt{R} package \texttt{raster} (\textbf{hijmans2017package?}).

\hypertarget{optimal-model-settings}{%
\subsection{2.4. Optimal model settings}\label{optimal-model-settings}}

Model performance and optimal settings configurations were assessed
using multiple metrics that reflected aspects of model (1)
discriminatory ability (AUC\textsubscript{test}), (2) overfitting
(AUC\textsubscript{diff}), (3) omission rates (OR\textsubscript{10}),
and (4) overall parsimony (AICc) (\textbf{Low2021?}).

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\item
  AUC\textsubscript{test} assesses the models ability to discriminate
  between predicted presence/absence at withheld portions of the data
  used to test the model versus background points. The area under the
  receiver operating characteristic curve (AUC) is one of the most
  popular metrics used to evaluate MaxEnt models. An AUC of less than
  0.8 is considered a poor model, between 0.8 and 0.9 is a fair model,
  between 0.9 and 0.995 a good model, and \textgreater{} 0.995 an
  excellent model (\textbf{Fielding1997?}). Thus, higher
  AUC\textsubscript{test} values indicate increased ability to
  discriminate between training and background points.
\item
  AUC\textsubscript{diff} calculates the difference between AUC values
  calculated on training points only (AUC\textsubscript{train}) and
  AUC\textsubscript{test} using cross-validation (\textbf{Warren2011?}).
  Thus, higher AUC\textsubscript{diff} values indicate whether the
  MaxEnt model is overfit on the training data, and thus, performs
  poorly when evaluated against testing points.
\item
  OR\textsubscript{10} is the proportion of testing points that are not
  predicted to fall within the projected model surface once the model is
  converted into a binary prediction output (\textbf{Boria2014?}).
  Overfit models have omission rates higher than the theoretical
  expectation for the threshold applied (\textbf{Shcheglovitova2013?}).
  OR10 sets the binary prediction threshold at a value that excludes the
  10\% of the calibration localities from the model with the lowest
  prediction values, and therefore has an expected omission rate of 0.10
  (\textbf{Boria2014?}). As such, the OR\textsubscript{10} criterion
  selected models calibrated with MaxEnt settings which best
  approximated the expected 0.10 omission rate. Models with omission
  rates increasingly higher than the expected value were considered as
  more overfit (\textbf{Boria2017?}).
\item
  Lastly, optimal model settings were determined by selecting model
  configurations which produced the lowest value for the Akaike
  Information Criterion corrected for small sample sizes (AICc)
  (i.e.~AICc=0; following (\textbf{Muscarella2014?})). The AICc
  criterion simultaneously scores models according to their complexity
  and goodness-of-fit, whereby models with the lowest AICc are selected
  as the best models. AICc was used as the primary evaluation metric as
  it is calculated using MaxEnt models built using the entire species
  occurrence dataset, unlike AUC and OR\textsubscript{10} (and numerous
  other metrics frequently used for model evaluation) which may be
  spatially biased due to the partitioning of the species occurrence
  dataset into training and evaluation sets (\textbf{Sanin2018?}).
\end{enumerate}

\hypertarget{assessing-model-extrapolation}{%
\subsection{2.5. Assessing model
extrapolation}\label{assessing-model-extrapolation}}

Several authors have raised concerns over the use of correlative SDM's,
such as MaxEnt, when projected into new geographic regions or time
periods due to issues with extrapolation into novel/non-analogous
climate (\textbf{Elith2009?}; \textbf{Elith2011?}; \textbf{Yates2018?}).
To address these concerns, Multivariate environmental similarity
surfaces (hereafter `MESS') (\textbf{Elith2010?}) were computed to
assess whether MaxEnt models were extrapolating or interpolating. MESS
analyses measure the similarity of any given point to a set of reference
points (reference points were defined as only the species presence
points used to calibrate models following \textbf{Kriticos2014?}).
Negative MESS values indicate geographic regions outside the range of
climate variables used to calibrate the model (i.e.~extrapolation space
or MESS-), while MESS values between 0 and 100 indicate geographic
regions inside the range of climatic variables used to calibrate the
model (i.e.~interpolation space or MESS+). MESS maps are used as a
measure of prediction uncertainty or caution against inferences in
extrapolation space (\textbf{Elith2010?}).

\hypertarget{statistical-analyses}{%
\subsection{2.6. Statistical analyses}\label{statistical-analyses}}

Five candidate MaxEnt models were specified using native-range
(Australia) training data. The five models differed in the feature
classes and regularisation multiplier settings used by MaxEnt to build
the models. The five models were specified with (1) default MaxEnt
settings, and four optimised settings configurations that maximised
performance based on (2) AUC\textsubscript{test}, (3)
AUC\textsubscript{diff}, (4) OR\textsubscript{10} and (5) AICc.

To evaluate the predictive performance and transferability of these
models, we evaluated whether MaxEnt climatic suitability scores were
correlated with increased probabilities of \emph{D. rubiformis}
establishment in South Africa. To do so, MaxEnt scores were specified as
a continuous fixed effect and whether \emph{D. rubiformis} was recorded
or not (present/absent) as a boolean response variable in a logistic
general linear model (GLM), using a logit link function. MaxEnt scores,
bounded between 0 (not climatically suitable) and 1 (highly climatically
suitable) indicating climatic similarity between the native range
occupied by \emph{D. rubiformis} in Australia and sites where it was
released as a biological control agent in South Africa, were derived by
projecting MaxEnt suitability rasters from models calibrated using
native-range GPS records only (Australian native-range records) over
South Africa, and extracting the MaxEnt suitability scores for each site
where \emph{D. rubiformis} was surveyed in South Africa using the
\texttt{extract} function from the \texttt{R} package \texttt{raster}
(\textbf{hijmans2017package?}). GLM's were specified using the
\texttt{rms} package in \texttt{R} (\textbf{Harrell2017package?}).
Likelihood Ratio Tests (LRT) were performed to determine whether there
was a statistical correlation between MaxEnt scores and the probability
of \emph{D. rubiformis} establishment. A statistically significant,
positive correlation between MaxEnt scores and insect establishment
rates would provide support that the SDM's were useful and predictive of
insect establishment, even when the insect was transferred in space.
Out-of-sample AUC was calculated using the \texttt{rms} package, using
500 bootstrap replicates (\textbf{Harrell2017package?}) to quantify the
predictive performance of MaxEnt models, whereby AUC values less than
0.80 are typically considered poor fitting models (lacks predictive
power) (\textbf{Fielding1997?}).

A similar approach was adopted to test if each occurrence record in
South Africa was classified as being in interpolation (within the
climate space occupied by the insect in its native range) versus
extrapolation (outside the climate space occupied by the insect in its
native range) was predictive of \emph{D. rubiformis} establishment.
Interpolation/extrapolation were quantified using MESS maps using the
\texttt{mess} function from the \texttt{dismo} package in \texttt{R}
(\textbf{hijmans2017package?}). Interpolation versus extrapolation was
fit as a categorical predictor and whether \emph{D. rubiformis} was
recorded or not (present/absent) as a boolean response variable in a
logistic GLM, using a logit link function.

All modelling and statistical analyses were conducted in \texttt{R} ver.
4.0.3 (\textbf{RCoreTeam2020?}).

\hypertarget{results}{%
\section{3. Results}\label{results}}

\hypertarget{native-range-models-model-calibration}{%
\subsection{3.1. Native-range models (model
calibration)}\label{native-range-models-model-calibration}}

Five candidate MaxEnt models were developed using native-range
(Australia) training data for \emph{D. rubiformis}. The five models
differed in the feature classes and regularisation multiplier settings
used by MaxEnt to build the models. The five models were specified with
(1) default MaxEnt settings, and four optimised settings configurations
that (2) maximised performance based on AUC\textsubscript{test}, (3)
AUC\textsubscript{diff}, (4) OR\textsubscript{10} and (5) AICc (Table
S1). The AUC values for these five models ranged were ranged between
0.82 and 0.97, depending on the MaxEnt settings configuration used. As
such, all five candidate models were considered useful and had high
predictive accuracy in distinguishing between \emph{D. rubiformis}
recorded presences and background points in its native range. While
default MaxEnt settings produced the highest discriminatory power when
tested in the training area of Australia (AUC = 0.97), models calibrated
with tuned settings configurations led to less overfitting (lower
AUC\textsubscript{diff}), better omission rates (best approximated the
expectation for OR\textsubscript{10}), and more parsimonious models
(AICc) (Fig. 2).

\hypertarget{model-transferability}{%
\subsection{3.2. Model transferability}\label{model-transferability}}

When evaluated on independent testing data in South Africa, the MaxEnt
models showed relatively high, but variable, predictive accuracy,
depending on the MaxEnt settings used to calibrate the model. For all
models, the probability that \emph{D. rubiformis} was recorded at a
field site in South Africa increased with higher MaxEnt climatic
suitability scores (Fig. 3). The default MaxEnt model found no
statistically significant relationship between MaxEnt climatic
suitability scores and the probability of \emph{D. rubiformis}
establishment in South Africa (Table. 1), and displayed low power to
discriminate between sites where \emph{D. rubiformis} had established or
not (AUC: 0.64 \(\pm\) 0.01) (Fig. 4). In contrast, the four models
calibrated using optimised model settings configurations all
demonstrated a statistically significant positive relationship between
MaxEnt suitability scores and the probability of \emph{D. rubiformis}
establishment in South Africa (Table. 1). All four tuned models
displayed high power to discriminate between sites where \emph{D.
rubiformis} had established or not (AUC: 0.83 - 0.84) than the MaxEnt
model calibrated with default settings (AUC: 0.64 \(\pm\) 0.01) (Fig.
4). The predicted climatic suitability for \emph{D. rubiformis} across
South Africa varied depending on the MaxEnt settings used to calibrate
the model (Fig. 5). Default MaxEnt settings yielded higher climatic
suitability estimates for \emph{D. rubiformis} than the other MaxEnt
models with tuned model settings, notably in regions where \emph{D.
rubiformis} is absent (e.g.~the north-west coastal region bordering
Namibia).

Despite the high performance of the MaxEnt models when transferred in
space, MESS maps indicated that the MaxEnt models were extrapolating
climatic suitability scores for \emph{D. rubiformis} over the vast
majority of South Africa (Fig. S1). Indeed, 54 of the 78 field sites
surveyed for \emph{D. rubiformis} in South Africa occur in geographic
regions where the MaxEnt models were extrapolating. \emph{Dasineura
rubiformis} established at 23 of 24 sites in interpolation (96\%), while
it established at 44 of 54 sites in extrapolation (82\%) (Fig. 6). We
found no evidence that a site being in extrapolation/interpolation was a
statistically significant predictor of \emph{D. rubiformis}
establishment in South Africa (X\textsubscript{2} = 3.40, d.f = 1,
\emph{P} = 0.07). Moreover, classifying sites as being in
extrapolation/interpolation yielded lower power to discriminate to
discriminate between sites where \emph{D. rubiformis} had established or
not (AUC: 0.67 \(\pm\) 0.02) than all of the MaxEnt models, barring the
model with default settings.

\hypertarget{discussion}{%
\section{4. Discussion}\label{discussion}}

Species distribution models (SDM's) are becoming increasingly important
tools to support and guide the development and implementation of
environmental policies and management programs (\textbf{Addison2013?};
\textbf{Schuwirth2019?}). The usefulness of SDM's is dependent on model
performance, which is typically measured by its ability to correctly
predict where a species is established and its capacity to distinguish
between sites where the species is established versus a site where the
species is absent (or assumed to be absent) (\textbf{Smith2021?}).
However, a number of authors have called in question the usefulness of
SDM's, notably when models are transferred in space or time
(\textbf{Yates2018?}; \textbf{Blasi2021?}). Poor transferability may
lead to erroneous model predictions that may have serious negative
consequences for the application of the model, and resulting policies
and management programmes (Soultan and Safi, 2017). Here, we report on
the validation of a SDM for an insect, (\emph{D. rubiformis}), native to
Australia, that has been intentionally introduced as a biological
control agent into South Africa to reduce the invasiveness of its
host-plant, \emph{A. mearnsii} (Impson et al., 2013). The the
availability of independent testing data from South Africa, a separate
continent to the data used to generate the models (Australia), makes
this an ideal study system for assessing the transferability of SDM's.
This study represents one of only a handful of studies to validate SDM's
using independent testing data (\textbf{Costa2010?};
\textbf{Rebelo2010?}; \textbf{West2016?}; \textbf{Smith2021?}).

Our results indicated that MaxEnt SDM's were able to provide robust
estimates of climatic suitability for \emph{D. rubiformis} in South
Africa. AUC values greater than 0.80 are typically considered good
fitting models (having high predictive power) (\textbf{Fielding1997?}).
When MaxEnt models, calibrated using native-range records for \emph{D.
rubiformis} from Australia, were projected over South Africa, they had
AUC scores of 0.83 - 0.84, indicative of high power to discriminate
between sites where \emph{D. rubiformis} had established in South Africa
or not. Additionally, we showed that the probability that \emph{D.
rubiformis} was recorded at a field site in South Africa increased with
higher MaxEnt climatic suitability scores. This result demonstrates the
usefulness of the MaxEnt SDM and that, at least for our case-study,
SDM's can be transferred into novel geographic regions, and provide
valuable information to guide policy making and the development of
management programmes. For example, biological control programmes such
as the programme in which \emph{D. rubiformis} is used to curtail the
spread of the highly invasive weed \emph{A. mearnsii} expend many
resources (time and money) to mass-rear insect populations that are then
released into the field. The positive correlation between MaxEnt
suitability scores and \emph{D. rubiformis} establishment means that
researchers can develop and implement mass-rearing and experimental
release programmes for the insect that prioritise releasing the insect
at sites that are highly climatically suitable, and thus, more likely
the insect will establish viable populations. This may assist in
improving the level of control over the target weed and increasing the
efficiency of resources expended by increasing the chances of
establishing viable insect populations.

Our study adds to a growing body of literature indicating that SDM model
complexity affects model performance and transferability when being
projected into novel geographic regions. While the default MaxEnt model
had the highest predictive power in the training area (Australia), this
model performed extremely poorly when transferred over South Africa. In
contrast, the four MaxEnt models calibrated with tuned settings showed
high predictive power when transferred. The poor transferability of the
Default model is likely explained by default settings overfitting the
model on the training points (\textbf{Warren2014?}). This finding
corroborates other studies showing that high predictive power in the
model training area is not a guarantee of transferability
(\textbf{Warren2014?}; \textbf{Duque-Lazo2016?}). We echo that
recommendations of (\textbf{Shcheglovitova2013?}) and
(\textbf{Muscarella2014?}) that model tuning should be performed to
estimate optimal model complexity when transferring MaxEnt models in
space and/or time in order to maximise their usefulness.

An additional concern with transferring MaxEnt models in space or time
is the potential for models to significantly extrapolate predictions
when the model is projected into non-analogous conditions to those under
which the model was calibrated (Elith et al., 2010; Mesgaran et al.,
2014). However, the applied use of SDM's in developing environmental
policy, developing management programmes for invasive species and
implementing conservation programmes often necessitate that models are
transferred and predictions made, even when models are extrapolating
(Elith et al., 2010). Here, we showed that \emph{D. rubiformis} was able
to establish at many sites (82\%) in South Africa that were outside the
range of climate space it occupies in its native range
(i.e.~extrapolation), which was not substantially different from
establishment rates at sites in interpolation. While our MaxEnt models
were able to make accurate predictions of \emph{D. rubiformis}
establishment in South Africa, even when models were extrapolating, this
finding in no way discounts the importance of assessing model
extrapolation when transferring models in space or time. A number of
tools have been developed in recent years to assist SDM users with
identifying model extrapolation, e.g.~MESS (\textbf{Elith2010?}) and
ExDet (Mesgaran et al., 2014), and which should be an integral component
of any SDM study.

Given that the \emph{D. rubiformis} MaxEnt models showed high predictive
accuracy when transferred in space over South Africa and the
statistically significant positive correlation between MaxEnt scores and
\emph{D. rubiformis} establishment, we wish to highlight how these
models can be used in an applied setting. To visualise the potential
geographic regions where \emph{D. rubiformis} may establish in South
Africa, we thresholded the MaxEnt map that optimised AICc where the
probability of establishment was \textgreater{} 50\% (i.e.~MaxEnt score
= 0.30), given in Fig. 9. The map demonstrates that the entire south and
east coast regions of South Africa are climatically suitable for
\emph{D. rubiformis}. Mass-releasing and releasing weed biological
control agents is a time-consuming and costly endeavor (Coetzee et al.,
2021). As such, our models provide a statistically validated approach to
optimising the redistribution of \emph{D. rubiformis} in South Africa by
prioritising releases at field sites where MaxEnt scores are
\textgreater{} 0.30, and therefore where the probability of
establishment was \textgreater{} 50\%.

\hypertarget{conclusions}{%
\section{5. Conclusions}\label{conclusions}}

Species distribution models are becoming increasingly popular with
researchers, stakeholders, policy makers and land managers alike,
particularly for making predictions in novel geographic geographic
regions or forecasting species responses to climate change. Transfering
SDM's in space and/or time, however, has been met with justifiable
resilience and concerns over how well these models perform upon
transfer. In this study, we presented a case-study demonstrating that
MaxEnt SDM's were able to accurately predict the distribution of an
Australian insect when introduced into South Africa. Our study is one of
only a few studies to date validating MaxEnt models using spatially
independent data. However, the accuracy of the MaxEnt models was
dependent on the settings used to calibrate the models, with default
MaxEnt programme settings having low predictive power. Our results
demonstrate that MaxEnt can be useful in predicting where a species may
establish, even when projected into a novel geographic area, but that
users need to be aware of how model settings can influence model outputs
and usefulness. Demonstrating how SDM's can be useful, while
communicating uncertainty in their predictions and outputs, is essential
for building confidence in the use of SDM's (Blasi et al.~2020). We
showed how validating predictive models can be used to guide the
implementation of a biological control programme by optimising the
redistribution and release of \emph{D. rubiformis} in geographic areas
that are climatically suitable and where the probability of the insect
establishing is greater than 50\%.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

GFS and GDM acknowledge funding from the South African Working for Water
(WfW) programme of the Department of Forestry, Fisheries and the
Environment: Natural Resource Management Programmes (DFFE: NRMP).
Funding was also provided by the South African Research Chairs
Initiative of the Department of Science and Technology and the National
Research Foundation (NRF) of South Africa. Any opinions, finding,
conclusions or recommendations expressed in this material are those of
the authors and the NRF does not accept any liability in this regard.

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\setlength{\parindent}{-0.0in}
\setlength{\leftskip}{0.0in}
\setlength{\parskip}{8pt}
\vspace*{-0.0in}

\noindent

\newlength{\cslhangindent}
\newenvironment{CSLReferences}%
{\setlength{\parindent}{0pt}%
\everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
{\par}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\end{CSLReferences}

\newpage

\hypertarget{figure-and-table-legend}{%
\section{Figure and table legend}\label{figure-and-table-legend}}

\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\vspace*{-0.2in}

\noindent

\newpage

\hypertarget{figure-1}{%
\subsection{\texorpdfstring{\emph{Figure 1}}{Figure 1}}\label{figure-1}}

\newpage

\hypertarget{figure-2}{%
\subsection{\texorpdfstring{\emph{Figure 2}}{Figure 2}}\label{figure-2}}

\newpage

\hypertarget{figure-3}{%
\subsection{\texorpdfstring{\emph{Figure 3}}{Figure 3}}\label{figure-3}}

\newpage

\hypertarget{figure-4}{%
\subsection{\texorpdfstring{\emph{Figure 4}}{Figure 4}}\label{figure-4}}

\newpage

\hypertarget{figure-5}{%
\subsection{\texorpdfstring{\emph{Figure 5}}{Figure 5}}\label{figure-5}}

\newpage

\hypertarget{figure-6}{%
\subsection{\texorpdfstring{\emph{Figure 6}}{Figure 6}}\label{figure-6}}

\newpage

\textbf{Table 1}

~

\begin{ThreePartTable}
\begin{TableNotes}
\item[a] Adjusted to indicate change in Y for each 1\% unit increase in covariate
\end{TableNotes}
\begin{longtable}[l]{lllll}
\toprule
Model & $\beta$\textsuperscript{a} & $\chi$\textsubscript{2} & d.f. & \textit{P}\\
\midrule
\endfirsthead
\multicolumn{5}{@{}l}{\textit{(continued)}}\\
\toprule
Model & $\beta$\textsuperscript{a} & $\chi$\textsubscript{2} & d.f. & \textit{P}\\
\midrule
\endhead

\endfoot
\bottomrule
\insertTableNotes
\endlastfoot
\cellcolor{gray!6}{Default settings} & \cellcolor{gray!6}{1.02} & \cellcolor{gray!6}{1.77} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0.18}\\
AUCtest & 1.07 & 13.00 & 1 & < 0.001\\
\cellcolor{gray!6}{AUCdiff} & \cellcolor{gray!6}{1.14} & \cellcolor{gray!6}{14.95} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{< 0.001}\\
OR10 & 1.10 & 14.38 & 1 & < 0.001\\
\cellcolor{gray!6}{AICc} & \cellcolor{gray!6}{1.09} & \cellcolor{gray!6}{14.38} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{< 0.001}\\*
\end{longtable}
\end{ThreePartTable}

\newpage

\textbf{Table S1}

~

\begin{longtable}[l]{llllll}
\toprule
Model & RM & FC & AUC\textsubscript{test} & AUC\textsubscript{diff} & OR\textsubscript{10}\\
\midrule
\endfirsthead
\multicolumn{6}{@{}l}{\textit{(continued)}}\\
\toprule
Model & RM & FC & AUC\textsubscript{test} & AUC\textsubscript{diff} & OR\textsubscript{10}\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
\cellcolor{gray!6}{Default} & \cellcolor{gray!6}{1.0} & \cellcolor{gray!6}{HLQPT} & \cellcolor{gray!6}{0.94} & \cellcolor{gray!6}{0.03} & \cellcolor{gray!6}{0.15}\\
AUCtest & 1.5 & HLQPT & 0.95 & 0.02 & 0.15\\
\cellcolor{gray!6}{AUCdiff} & \cellcolor{gray!6}{4.0} & \cellcolor{gray!6}{HLQPT} & \cellcolor{gray!6}{0.94} & \cellcolor{gray!6}{0.01} & \cellcolor{gray!6}{0.14}\\
OR10 & 6.0 & HLQ & 0.92 & 0.02 & 0.10\\
\cellcolor{gray!6}{AICc} & \cellcolor{gray!6}{4.0} & \cellcolor{gray!6}{H} & \cellcolor{gray!6}{0.94} & \cellcolor{gray!6}{0.02} & \cellcolor{gray!6}{0.12}\\*
\end{longtable}

\newpage
\singlespacing 
\end{document}

### 2022-02-06 ms_body/manuscript.tex
Add BibTex references to MS

@@
 
-Sutton, G. F.\textsuperscript{1}\(\dag\); Martin, G.D.
-\textsuperscript{1,2}
+Sutton, G. F.\textsuperscript{1}\(\dag\); Martin,
+G.D.\textsuperscript{1,2}
 
@@
 development and implementation of environmental policies and management
-programs (\textbf{Addison2013?}; \textbf{Schuwirth2019?}). For example,
-ecological models can be used for planning conservation programs
-(\textbf{Guisan2013?}), predicting the establishment of invasive species
-(\textbf{Martin2020?}), implementing biological control programmes
-(\textbf{Mukherjee2021?}), or forecasting species responses to
-environmental change (\textbf{Bocedi2014?}). Species distribution models
-(SDM's) are an example of ecological models that have become
-increasingly popular in recent years (\textbf{Elith2009?}). SDM's
-typically take the form of correlative or mechanistic models that relate
-species presence/absences to environmental covariates to map habitats
-that may be suitable for a taxon (\textbf{Elith2011?}). The Maximum
-Entropy Species Distribution Model (hereafter `MaxEnt')
-(\textbf{Phillips2017?}) is among the most popular methods for species
-distribution modeling and has been shown to perform well compared to
-alternative modeling algorithms (\textbf{Elith2006?}). It uses maximum
-entropy to distinguish between environmental conditions where the focal
-taxon is present from environmental conditions at background locations
-where the taxons' presence is absent (or assumed absent)
-(\textbf{Elith2011?}).
+programs (Addison et al., 2013; Schuwirth et al., 2019). For example,
+ecological models can be used for planning conservation programs (Guisan
+et al., 2013), predicting the establishment of invasive species (Martin
+et al., 2020), implementing biological control programmes (Mukherjee et
+al., 2021), or forecasting species responses to environmental change
+(Bocedi et al., 2014). Species distribution models (SDM's) are an
+example of ecological models that have become increasingly popular in
+recent years (Elith and Leathwick, 2009). SDM's typically take the form
+of correlative or mechanistic models that relate species
+presence/absences to environmental covariates to map habitats that may
+be suitable for a taxon (Elith et al., 2011). The Maximum Entropy
+Species Distribution Model (hereafter `MaxEnt') (Phillips et al., 2017)
+is among the most popular methods for species distribution modeling and
+has been shown to perform well compared to alternative modeling
+algorithms (Elith et al., 2006). It uses maximum entropy to distinguish
+between environmental conditions where the focal taxon is present from
+environmental conditions at background locations where the taxons'
+presence is absent (or assumed absent) (Elith et al., 2011).
 
@@
 (transferred in space) or time periods (transferred in time) from those
-for which the data used to build the model were collected
-(\textbf{Yates2018?}; \textbf{Blasi2021?}). To date, studies assessing
-transferability have produced contrasting results (\textbf{Randin2006?};
-\textbf{Bahn2013?}; \textbf{Capinha2018?}). Model transferability has
-typically been assessed by cross validation, whereby a fraction of the
-data is withheld to calibrate the model (i.e.~training data) and the
-remaining data is retained to validate the model (i.e.~the testing data)
-(\textbf{Wenger2012?}). However, cross validation may not provide an
-estimate of model transferability as a selected test sample may not
-provide an unbiased estimate of the full data set, and therefore may not
-reflect the heterogeneity that may exist in the underlying data in space
-or time (see \textbf{Wenger2012?}). Ideally, a direct test of model
+for which the data used to build the model were collected (Blasi et al.,
+2021; Yates et al., 2018). To date, studies assessing transferability
+have produced contrasting results (Bahn and McGill, 2013; Capinha et
+al., 2018; Randin et al., 2006). Model transferability has typically
+been assessed by cross validation, whereby a fraction of the data is
+withheld to calibrate the model (i.e.~training data) and the remaining
+data is retained to validate the model (i.e.~the testing data) (Wenger
+and Olden, 2012). However, cross validation may not provide an estimate
+of model transferability as a selected test sample may not provide an
+unbiased estimate of the full data set, and therefore may not reflect
+the heterogeneity that may exist in the underlying data in space or time
+(see Wenger and Olden, 2012). Ideally, a direct test of model
 transferability should use an independent dataset of species
-presence/absences and environmental covariates (\textbf{Fielding1997?};
-\textbf{West2016?}). Unfortunately, very few studies, to date, have
-tested model transferability using independent datasets (e.g.
-\textbf{Costa2010?}; \textbf{Rebelo2010?}), which may have limited our
+presence/absences and environmental covariates (Fielding and Bell, 1997;
+West et al., 2016). Unfortunately, very few studies, to date, have
+tested model transferability using independent datasets (e.g. Costa et
+al., 2010; Rebelo and Jones, 2010), which may have limited our
 understanding of how useful SDM's can be when transferred in space or
@@
 ecological questions (e.g.~where could an invasive species establish?,
-where should a conservation priority be translocated?)
-(\textbf{Merow2013?}).
+where should a conservation priority be translocated?) (Merow et al.,
+2013).
 
 Model complexity can affect uncertainty of model transferability in
-space and time (\textbf{Warren2014?}). Model complexity may be affected
+space and time (Warren et al., 2014). Model complexity may be affected
 by numerous parameters, notably the feature classes and regularisation
-(beta) multiplier settings used to specify the MaxEnt model
-(\textbf{Elith2011?}; \textbf{Merow2013?}). MaxEnt allows users to
-specify combinations of different feature classes that modify the shape
-and complexity of species-environment functions. Moreover, MaxEnt uses a
+(beta) multiplier settings used to specify the MaxEnt model (Elith et
+al., 2011; Merow et al., 2013). MaxEnt allows users to specify
+combinations of different feature classes that modify the shape and
+complexity of species-environment functions. Moreover, MaxEnt uses a
 regularisation multiplier (a form of L\textsubscript{1}-regularisation)
 to balance model complexity with the model fit, penalising unnecessarily
-complex models (\textbf{Merow2013?}). The default MaxEnt settings use
-all feature classes in model fitting except `threshold' features
-(depending on sample size of training data) and a regularisation
-multiplier of 1 (\textbf{Phillips2017?}). Several recent studies have
-highlighted the importance of species-specific tuning of feature classes
-and regularisation multipliers for optimising model complexity
-(\textbf{Shcheglovitova2013?}; \textbf{Radosavljevic2014?}), with
-default settings often producing models that are overly complex, and
-thus, may perform poorly when transferred (\textbf{Elith2011?};
-\textbf{Merow2013?}). Despite the potential implications, surprisingly
-few studies use user-defined feature class and regularisation settings
-(e.g. \textbf{Sutton2019b?}), and as such, little is known about how
-model settings configurations may influence model transferability
-(\textbf{Moreno-Amat2015?}; \textbf{Low2021?}).
+complex models (Merow et al., 2013). The default MaxEnt settings use all
+feature classes in model fitting except `threshold' features (depending
+on sample size of training data) and a regularisation multiplier of 1
+(Phillips et al., 2017). Several recent studies have highlighted the
+importance of species-specific tuning of feature classes and
+regularisation multipliers for optimising model complexity
+(Radosavljevic and Anderson, 2014; Shcheglovitova and Anderson, 2013),
+with default settings often producing models that are overly complex,
+and thus, may perform poorly when transferred (Elith et al., 2011; Merow
+et al., 2013). Despite the potential implications, surprisingly few
+studies use user-defined feature class and regularisation settings (e.g.
+Sutton, 2019), and as such, little is known about how model settings
+configurations may influence model transferability (Low et al., 2021;
+Moreno-Amat et al., 2015).
 
 In this study, we performed a retrospective assessment of the
-transferability of SDM's built using MaxEnt (\textbf{Phillips2017?}).
+transferability of SDM's built using MaxEnt (Phillips et al., 2017).
 Here, we modelled the potential climatic suitability of an insect that
@@
 Australia in 2013, to curb the spread of the invasive Australian
-\emph{Acacia mearnsii} De Wild (\textbf{Impson2013?}). Extensive field
+\emph{Acacia mearnsii} De Wild (Impson et al., 2013). Extensive field
 surveys have been performed in Australia (characterizing the native
-distribution of \emph{D. rubiformis}) (\textbf{Adair2004?}) and
-long-term post-release monitoring has been performed in South Africa to
-measure establishment rates of the insect, following experimental
-releases of the insect (Impson et al., 2021). As such, the availability
-of independent testing data from a separate continent to the data used
-to generate the models makes this an ideal study system for assessing
-the transferability of SDM's.
+distribution of \emph{D. rubiformis}) (Adair, 2004) and long-term
+post-release monitoring has been performed in South Africa to measure
+establishment rates of the insect, following experimental releases of
+the insect (Impson et al., 2021). As such, the availability of
+independent testing data from a separate continent to the data used to
+generate the models makes this an ideal study system for assessing the
+transferability of SDM's.
 
@@
 A total of 103 native-range records for \emph{D. rubiformis} were
-obtained from (\textbf{Adair2004?}). A total of 452 invaded-range
-records to evaluate model transferability were taken from Impson,
-Kleinjan and Hoffmann (unpublished). Distribution maps for \emph{D.
-rubiformis} for both native and invaded range are provided in Fig. 1.
+obtained from (Adair, 2004). A total of 452 invaded-range records to
+evaluate model transferability were taken from Impson, Kleinjan and
+Hoffmann (unpublished). Distribution maps for \emph{D. rubiformis} for
+both native and invaded range are provided in Fig. 1.
 
@@
 outputs. Filtering of species occurrence data may limit the inherent
-biases in the data and improve model quality (\textbf{Veloz2009?}). To
-avoid pseudo-replication, only one occurrence record per 2.5 minute grid
-cell was used for model calibration. Species occurrence datasets were
-thinned using the `spThin' package (\textbf{Aiello-Lammens2015?}), and
-spatial autocorrelation analyses were performed using the `ecospat'
-package (\textbf{DiCola2017?}).
+biases in the data and improve model quality (Veloz, 2009). To avoid
+pseudo-replication, only one occurrence record per 2.5 minute grid cell
+was used for model calibration. Species occurrence datasets were thinned
+using the `spThin' package (Aiello-Lammens et al., 2015), and spatial
+autocorrelation analyses were performed using the `ecospat' package (Di
+Cola et al., 2017).
 
@@
 Climate data were obtained by downloading the standard set of 19
-bioclimatic variables from the WorldClim ver. 1.4 database
-(\textbf{Hijmans2005?}) (data available at:
-www.worldclim.org/download.html). This dataset is representative of
-annual and seasonal means and variation of temperature and precipitation
-metrics averaged over the 1950--2000 time period (current climate) at a
-5 arc minute resolution.
+bioclimatic variables from the WorldClim ver. 1.4 database (Hijmans et
+al., 2005) (data available at: www.worldclim.org/download.html). This
+dataset is representative of annual and seasonal means and variation of
+temperature and precipitation metrics averaged over the 1950--2000 time
+period (current climate) at a 5 arc minute resolution.
 
@@
 predictors which were highly correlated (\textbar{}\emph{r}\textbar{}
-\textgreater{} 0.85) were excluded from the final predictor set
-(\textbf{Capinha2011?}). The reduced set of environmental predictors
+\textgreater{} 0.85) were excluded from the final predictor set (Capinha
+and Anastácio, 2011). The reduced set of environmental predictors
 consisted of seven climatic variables, including: bio1 -- annual mean
@@
 bio14 -- precipitation of the driest month and bio16 -- precipitation of
-the wettest quarter (see (\textbf{Hijmans2005?}) for further details).
+the wettest quarter (see Hijmans et al. (2005) for further details).
 
@@
 Maxent was implemented in the \texttt{dismo} package in \texttt{R}
-(\textbf{hijmans2017package?}). MaxEnt was selected as it consistently
-outperforms other modelling algorithms (\textbf{Wisz2008?}). Given that
+(Hijmans et al., 2017). MaxEnt was selected as it consistently
+outperforms other modelling algorithms (Wisz et al., 2008). Given that
 MaxEnt is a presence-only modelling algorithm, model calibration
@@
 (i.e.~background points). Background definition can have a significant
-effect on model output (\textbf{VanDerWal2009?}). The background should
+effect on model output (VanDerWal et al., 2009). The background should
 ideally represent the geographic areas available to the focal species,
 omitting areas where species absence is due to historical factors,
-dispersal constraints and/or biotic interactions (\textbf{Sanin2018?}).
-Following (\textbf{Webber2011?}), we defined the model background using
-the Koppen-Geiger climate classification (Available at:
+dispersal constraints and/or biotic interactions (Sanín and Anderson,
+2018). Following Webber et al. (2011), we defined the model background
+using the Koppen-Geiger climate classification (Available at:
 \url{http://koeppen-geiger.vu-wien.ac.at}). Only Koppen-Geiger climate
@@
 zones were intersected using the \texttt{extract} function from the
-\texttt{R} package \texttt{raster} (\textbf{hijmans2017package?}).
+\texttt{R} package \texttt{raster} (Hijmans et al., 2017).
 
@@
 (AUC\textsubscript{diff}), (3) omission rates (OR\textsubscript{10}),
-and (4) overall parsimony (AICc) (\textbf{Low2021?}).
+and (4) overall parsimony (AICc) (Low et al., 2021).
 
@@
   between 0.9 and 0.995 a good model, and \textgreater{} 0.995 an
-  excellent model (\textbf{Fielding1997?}). Thus, higher
+  excellent model (Fielding and Bell, 1997). Thus, higher
   AUC\textsubscript{test} values indicate increased ability to
@@
   calculated on training points only (AUC\textsubscript{train}) and
-  AUC\textsubscript{test} using cross-validation (\textbf{Warren2011?}).
-  Thus, higher AUC\textsubscript{diff} values indicate whether the
-  MaxEnt model is overfit on the training data, and thus, performs
+  AUC\textsubscript{test} using cross-validation (Warren and Seifert,
+  2011). Thus, higher AUC\textsubscript{diff} values indicate whether
+  the MaxEnt model is overfit on the training data, and thus, performs
   poorly when evaluated against testing points.
@@
   predicted to fall within the projected model surface once the model is
-  converted into a binary prediction output (\textbf{Boria2014?}).
+  converted into a binary prediction output (Boria et al., 2014).
   Overfit models have omission rates higher than the theoretical
-  expectation for the threshold applied (\textbf{Shcheglovitova2013?}).
-  OR10 sets the binary prediction threshold at a value that excludes the
-  10\% of the calibration localities from the model with the lowest
-  prediction values, and therefore has an expected omission rate of 0.10
-  (\textbf{Boria2014?}). As such, the OR\textsubscript{10} criterion
-  selected models calibrated with MaxEnt settings which best
+  expectation for the threshold applied (Shcheglovitova and Anderson,
+  2013). OR10 sets the binary prediction threshold at a value that
+  excludes the 10\% of the calibration localities from the model with
+  the lowest prediction values, and therefore has an expected omission
+  rate of 0.10 (Boria et al., 2014). As such, the OR\textsubscript{10}
+  criterion selected models calibrated with MaxEnt settings which best
   approximated the expected 0.10 omission rate. Models with omission
   rates increasingly higher than the expected value were considered as
-  more overfit (\textbf{Boria2017?}).
+  more overfit (Boria et al., 2017).
 \item
@@
   Information Criterion corrected for small sample sizes (AICc)
-  (i.e.~AICc=0; following (\textbf{Muscarella2014?})). The AICc
-  criterion simultaneously scores models according to their complexity
-  and goodness-of-fit, whereby models with the lowest AICc are selected
-  as the best models. AICc was used as the primary evaluation metric as
-  it is calculated using MaxEnt models built using the entire species
+  (i.e.~AICc=0; following Muscarella et al. (2014)). The AICc criterion
+  simultaneously scores models according to their complexity and
+  goodness-of-fit, whereby models with the lowest AICc are selected as
+  the best models. AICc was used as the primary evaluation metric as it
+  is calculated using MaxEnt models built using the entire species
   occurrence dataset, unlike AUC and OR\textsubscript{10} (and numerous
@@
   spatially biased due to the partitioning of the species occurrence
-  dataset into training and evaluation sets (\textbf{Sanin2018?}).
+  dataset into training and evaluation sets (Sanín and Anderson, 2018).
 \end{enumerate}
@@
 periods due to issues with extrapolation into novel/non-analogous
-climate (\textbf{Elith2009?}; \textbf{Elith2011?}; \textbf{Yates2018?}).
-To address these concerns, Multivariate environmental similarity
-surfaces (hereafter `MESS') (\textbf{Elith2010?}) were computed to
-assess whether MaxEnt models were extrapolating or interpolating. MESS
-analyses measure the similarity of any given point to a set of reference
-points (reference points were defined as only the species presence
-points used to calibrate models following \textbf{Kriticos2014?}).
-Negative MESS values indicate geographic regions outside the range of
-climate variables used to calibrate the model (i.e.~extrapolation space
-or MESS-), while MESS values between 0 and 100 indicate geographic
-regions inside the range of climatic variables used to calibrate the
-model (i.e.~interpolation space or MESS+). MESS maps are used as a
-measure of prediction uncertainty or caution against inferences in
-extrapolation space (\textbf{Elith2010?}).
+climate (Elith et al., 2011; Elith and Leathwick, 2009; Yates et al.,
+2018). To address these concerns, Multivariate environmental similarity
+surfaces (hereafter `MESS') (Elith et al., 2010) were computed to assess
+whether MaxEnt models were extrapolating or interpolating. MESS analyses
+measure the similarity of any given point to a set of reference points
+(reference points were defined as only the species presence points used
+to calibrate models following Kriticos et al., 2014). Negative MESS
+values indicate geographic regions outside the range of climate
+variables used to calibrate the model (i.e.~extrapolation space or
+MESS-), while MESS values between 0 and 100 indicate geographic regions
+inside the range of climatic variables used to calibrate the model
+(i.e.~interpolation space or MESS+). MESS maps are used as a measure of
+prediction uncertainty or caution against inferences in extrapolation
+space (Elith et al., 2010).
 
@@
 \texttt{extract} function from the \texttt{R} package \texttt{raster}
-(\textbf{hijmans2017package?}). GLM's were specified using the
-\texttt{rms} package in \texttt{R} (\textbf{Harrell2017package?}).
-Likelihood Ratio Tests (LRT) were performed to determine whether there
-was a statistical correlation between MaxEnt scores and the probability
-of \emph{D. rubiformis} establishment. A statistically significant,
-positive correlation between MaxEnt scores and insect establishment
-rates would provide support that the SDM's were useful and predictive of
-insect establishment, even when the insect was transferred in space.
-Out-of-sample AUC was calculated using the \texttt{rms} package, using
-500 bootstrap replicates (\textbf{Harrell2017package?}) to quantify the
-predictive performance of MaxEnt models, whereby AUC values less than
-0.80 are typically considered poor fitting models (lacks predictive
-power) (\textbf{Fielding1997?}).
+(Hijmans et al., 2017). GLM's were specified using the \texttt{rms}
+package in \texttt{R} (Harrell, 2017). Likelihood Ratio Tests (LRT) were
+performed to determine whether there was a statistical correlation
+between MaxEnt scores and the probability of \emph{D. rubiformis}
+establishment. A statistically significant, positive correlation between
+MaxEnt scores and insect establishment rates would provide support that
+the SDM's were useful and predictive of insect establishment, even when
+the insect was transferred in space. Out-of-sample AUC was calculated
+using the \texttt{rms} package, using 500 bootstrap replicates (Harrell,
+2017) to quantify the predictive performance of MaxEnt models, whereby
+AUC values less than 0.80 are typically considered poor fitting models
+(lacks predictive power) (Fielding and Bell, 1997).
 
@@
 \texttt{mess} function from the \texttt{dismo} package in \texttt{R}
-(\textbf{hijmans2017package?}). Interpolation versus extrapolation was
-fit as a categorical predictor and whether \emph{D. rubiformis} was
-recorded or not (present/absent) as a boolean response variable in a
-logistic GLM, using a logit link function.
+(Hijmans et al., 2017). Interpolation versus extrapolation was fit as a
+categorical predictor and whether \emph{D. rubiformis} was recorded or
+not (present/absent) as a boolean response variable in a logistic GLM,
+using a logit link function.
 
 All modelling and statistical analyses were conducted in \texttt{R} ver.
-4.0.3 (\textbf{RCoreTeam2020?}).
+4.0.3 (Team, 2020).
 
@@
 tools to support and guide the development and implementation of
-environmental policies and management programs (\textbf{Addison2013?};
-\textbf{Schuwirth2019?}). The usefulness of SDM's is dependent on model
+environmental policies and management programs (Addison et al., 2013;
+Schuwirth et al., 2019). The usefulness of SDM's is dependent on model
 performance, which is typically measured by its ability to correctly
@@
 between sites where the species is established versus a site where the
-species is absent (or assumed to be absent) (\textbf{Smith2021?}).
+species is absent (or assumed to be absent) (Smith et al., 2021).
 However, a number of authors have called in question the usefulness of
-SDM's, notably when models are transferred in space or time
-(\textbf{Yates2018?}; \textbf{Blasi2021?}). Poor transferability may
-lead to erroneous model predictions that may have serious negative
-consequences for the application of the model, and resulting policies
-and management programmes (Soultan and Safi, 2017). Here, we report on
-the validation of a SDM for an insect, (\emph{D. rubiformis}), native to
-Australia, that has been intentionally introduced as a biological
-control agent into South Africa to reduce the invasiveness of its
-host-plant, \emph{A. mearnsii} (Impson et al., 2013). The the
-availability of independent testing data from South Africa, a separate
-continent to the data used to generate the models (Australia), makes
-this an ideal study system for assessing the transferability of SDM's.
-This study represents one of only a handful of studies to validate SDM's
-using independent testing data (\textbf{Costa2010?};
-\textbf{Rebelo2010?}; \textbf{West2016?}; \textbf{Smith2021?}).
+SDM's, notably when models are transferred in space or time (Blasi et
+al., 2021; Yates et al., 2018). Poor transferability may lead to
+erroneous model predictions that may have serious negative consequences
+for the application of the model, and resulting policies and management
+programmes (Soultan and Safi, 2017). Here, we report on the validation
+of a SDM for an insect, (\emph{D. rubiformis}), native to Australia,
+that has been intentionally introduced as a biological control agent
+into South Africa to reduce the invasiveness of its host-plant, \emph{A.
+mearnsii} (Impson et al., 2013). The the availability of independent
+testing data from South Africa, a separate continent to the data used to
+generate the models (Australia), makes this an ideal study system for
+assessing the transferability of SDM's. This study represents one of
+only a handful of studies to validate SDM's using independent testing
+data (Costa et al., 2010; Rebelo and Jones, 2010; Smith et al., 2021;
+West et al., 2016).
 
@@
 Africa. AUC values greater than 0.80 are typically considered good
-fitting models (having high predictive power) (\textbf{Fielding1997?}).
+fitting models (having high predictive power) (Fielding and Bell, 1997).
 When MaxEnt models, calibrated using native-range records for \emph{D.
@@
 Default model is likely explained by default settings overfitting the
-model on the training points (\textbf{Warren2014?}). This finding
+model on the training points (Warren et al., 2014). This finding
 corroborates other studies showing that high predictive power in the
-model training area is not a guarantee of transferability
-(\textbf{Warren2014?}; \textbf{Duque-Lazo2016?}). We echo that
-recommendations of (\textbf{Shcheglovitova2013?}) and
-(\textbf{Muscarella2014?}) that model tuning should be performed to
-estimate optimal model complexity when transferring MaxEnt models in
-space and/or time in order to maximise their usefulness.
+model training area is not a guarantee of transferability (Duque-Lazo et
+al., 2016; Warren et al., 2014). We echo that recommendations of
+Shcheglovitova and Anderson (2013) and Muscarella et al. (2014) that
+model tuning should be performed to estimate optimal model complexity
+when transferring MaxEnt models in space and/or time in order to
+maximise their usefulness.
 
@@
 tools have been developed in recent years to assist SDM users with
-identifying model extrapolation, e.g.~MESS (\textbf{Elith2010?}) and
+identifying model extrapolation, e.g.~MESS (Elith et al., 2010) and
 ExDet (Mesgaran et al., 2014), and which should be an integral component
@@
 \hypertarget{refs}{}
-\begin{CSLReferences}{0}{0}
+\begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-Adair2004}{}%
+Adair, R.J., 2004. Seed-reducing {Cecidomyiidae} as potential biological
+control agents for invasive {Australian} wattles in {South Africa},
+particularly {\emph{Acacia}}{ \emph{Mearnsii}} and {\emph{A}}{\emph{.
+Cyclops}} (PhD thesis). {University of Cape Town}, {South Africa}.
+
+\leavevmode\hypertarget{ref-Addison2013}{}%
+Addison, P.F.E., Rumpff, L., Bau, S.S., Carey, J.M., Chee, Y.E., Jarrad,
+F.C., McBride, M.F., Burgman, M.A., 2013. Practical solutions for making
+models indispensable in conservation decision-making. Diversity Distrib.
+19, 490--502. \url{https://doi.org/10.1111/ddi.12054}
+
+\leavevmode\hypertarget{ref-Aiello-Lammens2015}{}%
+Aiello-Lammens, M.E., Boria, R.A., Radosavljevic, A., Vilela, B.,
+Anderson, R.P., 2015. {spThin}: An {R} package for spatial thinning of
+species occurrence records for use in ecological niche models. Ecography
+38, 541--545. \url{https://doi.org/10.1111/ecog.01132}
+
+\leavevmode\hypertarget{ref-Bahn2013}{}%
+Bahn, V., McGill, B.J., 2013. Testing the predictive performance of
+distribution models. Oikos 122, 321--331.
+\url{https://doi.org/10.1111/j.1600-0706.2012.00299.x}
+
+\leavevmode\hypertarget{ref-Blasi2021}{}%
+Blasi, M., Bartomeus, I., Bommarco, R., Gagic, V., Garratt, M.,
+Holzschuh, A., Kleijn, D., Lindström, S.A.M., Olsson, P., Polce, C.,
+Potts, S.G., Rundlöf, M., Scheper, J., Smith, H.G., Steffan‐Dewenter,
+I., Clough, Y., 2021. Evaluating predictive performance of statistical
+models explaining wild bee abundance in a mass‐flowering crop. Ecography
+ecog.05308. \url{https://doi.org/10.1111/ecog.05308}
+
+\leavevmode\hypertarget{ref-Bocedi2014}{}%
+Bocedi, G., Palmer, S.C.F., Pe'er, G., Heikkinen, R.K., Matsinos, Y.G.,
+Watts, K., Travis, J.M.J., 2014. {RangeShifter}: A platform for
+modelling spatial eco-evolutionary dynamics and species' responses to
+environmental changes. Methods in Ecology and Evolution 5, 388--396.
+\url{https://doi.org/10.1111/2041-210X.12162}
+
+\leavevmode\hypertarget{ref-Boria2017}{}%
+Boria, R.A., Olson, L.E., Goodman, S.M., Anderson, R.P., 2017. A
+single-algorithm ensemble approach to estimating suitability and
+uncertainty: Cross-time projections for four {Malagasy} tenrecs.
+Diversity and Distributions 23, 196--208.
+\url{https://doi.org/10.1111/ddi.12510}
+
+\leavevmode\hypertarget{ref-Boria2014}{}%
+Boria, R.A., Olson, L.E., Goodman, S.M., Anderson, R.P., 2014. Spatial
+filtering to reduce sampling bias can improve the performance of
+ecological niche models. Ecological Modelling 275, 73--77.
+\url{https://doi.org/10.1016/j.ecolmodel.2013.12.012}
+
+\leavevmode\hypertarget{ref-Capinha2011}{}%
+Capinha, C., Anastácio, P., 2011. Assessing the environmental
+requirements of invaders using ensembles of distribution models.
+Diversity and Distributions 17, 13--24.
+\url{https://doi.org/10.1111/j.1472-4642.2010.00727.x}
+
+\leavevmode\hypertarget{ref-Capinha2018}{}%
+Capinha, C., Essl, F., Seebens, H., Pereira, H.M., Kühn, I., 2018.
+Models of alien species richness show moderate predictive accuracy and
+poor transferability. NeoBiota 38, 77--96.
+\url{https://doi.org/10.3897/neobiota.38.23518}
+
+\leavevmode\hypertarget{ref-Costa2010}{}%
+Costa, G.C., Nogueira, C., Machado, R.B., Colli, G.R., 2010. Sampling
+bias and the use of ecological niche modeling in conservation planning:
+A field evaluation in a biodiversity hotspot. Biodivers Conserv 19,
+883--899. \url{https://doi.org/10.1007/s10531-009-9746-8}
+
+\leavevmode\hypertarget{ref-DiCola2017}{}%
+Di Cola, V., Broennimann, O., Petitpierre, B., Breiner, F.T., D'Amen,
+M., Randin, C., Engler, R., Pottier, J., Pio, D., Dubuis, A.,
+Pellissier, L., Mateo, R.G., Hordijk, W., Salamin, N., Guisan, A., 2017.
+Ecospat: An {R} package to support spatial analyses and modeling of
+species niches and distributions. Ecography 40, 774--787.
+\url{https://doi.org/10.1111/ecog.02671}
+
+\leavevmode\hypertarget{ref-Duque-Lazo2016}{}%
+Duque-Lazo, J., van Gils, H., Groen, T.A., Navarro-Cerrillo, R.M., 2016.
+Transferability of species distribution models: {The} case of
+{\emph{Phytophthora}}{ \emph{Cinnamomi}} in {Southwest Spain} and
+{Southwest Australia}. Ecological Modelling 320, 62--70.
+\url{https://doi.org/10.1016/j.ecolmodel.2015.09.019}
+
+\leavevmode\hypertarget{ref-Elith2006}{}%
+Elith, J., H. Graham*, C., P. Anderson, R., Dudík, M., Ferrier, S.,
+Guisan, A., J. Hijmans, R., Huettmann, F., R. Leathwick, J., Lehmann,
+A., Li, J., G. Lohmann, L., A. Loiselle, B., Manion, G., Moritz, C.,
+Nakamura, M., Nakazawa, Y., McC. M. Overton, J., Townsend Peterson, A.,
+J. Phillips, S., Richardson, K., Scachetti-Pereira, R., E. Schapire, R.,
+Soberón, J., Williams, S., S. Wisz, M., E. Zimmermann, N., 2006. Novel
+methods improve prediction of species' distributions from occurrence
+data. Ecography 29, 129--151.
+\url{https://doi.org/10.1111/j.2006.0906-7590.04596.x}
+
+\leavevmode\hypertarget{ref-Elith2010}{}%
+Elith, J., Kearney, M., Phillips, S., 2010. The art of modelling
+range-shifting species. Methods in Ecology and Evolution 1, 330--342.
+\url{https://doi.org/10.1111/j.2041-210X.2010.00036.x}
+
+\leavevmode\hypertarget{ref-Elith2009}{}%
+Elith, J., Leathwick, J., 2009. Species distribution models: Ecological
+explanation and prediction across space and time. Annual Review of
+Ecology, Evolution and Systematics 40, 677--697.
+\url{https://doi.org/10.1146/annurev.ecolsys.110308.120159}
+
+\leavevmode\hypertarget{ref-Elith2011}{}%
+Elith, J., Phillips, S.J., Hastie, T., Dudík, M., Chee, Y.E., Yates,
+C.J., 2011. A statistical explanation of {MaxEnt} for ecologists.
+Diversity and Distributions 17, 43--57.
+\url{https://doi.org/10.1111/j.1472-4642.2010.00725.x}
+
+\leavevmode\hypertarget{ref-Fielding1997}{}%
+Fielding, A.H., Bell, J.F., 1997. A review of methods for the assessment
+of prediction errors in conservation presence/absence models.
+Environmental Conservation 24, 38--49.
+\url{https://doi.org/10.1017/S0376892997000088}
+
+\leavevmode\hypertarget{ref-Guisan2013}{}%
+Guisan, A., Tingley, R., Baumgartner, J.B., Naujokaitis-Lewis, I.,
+Sutcliffe, P.R., Tulloch, A.I.T., Regan, T.J., Brotons, L.,
+McDonald-Madden, E., Mantyka-Pringle, C., Martin, T.G., Rhodes, J.R.,
+Maggini, R., Setterfield, S.A., Elith, J., Schwartz, M.W., Wintle, B.A.,
+Broennimann, O., Austin, M., Ferrier, S., Kearney, M.R., Possingham,
+H.P., Buckley, Y.M., 2013. Predicting species distributions for
+conservation decisions. Ecology Letters 16, 1424--1435.
+\url{https://doi.org/10.1111/ele.12189}
+
+\leavevmode\hypertarget{ref-Harrell2017package}{}%
+Harrell, F.E., 2017. Package {`rms.'} Vanderbilt University 229.
+
+\leavevmode\hypertarget{ref-Hijmans2005}{}%
+Hijmans, R.J., Cameron, S.E., Parra, J.L., Jones, P.G., Jarvis, A.,
+2005. Very high resolution interpolated climate surfaces for global land
+areas. International Journal of Climatology 25, 1965--1978.
+\url{https://doi.org/10.1002/joc.1276}
+
+\leavevmode\hypertarget{ref-hijmans2017package}{}%
+Hijmans, R.J., Phillips, S., Leathwick, J., Elith, J., Hijmans, M.R.J.,
+2017. Package {`dismo.'} Circles 9, 1--68.
+
+\leavevmode\hypertarget{ref-Impson2013}{}%
+Impson, F.A.C., Post, J.A., Hoffmann, J.H., 2013. Impact of the
+flower-galling midge, {\emph{Dasineura rubiformis}} {Kolesik}, on the
+growth of its host plant, {\emph{Acacia mearnsii}} {De Wild}, in {South
+Africa}. South African Journal of Botany 87, 118--121.
+\url{https://doi.org/10.1016/j.sajb.2013.04.006}
+
+\leavevmode\hypertarget{ref-Kriticos2014}{}%
+Kriticos, D., Morin, L., Webber, B., 2014. Taxonomic uncertainty in pest
+risks or modelling artefacts? {Implications} for biosecurity policy and
+practice. NB 23, 81--93. \url{https://doi.org/10.3897/neobiota.23.7496}
+
+\leavevmode\hypertarget{ref-Low2021}{}%
+Low, B.W., Zeng, Y., Tan, H.H., Yeo, D.C.J., 2021. Predictor complexity
+and feature selection affect {Maxent} model transferability: {Evidence}
+from global freshwater invasive species. Diversity and Distributions 27,
+497--511.
+
+\leavevmode\hypertarget{ref-Martin2020}{}%
+Martin, G.D., Magengelele, N.L., Paterson, I.D., Sutton, G.F., 2020.
+Climate modelling suggests a review of the legal status of {Brazilian}
+pepper {\emph{Schinus}}{ \emph{Terebinthifolia}} in {South Africa} is
+required. South African Journal of Botany 132, 95--102.
+\url{https://doi.org/10.1016/j.sajb.2020.04.019}
+
+\leavevmode\hypertarget{ref-Merow2013}{}%
+Merow, C., Smith, M.J., Silander Jr, J.A., 2013. A practical guide to
+{MaxEnt} for modeling species' distributions: What it does, and why
+inputs and settings matter. Ecography 36, 1058--1069.
+\url{https://doi.org/10.1111/j.1600-0587.2013.07872.x}
+
+\leavevmode\hypertarget{ref-Moreno-Amat2015}{}%
+Moreno-Amat, E., Mateo, R.G., Nieto-Lugilde, D., Morueta-Holme, N.,
+Svenning, J.-C., García-Amorena, I., 2015. Impact of model complexity on
+cross-temporal transferability in {Maxent} species distribution models:
+{An} assessment using paleobotanical data. Ecological Modelling 312,
+308--317. \url{https://doi.org/10.1016/j.ecolmodel.2015.05.035}
+
+\leavevmode\hypertarget{ref-Mukherjee2021}{}%
+Mukherjee, A., Banerjee, A.K., Raghu, S., 2021. Biological control of
+{\emph{Parkinsonia}}{ \emph{Aculeata}}: {Using} species distribution
+models to refine agent surveys and releases. Biological Control 159,
+104630. \url{https://doi.org/10.1016/j.biocontrol.2021.104630}
+
+\leavevmode\hypertarget{ref-Muscarella2014}{}%
+Muscarella, R., Galante, P.J., Soley-Guardia, M., Boria, R.A., Kass,
+J.M., Uriarte, M., Anderson, R.P., 2014. {ENMeval}: {An R} package for
+conducting spatially independent evaluations and estimating optimal
+model complexity for {Maxent} ecological niche models. Methods in
+Ecology and Evolution 5, 1198--1205.
+\url{https://doi.org/10.1111/2041-210X.12261}
+
+\leavevmode\hypertarget{ref-Phillips2017}{}%
+Phillips, S.J., Anderson, R.P., Dudík, M., Schapire, R.E., Blair, M.E.,
+2017. Opening the black box: An open-source release of {Maxent}.
+Ecography 40, 887--893. \url{https://doi.org/10.1111/ecog.03049}
+
+\leavevmode\hypertarget{ref-Radosavljevic2014}{}%
+Radosavljevic, A., Anderson, R.P., 2014. Making better {Maxent} models
+of species distributions: Complexity, overfitting and evaluation.
+Journal of Biogeography 41, 629--643.
+\url{https://doi.org/10.1111/jbi.12227}
+
+\leavevmode\hypertarget{ref-Randin2006}{}%
+Randin, C.F., Dirnböck, T., Dullinger, S., Zimmermann, N.E., Zappa, M.,
+Guisan, A., 2006. Are niche-based species distribution models
+transferable in space? Journal of Biogeography 33, 1689--1703.
+\url{https://doi.org/10.1111/j.1365-2699.2006.01466.x}
+
+\leavevmode\hypertarget{ref-Rebelo2010}{}%
+Rebelo, H., Jones, G., 2010. Ground validation of presence-only
+modelling with rare species: A case study on {\emph{Barbastella}}{
+\emph{Barbastellus}} ({Chiroptera}: {Vespertilionidae}). Journal of
+Applied Ecology 47, 410--420.
+\url{https://doi.org/10.1111/j.1365-2664.2009.01765.x}
+
+\leavevmode\hypertarget{ref-Sanin2018}{}%
+Sanín, C., Anderson, R.P., 2018. A framework for simultaneous tests of
+abiotic, biotic, and historical drivers of species distributions:
+Empirical tests for {North American Wood Warblers} based on climate and
+pollen. The American Naturalist 192, E48--E61.
+\url{https://doi.org/10.1086/697537}
+
+\leavevmode\hypertarget{ref-Schuwirth2019}{}%
+Schuwirth, N., Borgwardt, F., Domisch, S., Friedrichs, M., Kattwinkel,
+M., Kneis, D., Kuemmerlen, M., Langhans, S.D., Martínez-López, J.,
+Vermeiren, P., 2019. How to make ecological models useful for
+environmental management. Ecological Modelling 411, 108784.
+\url{https://doi.org/10.1016/j.ecolmodel.2019.108784}
+
+\leavevmode\hypertarget{ref-Shcheglovitova2013}{}%
+Shcheglovitova, M., Anderson, R.P., 2013. Estimating optimal complexity
+for ecological niche models: {A} jackknife approach for species with
+small sample sizes. Ecological Modelling 269, 9--17.
+\url{https://doi.org/10.1016/j.ecolmodel.2013.08.011}
+
+\leavevmode\hypertarget{ref-Smith2021}{}%
+Smith, J.N., Kelly, N., Renner, I.W., 2021. Validation of presence-only
+models for conservation planning and the application to whales in a
+multiple-use marine park. Ecological Applications 31, e02214.
+\url{https://doi.org/10.1002/eap.2214}
+
+\leavevmode\hypertarget{ref-Sutton2019b}{}%
+Sutton, G.F., 2019. Searching for a needle in a haystack: {Where} to
+survey for climatically-matched biological control agents for two
+grasses ({\emph{Sporobolus}} spp.) Invading {Australia}. Biological
+Control 129, 37--44.
+\url{https://doi.org/10.1016/j.biocontrol.2018.11.012}
+
+\leavevmode\hypertarget{ref-RCoreTeam2020}{}%
+Team, R.C., 2020. R: A Language and Environment for Statistical
+Computing. R Foundation for Statistical Computing, Vienna, Austria:
+Available at: https://www. R-project. org/.
+
+\leavevmode\hypertarget{ref-VanDerWal2009}{}%
+VanDerWal, J., Shoo, L.P., Graham, C., Williams, S.E., 2009. Selecting
+pseudo-absence data for presence-only distribution modeling: {How} far
+should you stray from what you know? Ecological Modelling 220, 589--594.
+\url{https://doi.org/10.1016/j.ecolmodel.2008.11.010}
+
+\leavevmode\hypertarget{ref-Veloz2009}{}%
+Veloz, S.D., 2009. Spatially autocorrelated sampling falsely inflates
+measures of accuracy for presence-only niche models. Journal of
+Biogeography 36, 2290--2299.
+\url{https://doi.org/10.1111/j.1365-2699.2009.02174.x}
+
+\leavevmode\hypertarget{ref-Warren2011}{}%
+Warren, D.L., Seifert, S.N., 2011. Ecological niche modeling in
+{Maxent}: The importance of model complexity and the performance of
+model selection criteria. Ecological Applications 21, 335--342.
+\url{https://doi.org/10.1890/10-1171.1}
+
+\leavevmode\hypertarget{ref-Warren2014}{}%
+Warren, D.L., Wright, A.N., Seifert, S.N., Shaffer, H.B., 2014.
+Incorporating model complexity and spatial sampling bias into ecological
+niche models of climate change risks faced by 90 {California} vertebrate
+species of concern. Diversity and Distributions 20, 334--343.
+\url{https://doi.org/10.1111/ddi.12160}
+
+\leavevmode\hypertarget{ref-Webber2011}{}%
+Webber, B.L., Yates, C.J., Le Maitre, D.C., Scott, J.K., Kriticos, D.J.,
+Ota, N., McNeill, A., Le Roux, J.J., Midgley, G.F., 2011. Modelling
+horses for novel climate courses: Insights from projecting potential
+distributions of native and alien {Australian} acacias with correlative
+and mechanistic models. Diversity and Distributions 17, 978--1000.
+\url{https://doi.org/10.1111/j.1472-4642.2011.00811.x}
+
+\leavevmode\hypertarget{ref-Wenger2012}{}%
+Wenger, S.J., Olden, J.D., 2012. Assessing transferability of ecological
+models: An underappreciated aspect of statistical validation. Methods in
+Ecology and Evolution 3, 260--267.
+\url{https://doi.org/10.1111/j.2041-210X.2011.00170.x}
+
+\leavevmode\hypertarget{ref-West2016}{}%
+West, A.M., Kumar, S., Brown, C.S., Stohlgren, T.J., Bromberg, J., 2016.
+Field validation of an invasive species {Maxent} model. Ecological
+Informatics 36, 126--134.
+\url{https://doi.org/10.1016/j.ecoinf.2016.11.001}
+
+\leavevmode\hypertarget{ref-Wisz2008}{}%
+Wisz, M.S., Hijmans, R.J., Li, J., Peterson, A.T., Graham, C.H., Guisan,
+A., Group, N.P.S.D.W., 2008. Effects of sample size on the performance
+of species distribution models. Diversity and Distributions 14,
+763--773. \url{https://doi.org/10.1111/j.1472-4642.2008.00482.x}
+
+\leavevmode\hypertarget{ref-Yates2018}{}%
+Yates, K.L., Bouchet, P.J., Caley, M.J., Mengersen, K., Randin, C.F.,
+Parnell, S., Fielding, A.H., Bamford, A.J., Ban, S., Barbosa, A.M.,
+Dormann, C.F., Elith, J., Embling, C.B., Ervin, G.N., Fisher, R., Gould,
+S., Graf, R.F., Gregr, E.J., Halpin, P.N., Heikkinen, R.K., Heinänen,
+S., Jones, A.R., Krishnakumar, P.K., Lauria, V., Lozano-Montes, H.,
+Mannocci, L., Mellin, C., Mesgaran, M.B., Moreno-Amat, E., Mormede, S.,
+Novaczek, E., Oppel, S., Ortuño Crespo, G., Peterson, A.T., Rapacciuolo,
+G., Roberts, J.J., Ross, R.E., Scales, K.L., Schoeman, D., Snelgrove,
+P., Sundblad, G., Thuiller, W., Torres, L.G., Verbruggen, H., Wang, L.,
+Wenger, S., Whittingham, M.J., Zharikov, Y., Zurell, D., Sequeira,
+A.M.M., 2018. Outstanding challenges in the transferability of
+ecological models. Trends in Ecology \& Evolution 33, 790--802.
+\url{https://doi.org/10.1016/j.tree.2018.08.001}
+
 \end{CSLReferences}

### 2022-02-07 R/scripts/figure-4-make.R
Make Figure 4 + 5

@@ -1,0 +1,433 @@
+# -----------------------------------------------------------------------------
+# Script: - Reproduce Figure 4
+#
+# AUTHOR: Guy F. Sutton
+# AFFILIATION: Centre for Biological Control, Rhodes University, South Africa
+# DATE MODIFIED: 03/02/2021
+# CONTACT: g.sutton@ru.ac.za
+# -----------------------------------------------------------------------------
+
+# -----------------------------------------------------------------------------
+# Session setup
+# -----------------------------------------------------------------------------
+
+# Load required packages
+if (!require("pacman"))
+  install.packages("pacman")
+pacman::p_load(
+  tidyverse,
+  spocc,
+  ggmap,
+  maptools,
+  maps,
+  ggplot2,
+  scrubr,
+  mapr,
+  tidyr,
+  stringr,
+  rnaturalearth,
+  rnaturalearthdata,
+  rlang,
+  sf,
+  ggspatial,
+  raster,
+  here,
+  tidytext,
+  ggtext,
+  tidymodels,
+  themis,
+  ranger,
+  yardstick,
+  vip,
+  DHARMa,
+  rms
+)
+
+# ----------------------------------------------------------------------------
+# Evaluate performance of default settings model: 
+# ----------------------------------------------------------------------------
+
+# Import Maxent prediction raster 
+clim_map <- raster::raster("./models/rasters/raster_default.asc")
+
+# Import invaded-range GPS points 
+inv_data <- readr::read_csv("./data/data_clean/dasi_rubi_invaded_range_gps_clean.csv")
+head(inv_data)
+
+# Keep just the columns of interest
+inv_clean <- inv_data %>%
+  dplyr::select(lon = lon_converted,
+                lat = lat_converted,
+                established) %>%
+  mutate(lat = lat * -1)
+inv_clean
+
+# Keep only the sites with establishment data
+inv_clean <- inv_clean %>%
+  tidyr::drop_na(established)
+
+# Extract the maxent scores for each site
+clim_data <- raster::extract(clim_map,
+                             inv_clean[, 1:2]) %>%
+  as.data.frame()
+
+# Bind the maxent scores to the site data
+clim_data <- dplyr::bind_cols(inv_clean, clim_data) %>%
+  dplyr::rename(clim_match = 4) %>%
+  tidyr::drop_na(clim_match) %>%
+  dplyr::mutate(establsihed = as.numeric(established))
+head(clim_data)
+
+# Run logistic regression
[…]

### 2022-02-07 R/scripts/figure-5-make.R

@@ -1,5 +1,5 @@
 # -----------------------------------------------------------------------------
-# Script: - Reproduce Figure 6
+# Script: - Reproduce Figure 5
 #
 # AUTHOR: Guy F. Sutton
 # AFFILIATION: Centre for Biological Control, Rhodes University, South Africa
@@ -79,7 +79,6 @@ theme_opts <- list(
 # ----------------------------------------------------------------------------
 # Plot default map: 
 # ----------------------------------------------------------------------------
-
 
 # Import Maxent prediction raster
 clim_map <- raster::raster("./models/rasters/raster_default.asc")
@@ -391,7 +390,7 @@ all_maps
 
 # Save the plot to your PC
 ggsave(
-  "./ms_body/ms_figs/fig_6_maxent_maps.png",
+  "./ms_body/ms_figs/fig_5_maxent_maps.png",
   width = 18,
   height = 12,
   dpi = 600

### 2022-02-07 R/scripts/figure-6-make.R
Add MESS analyses

@@ -1,0 +1,383 @@
+# -----------------------------------------------------------------------------
+# Script: - Reproduce Figure 6
+#
+# AUTHOR: Guy F. Sutton
+# AFFILIATION: Centre for Biological Control, Rhodes University, South Africa
+# DATE MODIFIED: 03/02/2021
+# CONTACT: g.sutton@ru.ac.za
+# -----------------------------------------------------------------------------
+
+# -----------------------------------------------------------------------------
+# Session setup
+# -----------------------------------------------------------------------------
+
+# Load packages
+if (!require("pacman"))
+  install.packages("pacman")
+pacman::p_load(
+  tidyverse,
+  spocc,
+  ggmap,
+  maptools,
+  maps,
+  ggplot2,
+  scrubr,
+  mapr,
+  tidyr,
+  stringr,
+  patchwork,
+  rnaturalearth,
+  rnaturalearthdata,
+  rlang,
+  sf,
+  ggspatial,
+  raster,
+  here,
+  tidytext,
+  ggtext,
+  dismo,
+  tidymodels,
+  themis,
+  ranger,
+  viridis,
+  yardstick,
+  vip,
+  DHARMa,
+  rms
+)
+
+# Set the theme for the figures
+theme_set(
+  theme_classic() +
+    theme(
+      panel.border = element_rect(colour = "black", fill = NA),
+      axis.text = element_text(colour = "black"),
+      axis.title.x = element_text(margin = unit(c(2, 0, 0, 0), "mm")),
+      axis.title.y = element_text(margin = unit(c(0, 4, 0, 0), "mm")),
+      legend.position = "none"
+    )
+)
+
+# Set the theme for the maps
+theme_opts <- list(
+  theme(
+    panel.grid.minor = element_blank(),
+    panel.grid.major = element_blank(),
+    panel.background = element_rect(fill = 'white', colour = NA),
+    plot.background = element_rect(),
+    axis.line = element_blank(),
+    axis.text.x = element_text(colour = "black"),
+    axis.text.y = element_text(colour = "black"),
+    axis.ticks = element_line(colour = "black"),
+    axis.title.x = element_text(colour = "black"),
+    axis.title.y = element_text(colour = "black"),
+    plot.title = element_text(colour = "black"),
+    panel.border = element_rect(fill = NA),
+    legend.key = element_blank()
+  )
+)
+
+# -----------------------------------------------------------------------------
+# Load WORLDCLIM layers:
+#
+# The standard 19 WORLDCLIM layers are imported here to be used as covariates.
[…]

### 2022-02-08 ms_body/manuscript.Rmd
Add BibTex references to MS
Update AICc model (originally picked wrong rm value)

@@
 
-Sutton, G. F.^1^$\dag$; Martin, G.D. ^1,^^2^
+Sutton, G. F.^1^$\dag$; Martin, G.D.^1,^^2^
 
@@
 The five models differed in the feature classes and regularisation multiplier settings used by MaxEnt to build the models.
-The five models were specified with (1) default MaxEnt settings, and four optimised settings configurations that (2) maximised performance based on AUC~test~, (3) AUC~diff~, (4) OR~10~ and (5) AICc (Table S1).
-The AUC values for these five models ranged were ranged between 0.82 and 0.97, depending on the MaxEnt settings configuration used.
+The five models were specified with (1) default MaxEnt settings, and four optimised settings configurations that (2) maximised model performance based on AUC~test~, (3) AUC~diff~, (4) OR~10~ and (5) AICc (Table S1).
+The AUC values for these five models ranged between 0.89 - 0.92, depending on the MaxEnt settings configuration used.
 As such, all five candidate models  were considered useful and had high predictive accuracy in distinguishing between *D.
 rubiformis* recorded presences and background points in its native range.
-While default MaxEnt settings produced the highest discriminatory power when tested in the training area of Australia (AUC = 0.97), models calibrated with tuned settings configurations led to less overfitting (lower AUC~diff~), better omission rates (best approximated the expectation for OR~10~), and more parsimonious models (AICc) (Fig.
+While default MaxEnt settings produced high discriminatory power when tested in the training area of Australia (AUC = 0.89), models calibrated with tuned settings configurations led to less overfitting (lower AUC~diff~), better omission rates (best approximated the expectation for OR~10~), and more parsimonious models (AICc) (Fig.
 2).
 
-## 3.2. Model transferability
+## 3.2. Model performance
 
-When evaluated on independent testing data in South Africa, the MaxEnt models showed relatively high, but variable, predictive accuracy, depending on the MaxEnt settings used to calibrate the model.
+When evaluated on independent testing data in South Africa, the MaxEnt models showed relatively high predictive accuracy, irrespective of the MaxEnt settings used to calibrate the model.
 For all models, the probability that *D.
@@
 3).
-The default MaxEnt model found no statistically significant relationship between MaxEnt climatic suitability scores and the probability of *D.
+All five MaxEnt models demonstrated a statistically significant positive relationship between MaxEnt suitability scores and the probability of *D.
 rubiformis* establishment in South Africa (Table.
-1), and displayed low power to discriminate between sites where *D.
-rubiformis* had established or not (AUC: 0.64 $\pm$ 0.01) (Fig.
-4).
-In contrast, the four models calibrated using optimised model settings configurations all demonstrated a statistically significant positive relationship between MaxEnt suitability scores and the probability of *D.
-rubiformis* establishment in South Africa (Table.
-1).
-All four tuned models displayed high power to discriminate between sites where *D.
-rubiformis* had established or not (AUC: 0.83 - 0.84) than the MaxEnt model calibrated with default settings (AUC: 0.64 $\pm$ 0.01) (Fig.
+1), and high power to discriminate between sites where *D.
+rubiformis* had established or not (AUC: 0.79 - 0.91) (Fig.
 4).
 The predicted climatic suitability for *D.
-rubiformis* across South Africa varied depending on the MaxEnt settings used to calibrate the model (Fig.
+rubiformis* across South Africa did vary somewhat depending on the MaxEnt settings used to calibrate the model (Fig.
 5).
-Default MaxEnt settings yielded higher climatic suitability estimates for *D.
-rubiformis* than the other MaxEnt models with tuned model settings, notably in regions where *D.
-rubiformis* is absent (e.g.
-the north-west coastal region bordering Namibia).
+All MaxEnt models predicted relatively high climatic suitability for *D.
+rubiformis* in the far south-west of South Africa and in the southern Cape region.
+However, the four MaxEnt models calibrated with tuned model settings differed from the default settings model by predicting climatic suitability for *D.
+rubiformis* along the entire south and east Coast of South Africa.
 
@@
 rubiformis* over the vast majority of South Africa (Fig.
-S1).
+6a).
 Indeed, 54 of the 78 field sites surveyed for *D.
@@
 *Dasineura rubiformis* established at 23 of 24 sites in interpolation (96%), while it established at 44 of 54 sites in extrapolation (82%) (Fig.
-6).
+6b).
 We found no evidence that a site being in extrapolation/interpolation was a statistically significant predictor of *D.
 rubiformis* establishment in South Africa (X~2~ = 3.40, d.f = 1, *P* = 0.07).
-Moreover, classifying sites as being in extrapolation/interpolation yielded lower power to discriminate to discriminate between sites where *D.
-rubiformis* had established or not (AUC: 0.67 $\pm$ 0.02) than all of the MaxEnt models, barring the model with default settings.
+Moreover, classifying sites as being in extrapolation/interpolation yielded lower power to discriminate between sites where *D.
+rubiformis* had established or not (AUC: 0.63) than all of the MaxEnt models.
 
@@
 rubiformis* in South Africa.
-AUC values greater than 0.80 are typically considered good fitting models (having high predictive power) [@Fielding1997].
+AUC values greater than 0.75 are typically considered good fitting models (having high predictive power) [@Fielding1997].
 When MaxEnt models, calibrated using native-range records for *D.
@@
 
-<div id="refs" ></div>
+<div id="refs"></div>
 

### 2022-02-08 R/scripts/02-maxent-native-range.R

@@ -695,7 +695,7 @@ model_OR10 <- dismo::maxent(
 )
 
 # Model 5: Optimise AICc
-# - The top performing model by AICc had rm = 8 and only H features.
+# - The top performing model by AICc had rm = 5 and only H features.
 # Fit model with optimally tuned settings based on AICc
 model_AICc <- dismo::maxent(
   x = trainData,
@@ -704,7 +704,7 @@ model_AICc <- dismo::maxent(
                './models/models/optimal_settings_AICc',
                sep = ''),
   args = c(
-    'betamultiplier=8.0',
+    'betamultiplier=5.0',
     'linear=false',
     'quadratic=false',
     'product=false',

### 2022-02-08 R/scripts/04-mess-analyses.R

@@ -1,0 +1,220 @@
+# -----------------------------------------------------------------------------
+# Script: 04 - MESS analysis  
+#
+# AUTHOR: Guy F. Sutton
+# AFFILIATION: Centre for Biological Control, Rhodes University, South Africa
+# DATE MODIFIED: 03/02/2021
+# CONTACT: g.sutton@ru.ac.za
+# -----------------------------------------------------------------------------
+
+
+# -----------------------------------------------------------------------------
+# Session setup
+# -----------------------------------------------------------------------------
+
+# Load required packages
+if (!require("pacman"))
+  install.packages("pacman")
+pacman::p_load(
+  tidyverse,
+  dismo,
+  raster,
+  here,
+  corrplot,
+  Hmisc,
+  patchwork,
+  ecospat,
+  gridSVG,
+  gridExtra,
+  grid,
+  ENMeval,
+  spThin,
+  viridis,
+  viridisLite,
+  mapdata,
+  maptools,
+  scales,
+  geosphere,
+  rgdal,
+  ggtext,
+  rJava,
+  rgeos,
+  sp,
+  sf,
+  ggspatial,
+  ecospat,
+  rnaturalearth,
+  rnaturalearthdata,
+  rms
+)
+
+# Change ggplot theme
+theme_set(
+  theme_classic() +
+    theme(
+      panel.border = element_rect(colour = "black",
+                                  fill = NA),
+      axis.text = element_text(colour = "black"),
+      axis.title.x = element_text(margin = unit(c(2, 0, 0, 0),
+                                                "mm")),
+      axis.title.y = element_text(margin = unit(c(0, 4, 0, 0),
+                                                "mm")),
+      legend.position = "none"
+    )
+)
+
+# Set the theme for the maps
+theme_opts <- list(
+  theme(
+    panel.grid.minor = element_blank(),
+    panel.grid.major = element_blank(),
+    panel.background = element_rect(fill = 'white', colour = NA),
+    plot.background = element_rect(),
+    axis.line = element_blank(),
+    axis.text.x = element_text(colour = "black"),
+    axis.text.y = element_text(colour = "black"),
+    axis.ticks = element_line(colour = "black"),
+    axis.title.x = element_text(colour = "black"),
+    axis.title.y = element_text(colour = "black"),
+    plot.title = element_text(colour = "black"),
+    panel.border = element_rect(fill = NA),
+    legend.key = element_blank()
+  )
+)
+
+
+# -----------------------------------------------------------------------------
[…]

### 2022-02-08 R/scripts/figure-6-make.R

@@ -161,10 +161,6 @@ mss <- dismo::mess(x = mess_pred,
 # Raster should be the MESS map raster
 test <- mss
 test <- raster::mask(test, world)
-
-tif=stars(rcp)
-sf=st_as_sf(tif)
-
 rcp <- raster::rasterToPoints(test)
 rcpdf <- base::data.frame(rcp)
 colnames(rcpdf) <- c("Longitude", "Latitude", "Species")

### 2022-02-08 ms_body/manuscript.Rmd
Make Figure 7

@@
 rubiformis* in South Africa.
-However, we found that this pattern held only when MaxEnt models were specified with user-defined settings configurations.
-A MaxEnt model calibrated with default settings was overfit to the training data, and performed poorly when transfered in space.
-Our study demonstrates that MaxEnt models can prove useful when transferred in space, however, user-optimised model settings likely need to be used to control model complexity, and reduce the likelihood of overfitting.
+MaxEnt model outputs and projections were relatively consistent, irrespective of the settings used.
+Despite MaxEnt having to extrapolate over much of South Africa, MaxEnt performance was higher than modelling climatic suitability as a function of whether models were in extrapolation or interpolation.
+Our study demonstrates that MaxEnt models can prove useful when transferred in space, however, users need to be aware of the effect of different model settings configurations and model extrapolation can have on model transferability and usefulness in applied ecological settings.
 
@@
 To avoid pseudo-replication, only one occurrence record per 2.5 minute grid cell was used for model calibration.
-Species occurrence datasets were thinned using the ‘spThin’ package [@Aiello-Lammens2015], and spatial autocorrelation analyses were performed using the ‘ecospat’ package [@DiCola2017].
+Species occurrence datasets were thinned using the `spThin` package [@Aiello-Lammens2015], and spatial autocorrelation analyses were performed using the `ecospat` package [@DiCola2017].
 
@@
 
-Climate data were obtained by downloading the standard set of 19 bioclimatic variables from the WorldClim ver.
-1.4 database [@Hijmans2005] (data available at: www.worldclim.org/download.html).
-This dataset is representative of annual and seasonal means and
-variation of temperature and precipitation metrics averaged over the 1950–2000 time period (current climate) at a 5 arc minute resolution.
+Climate data were obtained by downloading the standard set of 19 bioclimatic variables from the WorldClim ver. 1.4 database [@Hijmans2005]  
+(data available at:  www.worldclim.org/download.html).
+This dataset is representative of annual and seasonal means and variation of temperature and precipitation metrics averaged over the 1950–2000 time period (current climate) at a 5 arc minute resolution.
 
@@
 When MaxEnt models, calibrated using native-range records for *D.
-rubiformis* from Australia, were projected over South Africa, they had AUC scores of 0.83 - 0.84, indicative of high power to discriminate between sites where *D.
+rubiformis* from Australia, were projected over South Africa, they had AUC scores of 0.89 - 0.92, indicative of high power to discriminate between sites where *D.
 rubiformis* had established in South Africa or not.
@@
 Our study adds to a growing body of literature indicating that SDM model complexity affects model performance and transferability when being projected into novel geographic regions.
-While the default MaxEnt model had the highest predictive power in the training area (Australia), this model performed extremely poorly when transferred over South Africa.
-In contrast, the four MaxEnt models calibrated with tuned settings showed high predictive power when transferred.
-The poor transferability of the Default model is likely explained by default settings overfitting the model on the training points [@Warren2014].
+While the default MaxEnt model had the high predictive power in the training area (Australia), this model was somewhat overfit to the training data [@Warren2014].
 This finding corroborates other studies showing that high predictive power in the model training area is not a guarantee of transferability [@Warren2014; @Duque-Lazo2016].
@@
 To visualise the potential geographic regions where *D.
-rubiformis* may establish in South Africa, we thresholded the MaxEnt map that optimised AICc where the probability of establishment was > 50% (i.e.
+rubiformis* may establish in South Africa, we thresholded the MaxEnt map that optimised AICc where the probability of establishment was > 75% (i.e.
 MaxEnt score = 0.30), given in Fig.
-9.
+7.
 The map demonstrates that the entire south and east coast regions of South Africa are climatically suitable for *D.
@@
 As such, our models provide a statistically validated approach to optimising the redistribution of *D.
-rubiformis* in South Africa by prioritising releases at field sites where MaxEnt scores are > 0.30, and therefore where the probability of establishment was > 50%.
+rubiformis* in South Africa by prioritising releases at field sites where MaxEnt scores are > 0.30, and therefore where the probability of establishment was > 75%.
 
@@
 Our study is one of only a few studies to date validating MaxEnt models using spatially independent data.
-However, the accuracy of the MaxEnt models was dependent on the settings used to calibrate the models, with default MaxEnt programme settings having low predictive power.
-Our results demonstrate that MaxEnt can be useful in predicting where a species may establish, even when projected into a novel geographic area, but that users need to be aware of how model settings can influence model outputs and usefulness.
+However, the accuracy of the MaxEnt models and their climatic suitability projections were dependent on the settings used to calibrate the models.
+Our results demonstrate that MaxEnt can be useful in predicting where a species may establish, even when projected into a novel geographic area, but that users need to be aware of how model settings and model extrapolation can influence model outputs and usefulness.
 Demonstrating how SDM's can be useful, while communicating uncertainty in their predictions and outputs, is essential for building confidence in the use of SDM's (Blasi et al.
@@
 We showed how validating predictive models can be used to guide the implementation of a biological control programme by optimising the redistribution and release of *D.
-rubiformis* in geographic areas that are climatically suitable and where the probability of the insect establishing is greater than 50%.
+rubiformis* in geographic areas that are climatically suitable and where the probability of the insect establishing is greater than 75%.
 
@@
 ```{r import_fig_1, echo = FALSE}
-# knitr::include_graphics(here::here("./ms_body/ms_figs/fig_1.png"), dpi = NA)
+knitr::include_graphics(here::here("./ms_body/ms_figs/fig_1_distributions.png"), dpi = NA)
 ```
@@
 ```{r import_fig_2, echo = FALSE}
-# knitr::include_graphics(here::here("./ms_body/ms_figs/fig_2.png"), dpi = NA)
+knitr::include_graphics(here::here("./ms_body/ms_figs/fig_2_model_tuning.png"), dpi = NA)
 ```
@@
 ```{r import_fig_3, echo = FALSE}
-# knitr::include_graphics(here::here("./ms_body/ms_figs/fig_3.png"), dpi = NA)
+knitr::include_graphics(here::here("./ms_body/ms_figs/fig_3_logits.png"), dpi = NA)
 ```
@@
 ```{r import_fig_4, echo = FALSE}
-# knitr::include_graphics(here::here("./ms_body/ms_figs/fig_4.png"), dpi = NA)
+knitr::include_graphics(here::here("./ms_body/ms_figs/fig_4_auc.png"), dpi = NA)
 ```
@@
 ```{r import_fig_5, echo = FALSE}
-# knitr::include_graphics(here::here("./ms_body/ms_figs/fig_5.png"), dpi = NA)
+knitr::include_graphics(here::here("./ms_body/ms_figs/fig_5_maxent_maps.png"), dpi = NA)
 ```
@@
 ```{r import_fig_6, echo = FALSE}
-# knitr::include_graphics(here::here("./ms_body/ms_figs/fig_6.png"), dpi = NA)
+knitr::include_graphics(here::here("./ms_body/ms_figs/fig_6_mess.png"), dpi = NA)
 ```
@@
     footnote(
-    alphabet = c("Adjusted to indicate change in Y for each 1% unit increase in covariate"),
+    alphabet = c("Adjusted to indicate change in Y for each 1% unit increase in climatic similarity"),
     escape = TRUE,
@@
       "AUC\\textsubscript{diff}",
-      "OR\\textsubscript{10}"
+      "OR\\textsubscript{10}",
+      "AIC\\textsubscript{c}"
     )

### 2022-02-08 ms_body/manuscript.tex
Make Figure 7

@@
 \setcounter{secnumdepth}{0}
+
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
 
@@
 MaxEnt scores were positively correlated with increased establishment
-rates for \emph{D. rubiformis} in South Africa. However, we found that
-this pattern held only when MaxEnt models were specified with
-user-defined settings configurations. A MaxEnt model calibrated with
-default settings was overfit to the training data, and performed poorly
-when transfered in space. Our study demonstrates that MaxEnt models can
-prove useful when transferred in space, however, user-optimised model
-settings likely need to be used to control model complexity, and reduce
-the likelihood of overfitting.
+rates for \emph{D. rubiformis} in South Africa. MaxEnt model outputs and
+projections were relatively consistent, irrespective of the settings
+used. Despite MaxEnt having to extrapolate over much of South Africa,
+MaxEnt performance was higher than modelling climatic suitability as a
+function of whether models were in extrapolation or interpolation. Our
+study demonstrates that MaxEnt models can prove useful when transferred
+in space, however, users need to be aware of the effect of different
+model settings configurations and model extrapolation can have on model
+transferability and usefulness in applied ecological settings.
 
@@
 was used for model calibration. Species occurrence datasets were thinned
-using the `spThin' package (Aiello-Lammens et al., 2015), and spatial
-autocorrelation analyses were performed using the `ecospat' package (Di
-Cola et al., 2017).
+using the \texttt{spThin} package (Aiello-Lammens et al., 2015), and
+spatial autocorrelation analyses were performed using the
+\texttt{ecospat} package (Di Cola et al., 2017).
 
@@
 bioclimatic variables from the WorldClim ver. 1.4 database (Hijmans et
-al., 2005) (data available at: www.worldclim.org/download.html). This
-dataset is representative of annual and seasonal means and variation of
-temperature and precipitation metrics averaged over the 1950--2000 time
-period (current climate) at a 5 arc minute resolution.
+al., 2005)\\
+(data available at: www.worldclim.org/download.html). This dataset is
+representative of annual and seasonal means and variation of temperature
+and precipitation metrics averaged over the 1950--2000 time period
+(current climate) at a 5 arc minute resolution.
 
@@
 (1) default MaxEnt settings, and four optimised settings configurations
-that (2) maximised performance based on AUC\textsubscript{test}, (3)
-AUC\textsubscript{diff}, (4) OR\textsubscript{10} and (5) AICc (Table
-S1). The AUC values for these five models ranged were ranged between
-0.82 and 0.97, depending on the MaxEnt settings configuration used. As
-such, all five candidate models were considered useful and had high
-predictive accuracy in distinguishing between \emph{D. rubiformis}
-recorded presences and background points in its native range. While
-default MaxEnt settings produced the highest discriminatory power when
-tested in the training area of Australia (AUC = 0.97), models calibrated
-with tuned settings configurations led to less overfitting (lower
+that (2) maximised model performance based on AUC\textsubscript{test},
+(3) AUC\textsubscript{diff}, (4) OR\textsubscript{10} and (5) AICc
+(Table S1). The AUC values for these five models ranged between 0.89 -
+0.92, depending on the MaxEnt settings configuration used. As such, all
+five candidate models were considered useful and had high predictive
+accuracy in distinguishing between \emph{D. rubiformis} recorded
+presences and background points in its native range. While default
+MaxEnt settings produced high discriminatory power when tested in the
+training area of Australia (AUC = 0.89), models calibrated with tuned
+settings configurations led to less overfitting (lower
 AUC\textsubscript{diff}), better omission rates (best approximated the
@@
 
-\hypertarget{model-transferability}{%
-\subsection{3.2. Model transferability}\label{model-transferability}}
+\hypertarget{model-performance}{%
+\subsection{3.2. Model performance}\label{model-performance}}
 
 When evaluated on independent testing data in South Africa, the MaxEnt
-models showed relatively high, but variable, predictive accuracy,
-depending on the MaxEnt settings used to calibrate the model. For all
-models, the probability that \emph{D. rubiformis} was recorded at a
-field site in South Africa increased with higher MaxEnt climatic
-suitability scores (Fig. 3). The default MaxEnt model found no
-statistically significant relationship between MaxEnt climatic
-suitability scores and the probability of \emph{D. rubiformis}
-establishment in South Africa (Table. 1), and displayed low power to
-discriminate between sites where \emph{D. rubiformis} had established or
-not (AUC: 0.64 \(\pm\) 0.01) (Fig. 4). In contrast, the four models
-calibrated using optimised model settings configurations all
-demonstrated a statistically significant positive relationship between
-MaxEnt suitability scores and the probability of \emph{D. rubiformis}
-establishment in South Africa (Table. 1). All four tuned models
-displayed high power to discriminate between sites where \emph{D.
-rubiformis} had established or not (AUC: 0.83 - 0.84) than the MaxEnt
-model calibrated with default settings (AUC: 0.64 \(\pm\) 0.01) (Fig.
-4). The predicted climatic suitability for \emph{D. rubiformis} across
-South Africa varied depending on the MaxEnt settings used to calibrate
-the model (Fig. 5). Default MaxEnt settings yielded higher climatic
-suitability estimates for \emph{D. rubiformis} than the other MaxEnt
-models with tuned model settings, notably in regions where \emph{D.
-rubiformis} is absent (e.g.~the north-west coastal region bordering
-Namibia).
+models showed relatively high predictive accuracy, irrespective of the
+MaxEnt settings used to calibrate the model. For all models, the
+probability that \emph{D. rubiformis} was recorded at a field site in
+South Africa increased with higher MaxEnt climatic suitability scores
+(Fig. 3). All five MaxEnt models demonstrated a statistically
+significant positive relationship between MaxEnt suitability scores and
+the probability of \emph{D. rubiformis} establishment in South Africa
+(Table. 1), and high power to discriminate between sites where \emph{D.
+rubiformis} had established or not (AUC: 0.79 - 0.91) (Fig. 4). The
+predicted climatic suitability for \emph{D. rubiformis} across South
+Africa did vary somewhat depending on the MaxEnt settings used to
+calibrate the model (Fig. 5). All MaxEnt models predicted relatively
+high climatic suitability for \emph{D. rubiformis} in the far south-west
+of South Africa and in the southern Cape region. However, the four
+MaxEnt models calibrated with tuned model settings differed from the
+default settings model by predicting climatic suitability for \emph{D.
+rubiformis} along the entire south and east Coast of South Africa.
 
@@
 climatic suitability scores for \emph{D. rubiformis} over the vast
-majority of South Africa (Fig. S1). Indeed, 54 of the 78 field sites
+majority of South Africa (Fig. 6a). Indeed, 54 of the 78 field sites
 surveyed for \emph{D. rubiformis} in South Africa occur in geographic
@@
 rubiformis} established at 23 of 24 sites in interpolation (96\%), while
-it established at 44 of 54 sites in extrapolation (82\%) (Fig. 6). We
+it established at 44 of 54 sites in extrapolation (82\%) (Fig. 6b). We
 found no evidence that a site being in extrapolation/interpolation was a
@@
 \emph{P} = 0.07). Moreover, classifying sites as being in
-extrapolation/interpolation yielded lower power to discriminate to
-discriminate between sites where \emph{D. rubiformis} had established or
-not (AUC: 0.67 \(\pm\) 0.02) than all of the MaxEnt models, barring the
-model with default settings.
+extrapolation/interpolation yielded lower power to discriminate between
+sites where \emph{D. rubiformis} had established or not (AUC: 0.63) than
+all of the MaxEnt models.
 
@@
 estimates of climatic suitability for \emph{D. rubiformis} in South
-Africa. AUC values greater than 0.80 are typically considered good
+Africa. AUC values greater than 0.75 are typically considered good
 fitting models (having high predictive power) (Fielding and Bell, 1997).
@@
 rubiformis} from Australia, were projected over South Africa, they had
-AUC scores of 0.83 - 0.84, indicative of high power to discriminate
+AUC scores of 0.89 - 0.92, indicative of high power to discriminate
 between sites where \emph{D. rubiformis} had established in South Africa
@@
 projected into novel geographic regions. While the default MaxEnt model
-had the highest predictive power in the training area (Australia), this
-model performed extremely poorly when transferred over South Africa. In
-contrast, the four MaxEnt models calibrated with tuned settings showed
-high predictive power when transferred. The poor transferability of the
-Default model is likely explained by default settings overfitting the
-model on the training points (Warren et al., 2014). This finding
-corroborates other studies showing that high predictive power in the
-model training area is not a guarantee of transferability (Duque-Lazo et
-al., 2016; Warren et al., 2014). We echo that recommendations of
-Shcheglovitova and Anderson (2013) and Muscarella et al. (2014) that
-model tuning should be performed to estimate optimal model complexity
-when transferring MaxEnt models in space and/or time in order to
-maximise their usefulness.
+had the high predictive power in the training area (Australia), this
+model was somewhat overfit to the training data (Warren et al., 2014).
+This finding corroborates other studies showing that high predictive
+power in the model training area is not a guarantee of transferability
+(Duque-Lazo et al., 2016; Warren et al., 2014). We echo that
+recommendations of Shcheglovitova and Anderson (2013) and Muscarella et
+al. (2014) that model tuning should be performed to estimate optimal
+model complexity when transferring MaxEnt models in space and/or time in
+order to maximise their usefulness.
 
@@
 only a few studies to date validating MaxEnt models using spatially
-independent data. However, the accuracy of the MaxEnt models was
-dependent on the settings used to calibrate the models, with default
-MaxEnt programme settings having low predictive power. Our results
-demonstrate that MaxEnt can be useful in predicting where a species may
-establish, even when projected into a novel geographic area, but that
-users need to be aware of how model settings can influence model outputs
-and usefulness. Demonstrating how SDM's can be useful, while
-communicating uncertainty in their predictions and outputs, is essential
-for building confidence in the use of SDM's (Blasi et al.~2020). We
-showed how validating predictive models can be used to guide the
-implementation of a biological control programme by optimising the
-redistribution and release of \emph{D. rubiformis} in geographic areas
-that are climatically suitable and where the probability of the insect
+independent data. However, the accuracy of the MaxEnt models and their
+climatic suitability projections were dependent on the settings used to
+calibrate the models. Our results demonstrate that MaxEnt can be useful
+in predicting where a species may establish, even when projected into a
+novel geographic area, but that users need to be aware of how model
+settings and model extrapolation can influence model outputs and
+usefulness. Demonstrating how SDM's can be useful, while communicating
+uncertainty in their predictions and outputs, is essential for building
+confidence in the use of SDM's (Blasi et al.~2020). We showed how
+validating predictive models can be used to guide the implementation of
+a biological control programme by optimising the redistribution and
+release of \emph{D. rubiformis} in geographic areas that are
+climatically suitable and where the probability of the insect
 establishing is greater than 50\%.
@@
 
+\includegraphics{C:/Users/s1900332/OneDrive - Rhodes University/Manuscripts/MS-Dasinuera-MaxEnt-Transfer/./ms_body/ms_figs/fig_1_distributions.png}
+
 \newpage
@@
 
+\includegraphics{C:/Users/s1900332/OneDrive - Rhodes University/Manuscripts/MS-Dasinuera-MaxEnt-Transfer/./ms_body/ms_figs/fig_2_model_tuning.png}
+
 \newpage
@@
 
+\includegraphics{C:/Users/s1900332/OneDrive - Rhodes University/Manuscripts/MS-Dasinuera-MaxEnt-Transfer/./ms_body/ms_figs/fig_3_logits.png}
+
 \newpage
@@
 
+\includegraphics{C:/Users/s1900332/OneDrive - Rhodes University/Manuscripts/MS-Dasinuera-MaxEnt-Transfer/./ms_body/ms_figs/fig_4_auc.png}
+
 \newpage
@@
 
+\includegraphics{C:/Users/s1900332/OneDrive - Rhodes University/Manuscripts/MS-Dasinuera-MaxEnt-Transfer/./ms_body/ms_figs/fig_5_maxent_maps.png}
+
 \newpage
@@
 
+\includegraphics{C:/Users/s1900332/OneDrive - Rhodes University/Manuscripts/MS-Dasinuera-MaxEnt-Transfer/./ms_body/ms_figs/fig_6_mess.png}
+
 \newpage
@@
 \begin{TableNotes}
-\item[a] Adjusted to indicate change in Y for each 1\% unit increase in covariate
+\item[a] Adjusted to indicate change in Y for each 1\% unit increase in climatic similarity
 \end{TableNotes}
@@
 \endlastfoot
-\cellcolor{gray!6}{Default settings} & \cellcolor{gray!6}{1.02} & \cellcolor{gray!6}{1.77} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0.18}\\
-AUCtest & 1.07 & 13.00 & 1 & < 0.001\\
-\cellcolor{gray!6}{AUCdiff} & \cellcolor{gray!6}{1.14} & \cellcolor{gray!6}{14.95} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{< 0.001}\\
-OR10 & 1.10 & 14.38 & 1 & < 0.001\\
-\cellcolor{gray!6}{AICc} & \cellcolor{gray!6}{1.09} & \cellcolor{gray!6}{14.38} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{< 0.001}\\*
+\cellcolor{gray!6}{Default settings} & \cellcolor{gray!6}{1.11} & \cellcolor{gray!6}{26.21} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{< 0.001}\\
+AUCtest & 1.04 & 17.50 & 1 & < 0.001\\
+\cellcolor{gray!6}{AUCdiff} & \cellcolor{gray!6}{1.03} & \cellcolor{gray!6}{12.29} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{< 0.001}\\
+OR10 & 1.02 & 9.01 & 1 & 0.003\\
+\cellcolor{gray!6}{AICc} & \cellcolor{gray!6}{1.04} & \cellcolor{gray!6}{16.21} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{< 0.001}\\*
 \end{longtable}
@@
 
-\begin{longtable}[l]{llllll}
+\begin{longtable}[l]{lllllll}
 \toprule
-Model & RM & FC & AUC\textsubscript{test} & AUC\textsubscript{diff} & OR\textsubscript{10}\\
+Model & RM & FC & AUC\textsubscript{test} & AUC\textsubscript{diff} & OR\textsubscript{10} & AIC\textsubscript{c}\\
 \midrule
 \endfirsthead
-\multicolumn{6}{@{}l}{\textit{(continued)}}\\
+\multicolumn{7}{@{}l}{\textit{(continued)}}\\
 \toprule
-Model & RM & FC & AUC\textsubscript{test} & AUC\textsubscript{diff} & OR\textsubscript{10}\\
+Model & RM & FC & AUC\textsubscript{test} & AUC\textsubscript{diff} & OR\textsubscript{10} & AIC\textsubscript{c}\\
 \midrule
@@
 \endlastfoot
-\cellcolor{gray!6}{Default} & \cellcolor{gray!6}{1.0} & \cellcolor{gray!6}{HLQPT} & \cellcolor{gray!6}{0.94} & \cellcolor{gray!6}{0.03} & \cellcolor{gray!6}{0.15}\\
-AUCtest & 1.5 & HLQPT & 0.95 & 0.02 & 0.15\\
-\cellcolor{gray!6}{AUCdiff} & \cellcolor{gray!6}{4.0} & \cellcolor{gray!6}{HLQPT} & \cellcolor{gray!6}{0.94} & \cellcolor{gray!6}{0.01} & \cellcolor{gray!6}{0.14}\\
-OR10 & 6.0 & HLQ & 0.92 & 0.02 & 0.10\\
-\cellcolor{gray!6}{AICc} & \cellcolor{gray!6}{4.0} & \cellcolor{gray!6}{H} & \cellcolor{gray!6}{0.94} & \cellcolor{gray!6}{0.02} & \cellcolor{gray!6}{0.12}\\*
+\cellcolor{gray!6}{Default} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{HLQPT} & \cellcolor{gray!6}{0.89} & \cellcolor{gray!6}{0.04} & \cellcolor{gray!6}{0.18} & \cellcolor{gray!6}{564.46}\\
+AUCtest & 6 & H & 0.89 & 0.02 & 0.09 & 15.56\\
+\cellcolor{gray!6}{AUCdiff} & \cellcolor{gray!6}{4} & \cellcolor{gray!6}{HLQPT} & \cellcolor{gray!6}{0.89} & \cellcolor{gray!6}{0.02} & \cellcolor{gray!6}{0.13} & \cellcolor{gray!6}{485.91}\\
+OR10 & 6 & HLQPT & 0.92 & 0.02 & 0.12 & 144.39\\
+\cellcolor{gray!6}{AICc} & \cellcolor{gray!6}{5} & \cellcolor{gray!6}{H} & \cellcolor{gray!6}{0.90} & \cellcolor{gray!6}{0.02} & \cellcolor{gray!6}{0.09} & \cellcolor{gray!6}{0.00}\\*
 \end{longtable}

### 2022-02-08 R/scripts/figure-7-make.R

@@ -1,0 +1,142 @@
+# -----------------------------------------------------------------------------
+# Script: - Reproduce Figure 7
+#
+# AUTHOR: Guy F. Sutton
+# AFFILIATION: Centre for Biological Control, Rhodes University, South Africa
+# DATE MODIFIED: 03/02/2021
+# CONTACT: g.sutton@ru.ac.za
+# -----------------------------------------------------------------------------
+
+# -----------------------------------------------------------------------------
+# Session setup
+# -----------------------------------------------------------------------------
+
+# Load packages
+if (!require("pacman"))
+  install.packages("pacman")
+pacman::p_load(
+  tidyverse,
+  spocc,
+  ggmap,
+  maptools,
+  maps,
+  ggplot2,
+  scrubr,
+  mapr,
+  tidyr,
+  stringr,
+  patchwork,
+  rnaturalearth,
+  rnaturalearthdata,
+  rlang,
+  sf,
+  ggspatial,
+  raster,
+  here,
+  tidytext,
+  ggtext,
+  tidymodels,
+  themis,
+  ranger,
+  viridis,
+  yardstick,
+  vip,
+  DHARMa,
+  rms
+)
+
+# Set the theme for the figures
+theme_set(
+  theme_classic() +
+    theme(
+      panel.border = element_rect(colour = "black", fill = NA),
+      axis.text = element_text(colour = "black"),
+      axis.title.x = element_text(margin = unit(c(2, 0, 0, 0), "mm")),
+      axis.title.y = element_text(margin = unit(c(0, 4, 0, 0), "mm")),
+      legend.position = "none"
+    )
+)
+
+# Set the theme for the maps
+theme_opts <- list(
+  theme(
+    panel.grid.minor = element_blank(),
+    panel.grid.major = element_blank(),
+    panel.background = element_rect(fill = 'white', colour = NA),
+    plot.background = element_rect(),
+    axis.line = element_blank(),
+    axis.text.x = element_text(colour = "black"),
+    axis.text.y = element_text(colour = "black"),
+    axis.ticks = element_line(colour = "black"),
+    axis.title.x = element_text(colour = "black"),
+    axis.title.y = element_text(colour = "black"),
+    plot.title = element_text(colour = "black"),
+    panel.border = element_rect(fill = NA),
+    legend.key = element_blank()
+  )
+)
+
+# ----------------------------------------------------------------------------
+# Plot map: 
[…]

### 2022-02-08 ms_body/manuscript.Rmd
Editing MS - add Fig. 7, add new refs

@@
 However, a number of authors have called in question the usefulness of SDM's, notably when models are transferred in space or time [@Yates2018; @Blasi2021].
-Poor transferability may lead to erroneous model predictions that may have serious negative consequences for the application of the model, and resulting policies and management programmes (Soultan and Safi, 2017).
+Poor transferability may lead to erroneous model predictions that may have serious negative consequences for the application of the model, and resulting policies and management programmes.
 Here, we report on the validation of a SDM for an insect, (*D.
 rubiformis*), native to Australia, that has been intentionally introduced as a biological control agent into South Africa to reduce the invasiveness of its host-plant, *A.
-mearnsii* (Impson et al., 2013).
+mearnsii* [@Impson2013].
 The the availability of independent testing data from South Africa, a separate continent to the data used to generate the models (Australia), makes this an ideal study system for assessing the transferability of SDM’s.
@@
 
-An additional concern with transferring MaxEnt models in space or time is the potential for models to significantly extrapolate predictions when the model is projected into non-analogous conditions to those under which the model was calibrated (Elith et al., 2010; Mesgaran et al., 2014).
-However, the applied use of SDM's in developing environmental policy, developing management programmes for invasive species and implementing conservation programmes often necessitate that models are transferred and predictions made, even when models are extrapolating (Elith et al., 2010).
+An additional concern with transferring MaxEnt models in space or time is the potential for models to significantly extrapolate predictions when the model is projected into non-analogous conditions to those under which the model was calibrated [@Elith2010; @Mesgaran2014].
+However, the applied use of SDM's in developing environmental policy, developing management programmes for invasive species and implementing conservation programmes often necessitate that models are transferred and predictions made, even when models are extrapolating [@Elith2010].
 Here, we showed that *D.
@@
 A number of tools have been developed in recent years to assist SDM users with identifying model extrapolation, e.g.
-MESS [@Elith2010] and ExDet (Mesgaran et al., 2014), and which should be an integral component of any SDM study.
+MESS [@Elith2010] and ExDet [@Mesgaran2014], and which should be an integral component of any SDM study.
 
@@
 rubiformis*.
-Mass-releasing and releasing weed biological control agents is a time-consuming and costly endeavor (Coetzee et al., 2021).
+Mass-releasing and releasing weed biological control agents is a time-consuming and costly endeavor [@Hill2021].
 As such, our models provide a statistically validated approach to optimising the redistribution of *D.

### 2022-02-08 ms_body/manuscript.tex
Editing MS - add Fig. 7, add new refs

@@
 for the application of the model, and resulting policies and management
-programmes (Soultan and Safi, 2017). Here, we report on the validation
-of a SDM for an insect, (\emph{D. rubiformis}), native to Australia,
-that has been intentionally introduced as a biological control agent
-into South Africa to reduce the invasiveness of its host-plant, \emph{A.
-mearnsii} (Impson et al., 2013). The the availability of independent
-testing data from South Africa, a separate continent to the data used to
-generate the models (Australia), makes this an ideal study system for
-assessing the transferability of SDM's. This study represents one of
-only a handful of studies to validate SDM's using independent testing
-data (Costa et al., 2010; Rebelo and Jones, 2010; Smith et al., 2021;
-West et al., 2016).
+programmes. Here, we report on the validation of a SDM for an insect,
+(\emph{D. rubiformis}), native to Australia, that has been intentionally
+introduced as a biological control agent into South Africa to reduce the
+invasiveness of its host-plant, \emph{A. mearnsii} (Impson et al.,
+2013). The the availability of independent testing data from South
+Africa, a separate continent to the data used to generate the models
+(Australia), makes this an ideal study system for assessing the
+transferability of SDM's. This study represents one of only a handful of
+studies to validate SDM's using independent testing data (Costa et al.,
+2010; Rebelo and Jones, 2010; Smith et al., 2021; West et al., 2016).
 
@@
 Africa, we thresholded the MaxEnt map that optimised AICc where the
-probability of establishment was \textgreater{} 50\% (i.e.~MaxEnt score
-= 0.30), given in Fig. 9. The map demonstrates that the entire south and
+probability of establishment was \textgreater{} 75\% (i.e.~MaxEnt score
+= 0.30), given in Fig. 7. The map demonstrates that the entire south and
 east coast regions of South Africa are climatically suitable for
 \emph{D. rubiformis}. Mass-releasing and releasing weed biological
-control agents is a time-consuming and costly endeavor (Coetzee et al.,
+control agents is a time-consuming and costly endeavor (Hill et al.,
 2021). As such, our models provide a statistically validated approach to
@@
 \textgreater{} 0.30, and therefore where the probability of
-establishment was \textgreater{} 50\%.
+establishment was \textgreater{} 75\%.
 
@@
 climatically suitable and where the probability of the insect
-establishing is greater than 50\%.
+establishing is greater than 75\%.
 
@@
 
+\leavevmode\hypertarget{ref-Hill2021}{}%
+Hill, M., Conlong, D., Zachariades, C., Coetzee, J., Paterson, I.,
+Miller, B., Foxcroft, L., Van Der Westhuizen, L., 2021. The role of
+mass-rearing in weed biological control projects in {South Africa}.
+African Entomology 29, 1030--1044.
+
 \leavevmode\hypertarget{ref-Impson2013}{}%
@@
 \url{https://doi.org/10.1111/j.1600-0587.2013.07872.x}
+
+\leavevmode\hypertarget{ref-Mesgaran2014}{}%
+Mesgaran, M., Cousens, R., Webber, B., 2014. Here be dragons: {A} tool
+for quantifying novelty due to covariate range and correlation change
+when projecting species distribution models. Diversity and Distributions
+20, 1147--1159. \url{https://doi.org/10.1111/ddi.12209}
 
@@
 \cellcolor{gray!6}{AUCdiff} & \cellcolor{gray!6}{4} & \cellcolor{gray!6}{HLQPT} & \cellcolor{gray!6}{0.89} & \cellcolor{gray!6}{0.02} & \cellcolor{gray!6}{0.13} & \cellcolor{gray!6}{485.91}\\
-OR10 & 6 & HLQPT & 0.92 & 0.02 & 0.12 & 144.39\\
+OR10 & 6 & HLQPT & 0.92 & 0.02 & 0.11 & 144.39\\
 \cellcolor{gray!6}{AICc} & \cellcolor{gray!6}{5} & \cellcolor{gray!6}{H} & \cellcolor{gray!6}{0.90} & \cellcolor{gray!6}{0.02} & \cellcolor{gray!6}{0.09} & \cellcolor{gray!6}{0.00}\\*

### 2022-02-15 ms_body/manuscript.Rmd
Update Abstract
Add acknowledgements and minor edits from FI

@@
 - \renewcommand\familydefault{\sfdefault}
+- \usepackage{fullpage}
+- \usepackage{pdflscape}
 biblio-style: apsr
@@
 Our study demonstrates that MaxEnt models can prove useful when transferred in space, however, users need to be aware of the effect of different model settings configurations and model extrapolation can have on model transferability and usefulness in applied ecological settings.
+We demonstrate how validating the *D.
+rubiformis* MaxEnt models can be used to guide the biological control of *A.
+mearnsii* using *D.
+rubiformis*.
 
@@
 Here, we modelled the potential climatic suitability of an insect that is native to Australia, and that has been introduced into South Africa as a biological control agent.
-The insect, *Dasineura rubiformis* Kolesik (Diptera: Cecidomyiidae) was introduced into South Africa from Australia in 2013, to curb the spread of the invasive Australian *Acacia mearnsii* De Wild [@Impson2013].
+The insect, *Dasineura rubiformis* Kolesik (Diptera: Cecidomyiidae) was introduced into South Africa from Australia in 2001, and found to be established in 2006, to curb the spread of the invasive Australian *Acacia mearnsii* De Wild [@Impson2013].
 Extensive field surveys have been performed in Australia (characterizing the native distribution of *D.
@@
 To address these concerns, Multivariate environmental similarity surfaces (hereafter ‘MESS’) [@Elith2010] were computed to assess whether MaxEnt models were extrapolating or interpolating.
-MESS analyses measure the similarity of any given point to a set of reference points [reference points were defined as only the species presence points used to calibrate models following @Kriticos2014].
+MESS analyses measure the similarity of any given point to a set of reference points (reference points were defined as only the species presence points used to calibrate models following @Kriticos2014).
 Negative MESS values indicate geographic regions outside the range of climate variables used to calibrate the model (i.e.
@@
 rubiformis* is used to curtail the spread of the highly invasive weed *A.
-mearnsii* expend many resources (time and money) to mass-rear insect populations that are then released into the field.
+mearnsii* expend many resources (time and money) to develop effective biological control agents.
 The positive correlation between MaxEnt suitability scores and *D.
-rubiformis* establishment means that researchers can develop and implement mass-rearing and experimental release programmes for the insect that prioritise releasing the insect at sites that are highly climatically suitable, and thus, more likely the insect will establish viable populations.
+rubiformis* establishment means that researchers can develop experimental release programmes that prioritise releasing the insect at sites that are highly climatically suitable, and thus, more likely the insect will establish viable populations.
 This may assist in improving the level of control over the target weed and increasing the efficiency of resources expended by increasing the chances of establishing viable insect populations.
@@
 7.
-The map demonstrates that the entire south and east coast regions of South Africa are climatically suitable for *D.
+The map demonstrates that the south and east coast regions of South Africa are climatically suitable for *D.
 rubiformis*.
@@
 Any opinions, finding, conclusions or recommendations expressed in this material are those of the authors and the NRF does not accept any liability in this regard.
+We thank Fiona Impson, Robin Adair, Catharina Kleinjan and John Hoffmann for collecting and providing the invaluable data, and Fiona Impson for advice on a previous draft of this manuscript.
 

### 2022-02-15 ms_body/manuscript.tex
Add acknowledgements and minor edits from FI

@@
 \renewcommand\familydefault{\sfdefault}
+\usepackage{fullpage}
+\usepackage{pdflscape}
 \usepackage{booktabs}
@@
 model settings configurations and model extrapolation can have on model
-transferability and usefulness in applied ecological settings.
+transferability and usefulness in applied ecological settings. We
+demonstrate how validating the \emph{D. rubiformis} MaxEnt models can be
+used to guide the biological control of \emph{A. mearnsii} using
+\emph{D. rubiformis}.
 
@@
 Kolesik (Diptera: Cecidomyiidae) was introduced into South Africa from
-Australia in 2013, to curb the spread of the invasive Australian
-\emph{Acacia mearnsii} De Wild (Impson et al., 2013). Extensive field
-surveys have been performed in Australia (characterizing the native
-distribution of \emph{D. rubiformis}) (Adair, 2004) and long-term
-post-release monitoring has been performed in South Africa to measure
-establishment rates of the insect, following experimental releases of
-the insect (Impson et al., 2021). As such, the availability of
-independent testing data from a separate continent to the data used to
-generate the models makes this an ideal study system for assessing the
-transferability of SDM's.
+Australia in 2001, and found to be established in 2006, to curb the
+spread of the invasive Australian \emph{Acacia mearnsii} De Wild (Impson
+et al., 2013). Extensive field surveys have been performed in Australia
+(characterizing the native distribution of \emph{D. rubiformis}) (Adair,
+2004) and long-term post-release monitoring has been performed in South
+Africa to measure establishment rates of the insect, following
+experimental releases of the insect (Impson et al., 2021). As such, the
+availability of independent testing data from a separate continent to
+the data used to generate the models makes this an ideal study system
+for assessing the transferability of SDM's.
 
@@
 (reference points were defined as only the species presence points used
-to calibrate models following Kriticos et al., 2014). Negative MESS
+to calibrate models following Kriticos et al. (2014)). Negative MESS
 values indicate geographic regions outside the range of climate
@@
 spread of the highly invasive weed \emph{A. mearnsii} expend many
-resources (time and money) to mass-rear insect populations that are then
-released into the field. The positive correlation between MaxEnt
-suitability scores and \emph{D. rubiformis} establishment means that
-researchers can develop and implement mass-rearing and experimental
-release programmes for the insect that prioritise releasing the insect
-at sites that are highly climatically suitable, and thus, more likely
-the insect will establish viable populations. This may assist in
-improving the level of control over the target weed and increasing the
-efficiency of resources expended by increasing the chances of
-establishing viable insect populations.
+resources (time and money) to develop effective biological control
+agents. The positive correlation between MaxEnt suitability scores and
+\emph{D. rubiformis} establishment means that researchers can develop
+experimental release programmes that prioritise releasing the insect at
+sites that are highly climatically suitable, and thus, more likely the
+insect will establish viable populations. This may assist in improving
+the level of control over the target weed and increasing the efficiency
+of resources expended by increasing the chances of establishing viable
+insect populations.
 
@@
 probability of establishment was \textgreater{} 75\% (i.e.~MaxEnt score
-= 0.30), given in Fig. 7. The map demonstrates that the entire south and
-east coast regions of South Africa are climatically suitable for
-\emph{D. rubiformis}. Mass-releasing and releasing weed biological
-control agents is a time-consuming and costly endeavor (Hill et al.,
-2021). As such, our models provide a statistically validated approach to
-optimising the redistribution of \emph{D. rubiformis} in South Africa by
-prioritising releases at field sites where MaxEnt scores are
-\textgreater{} 0.30, and therefore where the probability of
-establishment was \textgreater{} 75\%.
+= 0.30), given in Fig. 7. The map demonstrates that the south and east
+coast regions of South Africa are climatically suitable for \emph{D.
+rubiformis}. Mass-releasing and releasing weed biological control agents
+is a time-consuming and costly endeavor (Hill et al., 2021). As such,
+our models provide a statistically validated approach to optimising the
+redistribution of \emph{D. rubiformis} in South Africa by prioritising
+releases at field sites where MaxEnt scores are \textgreater{} 0.30, and
+therefore where the probability of establishment was \textgreater{}
+75\%.
 
@@
 conclusions or recommendations expressed in this material are those of
-the authors and the NRF does not accept any liability in this regard.
+the authors and the NRF does not accept any liability in this regard. We
+thank Fiona Impson, Robin Adair, Catharina Kleinjan and John Hoffmann
+for collecting and providing the invaluable data, and Fiona Impson for
+advice on a previous draft of this manuscript.
 

### 2022-02-16 ms_body/manuscript.Rmd
Push MS edits before submission

@@
 A total of 103 native-range records for *D.
-rubiformis* were obtained from [@Adair2004].
+rubiformis* were obtained from @Adair2004.
 A total of 452 invaded-range records to evaluate model transferability were taken from Impson, Kleinjan and Hoffmann (unpublished).
@@
 
+**Figure 1:** Distribution records for *Dasineura rubiformis* in its (a) native range (Australia), and (b) introduced range (South Africa).
+Filled black circles indicate field sites where *D.
+rubiformis* has been recorded, and filled grey circles indicate field sites where it was not found.
+
+**Figure 2:** Model tuning experiments to identify optimal MaxEnt model settings configurations for *D.
+rubiformis* native-range (Australia) models.
+Model tuning was performed across a range of regularisation multipliers (0.5 - 10), and feature classes (H = Hinge, L = Linear, Q = Quadratic).
+Model performance and optimal settings configurations were assessed using multiple metrics that reflected aspects of model (1) discriminatory ability (AUC~test~), (2) overfitting (AUC~diff~), (3) omission rates (OR~10~), and (4) overall parsimony (AICc).
+
+**Figure 3:** Relationships between MaxEnt climatic suitability scores (between the native and introduced range) and the probability of establishment for *D.
+rubiformis* in its introduced range.
+Five different MaxEnt models are compared, specified with either default MaxEnt settings, or four optimised settings configurations.
+Mean prediction curves $\pm$ 95% confidence intervals (indicated by grey-shaded ribbons) are derived from logistic General Linear Model's (GLM's).
+
+**Figure 4::** MaxEnt model validation using out-of-sample Area under the Curve (AUC) and independent testing data from the introduced geographic range of *D.
+rubiformis* in South Africa.
+Five different MaxEnt models are compared, specified with either default MaxEnt settings, or four optimised settings configurations.
+The dashed horizontal line at y = 0.75 indicates the lowest bounds of acceptable model accuracy according to Fielding and Bell (1997).
+AUC values closer to 1 indicate improved model accuracy.
+Boxplots show the results obtained from n = 500 bootstrap resamples of the fitted logistic Generali Linear Models (GLM's).
+See text for further details.
+
+**Figure 5::** Projected MaxEnt climatic similarity for *D.
+rubiformis* in its introduced range (South Africa) based on native-range MaxEnt models (Australia).
+Five different MaxEnt models are compared, specified with either default MaxEnt settings, or four optimised settings configurations.
+Climatic suitability scores closer to 1 indicate higher climatic similarity between climates in the introduced vs native-range.
+
+**Figure 6:** Comparing *D.
+rubiformis* establishment in its introduced range (South Africa) given by MaxEnt model extrapolation.
+(a) Multivariate Environmental Similarity Surfaces (MESS) for *D.
+rubiformis*.
+MESS maps were thresholded to indicate regions where MaxEnt models were extrapolating (extrapolation - light grey fill) or not (interpolation - dark grey fill).
+(b) The probability that *D.
+rubiformis* was established or not at field sites in South Africa depending on whether the field site was in interpolation/extrapolation.
+
+**Figure 7:** Projected MaxEnt climatic similarity for *D.
+rubiformis* in its introduced range (South Africa), thresholded to show geographic regions where the estimated probability of establishment for *D.
+rubiformis* was greater than 75% (i.e.
+MaxEnt score = 0.30, taken from Figure 4) (Suitable regions - dark grey fill) or not (unsuitable regions - light grey fill).
+
+**Table 1:** Parameter estimates and Likelihood Ratio Tests (LRT's) comparing the relationships between MaxEnt climatic suitability scores (between the native and introduced range) and the probability of establishment for *D.
+rubiformis* in its introduced range.
+Five different MaxEnt models are compared, specified with either default MaxEnt settings, or four optimised settings configurations.
+$\\beta$\\textsuperscript{a} indicate the factor by which the probability of *D.
+rubiformis* establishment increases in South Africa for a 1% increase in climatic similarity between South Africa and its native-range (Australia).
+LRT's compare logistic General Linear Model's (GLM's) fitted with the climatic match score as a contiunous covariate and *D.
+rubiformis* establishment as a Boolean response variable against an analogous GLM fitted with a random intercept term only.
+
 \newpage
@@
 
-```{r import_fig_4, echo = FALSE}
+```{r import_fig_4, out.width="80%", out.height="80%",  echo = FALSE}
 knitr::include_graphics(here::here("./ms_body/ms_figs/fig_4_auc.png"), dpi = NA)
@@
 
-```{r import_fig_5, echo = FALSE}
+```{r import_fig_5, out.width="100%", out.height="100%", echo = FALSE}
 knitr::include_graphics(here::here("./ms_body/ms_figs/fig_5_maxent_maps.png"), dpi = NA)
@@
 knitr::include_graphics(here::here("./ms_body/ms_figs/fig_6_mess.png"), dpi = NA)
+```
+
+\newpage
+
+## *Figure 7*
+
+```{r import_fig_7, echo = FALSE}
+knitr::include_graphics(here::here("./ms_body/ms_figs/fig_7_biocontrol.png"), dpi = NA)
 ```
@@
 ```
-
-\newpage
-
-**Table S1**
-
- \
-
-```{r table_S1, echo = FALSE}
-# Import data to make Table S1
-tableS1_data <- readr::read_csv(here::here("./data/data_clean/table2_data.csv")) %>%
-  janitor::clean_names()
-
-# Create Table S1
-tableS1 <- tableS1_data %>%
-  knitr::kable(
-    format = "latex",
-    align = "l",
-    booktabs = TRUE,
-    longtable = TRUE,
-    linesep = "",
-    escape = FALSE,
-    col.names = c(
-      "Model",
-      "RM",
-      "FC",
-      "AUC\\textsubscript{test}",
-      "AUC\\textsubscript{diff}",
-      "OR\\textsubscript{10}",
-      "AIC\\textsubscript{c}"
-    )
-  ) %>%
-  kableExtra::kable_styling(
-    position = "left",
-    latex_options = c("striped", "repeat_header")
-    #,    stripe_color = "gray!15"
-  )
-
-# Display Table S1
-tableS1
-```

### 2022-02-16 ms_body/manuscript.tex
Push MS edits before submission

@@
 A total of 103 native-range records for \emph{D. rubiformis} were
-obtained from (Adair, 2004). A total of 452 invaded-range records to
+obtained from Adair (2004). A total of 452 invaded-range records to
 evaluate model transferability were taken from Impson, Kleinjan and
@@
 
+\textbf{Figure 1:} Distribution records for \emph{Dasineura rubiformis}
+in its (a) native range (Australia), and (b) introduced range (South
+Africa). Filled black circles indicate field sites where \emph{D.
+rubiformis} has been recorded, and filled grey circles indicate field
+sites where it was not found.
+
+\textbf{Figure 2:} Model tuning experiments to identify optimal MaxEnt
+model settings configurations for \emph{D. rubiformis} native-range
+(Australia) models. Model tuning was performed across a range of
+regularisation multipliers (0.5 - 10), and feature classes (H = Hinge, L
+= Linear, Q = Quadratic). Model performance and optimal settings
+configurations were assessed using multiple metrics that reflected
+aspects of model (1) discriminatory ability (AUC\textsubscript{test}),
+(2) overfitting (AUC\textsubscript{diff}), (3) omission rates
+(OR\textsubscript{10}), and (4) overall parsimony (AICc).
+
+\textbf{Figure 3:} Relationships between MaxEnt climatic suitability
+scores (between the native and introduced range) and the probability of
+establishment for \emph{D. rubiformis} in its introduced range. Five
+different MaxEnt models are compared, specified with either default
+MaxEnt settings, or four optimised settings configurations. Mean
+prediction curves \(\pm\) 95\% confidence intervals (indicated by
+grey-shaded ribbons) are derived from logistic General Linear Model's
+(GLM's).
+
+\textbf{Figure 4::} MaxEnt model validation using out-of-sample Area
+under the Curve (AUC) and independent testing data from the introduced
+geographic range of \emph{D. rubiformis} in South Africa. Five different
+MaxEnt models are compared, specified with either default MaxEnt
+settings, or four optimised settings configurations. The dashed
+horizontal line at y = 0.75 indicates the lowest bounds of acceptable
+model accuracy according to Fielding and Bell (1997). AUC values closer
+to 1 indicate improved model accuracy. Boxplots show the results
+obtained from n = 500 bootstrap resamples of the fitted logistic
+Generali Linear Models (GLM's). See text for further details.
+
+\textbf{Figure 5::} Projected MaxEnt climatic similarity for \emph{D.
+rubiformis} in its introduced range (South Africa) based on native-range
+MaxEnt models (Australia). Five different MaxEnt models are compared,
+specified with either default MaxEnt settings, or four optimised
+settings configurations. Climatic suitability scores closer to 1
+indicate higher climatic similarity between climates in the introduced
+vs native-range.
+
+\textbf{Figure 6:} Comparing \emph{D. rubiformis} establishment in its
+introduced range (South Africa) given by MaxEnt model extrapolation. (a)
+Multivariate Environmental Similarity Surfaces (MESS) for \emph{D.
+rubiformis}. MESS maps were thresholded to indicate regions where MaxEnt
+models were extrapolating (extrapolation - light grey fill) or not
+(interpolation - dark grey fill). (b) The probability that \emph{D.
+rubiformis} was established or not at field sites in South Africa
+depending on whether the field site was in interpolation/extrapolation.
+
+\textbf{Figure 7:} Projected MaxEnt climatic similarity for \emph{D.
+rubiformis} in its introduced range (South Africa), thresholded to show
+geographic regions where the estimated probability of establishment for
+\emph{D. rubiformis} was greater than 75\% (i.e.~MaxEnt score = 0.30,
+taken from Figure 4) (Suitable regions - dark grey fill) or not
+(unsuitable regions - light grey fill).
+
+\textbf{Table 1:} Parameter estimates and Likelihood Ratio Tests (LRT's)
+comparing the relationships between MaxEnt climatic suitability scores
+(between the native and introduced range) and the probability of
+establishment for \emph{D. rubiformis} in its introduced range. Five
+different MaxEnt models are compared, specified with either default
+MaxEnt settings, or four optimised settings configurations.
+\(\\beta\)\textbackslash textsuperscript\{a\} indicate the factor by
+which the probability of \emph{D. rubiformis} establishment increases in
+South Africa for a 1\% increase in climatic similarity between South
+Africa and its native-range (Australia). LRT's compare logistic General
+Linear Model's (GLM's) fitted with the climatic match score as a
+contiunous covariate and \emph{D. rubiformis} establishment as a Boolean
+response variable against an analogous GLM fitted with a random
+intercept term only.
+
 \newpage
@@
 
-\includegraphics{C:/Users/s1900332/OneDrive - Rhodes University/Manuscripts/MS-Dasinuera-MaxEnt-Transfer/./ms_body/ms_figs/fig_4_auc.png}
+\includegraphics[width=0.8\linewidth,height=0.8\textheight]{C:/Users/s1900332/OneDrive - Rhodes University/Manuscripts/MS-Dasinuera-MaxEnt-Transfer/./ms_body/ms_figs/fig_4_auc}
 
@@
 
-\includegraphics{C:/Users/s1900332/OneDrive - Rhodes University/Manuscripts/MS-Dasinuera-MaxEnt-Transfer/./ms_body/ms_figs/fig_5_maxent_maps.png}
+\includegraphics[width=1\linewidth,height=1\textheight]{C:/Users/s1900332/OneDrive - Rhodes University/Manuscripts/MS-Dasinuera-MaxEnt-Transfer/./ms_body/ms_figs/fig_5_maxent_maps}
 
@@
 \includegraphics{C:/Users/s1900332/OneDrive - Rhodes University/Manuscripts/MS-Dasinuera-MaxEnt-Transfer/./ms_body/ms_figs/fig_6_mess.png}
+
+\newpage
+
+\hypertarget{figure-7}{%
+\subsection{\texorpdfstring{\emph{Figure 7}}{Figure 7}}\label{figure-7}}
+
+\includegraphics{C:/Users/s1900332/OneDrive - Rhodes University/Manuscripts/MS-Dasinuera-MaxEnt-Transfer/./ms_body/ms_figs/fig_7_biocontrol.png}
 
@@
 \newpage
-
-\textbf{Table S1}
-
-~
-
-\begin{longtable}[l]{lllllll}
-\toprule
-Model & RM & FC & AUC\textsubscript{test} & AUC\textsubscript{diff} & OR\textsubscript{10} & AIC\textsubscript{c}\\
-\midrule
-\endfirsthead
-\multicolumn{7}{@{}l}{\textit{(continued)}}\\
-\toprule
-Model & RM & FC & AUC\textsubscript{test} & AUC\textsubscript{diff} & OR\textsubscript{10} & AIC\textsubscript{c}\\
-\midrule
-\endhead
-
-\endfoot
-\bottomrule
-\endlastfoot
-\cellcolor{gray!6}{Default} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{HLQPT} & \cellcolor{gray!6}{0.89} & \cellcolor{gray!6}{0.04} & \cellcolor{gray!6}{0.18} & \cellcolor{gray!6}{564.46}\\
-AUCtest & 6 & H & 0.89 & 0.02 & 0.09 & 15.56\\
-\cellcolor{gray!6}{AUCdiff} & \cellcolor{gray!6}{4} & \cellcolor{gray!6}{HLQPT} & \cellcolor{gray!6}{0.89} & \cellcolor{gray!6}{0.02} & \cellcolor{gray!6}{0.13} & \cellcolor{gray!6}{485.91}\\
-OR10 & 6 & HLQPT & 0.92 & 0.02 & 0.11 & 144.39\\
-\cellcolor{gray!6}{AICc} & \cellcolor{gray!6}{5} & \cellcolor{gray!6}{H} & \cellcolor{gray!6}{0.90} & \cellcolor{gray!6}{0.02} & \cellcolor{gray!6}{0.09} & \cellcolor{gray!6}{0.00}\\*
-\end{longtable}
-
-\newpage
 \singlespacing 

### 2022-02-16 R/scripts/figure-5-make.R

@@ -391,7 +391,7 @@ all_maps
 # Save the plot to your PC
 ggsave(
   "./ms_body/ms_figs/fig_5_maxent_maps.png",
-  width = 18,
-  height = 12,
+  width = 24, # 18
+  height = 16, # 12
   dpi = 600
 )

### 2022-02-16 ms_body/ms_tables/script-make-table-s1.R

@@ -1,0 +1,44 @@
+
+\newpage
+
+**Table S1**  
+  
+  \  
+
+```{r table_S1, echo = FALSE}
+# Import data to make Table S1
+tableS1_data <- readr::read_csv(here::here("./data/data_clean/table2_data.csv")) %>%
+  janitor::clean_names()
+
+
+# Create Table S1
+tableS1 <- tableS1_data %>%
+  knitr::kable(
+    format = "latex",
+    align = "l",
+    booktabs = TRUE,
+    longtable = TRUE,
+    linesep = "",
+    escape = FALSE,
+    col.names = c(
+      "Model",
+      "RM",
+      "FC",
+      "AUC\\textsubscript{test}",
+      "AUC\\textsubscript{diff}",
+      "OR\\textsubscript{10}",
+      "AIC\\textsubscript{c}"
+    )
+  ) %>%
+  kableExtra::kable_styling(
+    position = "left",
+    latex_options = c("striped", "repeat_header")
+    #,    stripe_color = "gray!15"
+  )
+
+# Display Table S1
+tableS1
+```
+
+
+
